### 2013-04-28 Documents/ms.tex
Writing

\documentclass[letterpaper, 11pt]{article}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage[left=3cm,top=3cm,right=3cm]{geometry}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\parindent=0cm

\title{Document Title}
\author{Brendon J. Brewer}

\begin{document}
\maketitle

\section{Introduction}

\section{Spatial Distribution}
Suppose the two uniform variables are $u_1$ and $u_2$.
These get turned into an elliptical thing by the following
one-to-one transformation:
\begin{eqnarray}
x &=& x_c + \frac{\mathcal{R}(u_1)}{\sqrt{q}}\cos\left(2\pi u_2\right)\\
y &=& y_c + \mathcal{R}(u_1)\sqrt{q}\sin\left(2\pi u_2\right)
\end{eqnarray}
where $\mathcal{R}(u)$ is some function that determines the
radial profile. The probability density for $x$ and $y$ is given by:
\begin{eqnarray}
p(x, y)
&=& \frac{p(u_1, u_2)}{\left|\det \mathbf{J}\right|}
\end{eqnarray}
where $\mathbf{J}$ is the Jacobian of the transformation:
\begin{eqnarray}
\mathbf{J} &=&
\left(
\begin{array}{cc}
\frac{\partial x}{\partial u_1} & \frac{\partial x}{\partial u_2}\\
\frac{\partial y}{\partial u_1} & \frac{\partial y}{\partial u_2}
\end{array}
\right)
\end{eqnarray}
Let's calculate the terms. Here's the first one:
\begin{eqnarray}
\frac{\partial x}{\partial u_1} &=& \frac{\mathcal{R}'(u_1)}{\sqrt{q}}\cos\left(2\pi u_2\right)
\end{eqnarray}
and the second:
\begin{eqnarray}
\frac{\partial x}{\partial u_2} &=& \frac{\mathcal{R}(u_1)}{\sqrt{q}}\cos\left(2\pi u_2\right)
\end{eqnarray}

\end{document}

### 2013-04-28 Documents/ms.tex
The derivatives

@@
 Suppose the two uniform variables are $u_1$ and $u_2$.
-These get turned into an elliptical thing by the following
-one-to-one transformation:
+These get turned into an elliptical thing 
+(centered at zero, aligned along the $x$-axis)
+by the following one-to-one transformation:
 \begin{eqnarray}
-x &=& x_c + \frac{\mathcal{R}(u_1)}{\sqrt{q}}\cos\left(2\pi u_2\right)\\
-y &=& y_c + \mathcal{R}(u_1)\sqrt{q}\sin\left(2\pi u_2\right)
+x &=& \frac{\mathcal{R}(u_1)}{\sqrt{q}}\cos\left(2\pi u_2\right)\\
+y &=& \mathcal{R}(u_1)\sqrt{q}\sin\left(2\pi u_2\right)
 \end{eqnarray}
 where $\mathcal{R}(u)$ is some function that determines the
-radial profile. The probability density for $x$ and $y$ is given by:
+radial profile. So far I am using an exponential distribution:
+\begin{eqnarray}
+\mathcal{R}(u) = -\mu\log(1 - u).
+\end{eqnarray}
+The probability density for $x$ and $y$ is given by:
 \begin{eqnarray}
@@
 \end{eqnarray}
-Let's calculate the terms. Here's the first one:
+Let's calculate the terms. Here they are:
 \begin{eqnarray}
-\frac{\partial x}{\partial u_1} &=& \frac{\mathcal{R}'(u_1)}{\sqrt{q}}\cos\left(2\pi u_2\right)
-\end{eqnarray}
-and the second:
-\begin{eqnarray}
-\frac{\partial x}{\partial u_2} &=& \frac{\mathcal{R}(u_1)}{\sqrt{q}}\cos\left(2\pi u_2\right)
+\frac{\partial x}{\partial u_1} &=& \frac{\mathcal{R}'(u_1)}{\sqrt{q}}\cos\left(2\pi u_2\right)\\
+\frac{\partial x}{\partial u_2} &=& -\frac{2\pi\mathcal{R}(u_1)}{\sqrt{q}}\sin\left(2\pi u_2\right)\\
+\frac{\partial y}{\partial u_1} &=& \mathcal{R}'(u_1)\sqrt{q}\sin\left(2\pi u_2\right)\\
+\frac{\partial y}{\partial u_2} &=&  2\pi\mathcal{R}(u_1)\sqrt{q}\cos\left(2\pi u_2\right)
 \end{eqnarray}

### 2013-04-28 main.cpp
Save more often

@@ -28,7 +28,7 @@ int main()
 	fstream fout("output.txt", ios::out);
 
 	// How many MCMC steps to do
-	int steps = 10000;
+	int steps = 5000;
 
 	for(int i=0; i<steps; i++)
 	{
@@ -40,7 +40,7 @@ int main()
 		if(randomU() <= exp(logH))
 			r = r2;
 
-		if(i%100 == 0)
+		if(i%5 == 0)
 		{
 			r.print(fout);
 			fout<<endl;

### 2013-04-28 display.py
More transparent dots

@@ -16,7 +16,7 @@ for i in xrange(0, output.shape[0]):
 	m = m[m > 0.]
 
 	if x.size > 0:
-		plt.scatter(x, y, s=100.*m/m.max(), alpha=0.25, color='k')
+		plt.scatter(x, y, s=100.*m/m.max(), alpha=0.1, color='k')
 	plt.axis('scaled')
 	plt.axis([-1., 1., -1., 1.])
 	plt.title('{i}/{N}'.format(i=(i+1), N=output.shape[0]))

### 2013-04-28 SpatialDistributions/Elliptical2D.cpp
I think that's the right log density

@@ -83,28 +83,7 @@ double Elliptical2D::position_log_pdf(const vector<double>& position) const
 {
 	vector<double> vec = position;
 	position_to_uniform(vec);
-
-	double r = -mu_r*log(1. - vec[0]);
-	double phi = 2.*M_PI*vec[1];
-
-	double J11, J12, J21, J22;
-	J11 = (1./sqrt(q)*cos_theta*cos(phi) - sqrt(q)*sin_theta*sin(phi))*
-			(mu_r/(vec[0] - 1.));
-	J12 = -2.*M_PI*r/sqrt(q)*cos_theta*sin(phi) -
-		2.*M_PI*r*sqrt(q)*sin_theta*cos(phi);
-	J21 = J12;
-	J22 = -2.*M_PI*r/sqrt(q)*sin_theta*sin(phi) -
-		2.*M_PI*r*sqrt(q)*cos_theta*cos(phi);
-
-	vector<double> test1 = position;
-	vector<double> test2 = position;
-	test1[0] += 1E-7;
-	test2[0] -= 1E-7;
-	position_to_uniform(test1);
-	position_to_uniform(test2);
-	cout<<(test2[0] - test1[0])/2E-7<<' '<<1./J11<<endl;
-
-	return 1./abs(J11*J22 - J12*J21);
+	return -log(abs(mu_r*log(1. - vec[0])/(1. - vec[0])));
 }
 
 void Elliptical2D::position_from_uniform(vector<double>& vec) const

### 2014-06-08 Documents/Paper/ms.tex
paper

--- 2013-04-28 Documents/ms.tex
@@
 
-\title{Document Title}
+\title{RJObject Paper}
 \author{Brendon J. Brewer}
@@
 \maketitle
+\abstract{This is the abstract.}
 
 \section{Introduction}
+Many inference problems have the following structure.
 
-\section{Spatial Distribution}
-Suppose the two uniform variables are $u_1$ and $u_2$.
-These get turned into an elliptical thing by the following
-one-to-one transformation:
-\begin{eqnarray}
-x &=& x_c + \frac{\mathcal{R}(u_1)}{\sqrt{q}}\cos\left(2\pi u_2\right)\\
-y &=& y_c + \mathcal{R}(u_1)\sqrt{q}\sin\left(2\pi u_2\right)
-\end{eqnarray}
-where $\mathcal{R}(u)$ is some function that determines the
-radial profile. The probability density for $x$ and $y$ is given by:
-\begin{eqnarray}
-p(x, y)
-&=& \frac{p(u_1, u_2)}{\left|\det \mathbf{J}\right|}
-\end{eqnarray}
-where $\mathbf{J}$ is the Jacobian of the transformation:
-\begin{eqnarray}
-\mathbf{J} &=&
-\left(
-\begin{array}{cc}
-\frac{\partial x}{\partial u_1} & \frac{\partial x}{\partial u_2}\\
-\frac{\partial y}{\partial u_1} & \frac{\partial y}{\partial u_2}
-\end{array}
-\right)
-\end{eqnarray}
-Let's calculate the terms. Here's the first one:
-\begin{eqnarray}
-\frac{\partial x}{\partial u_1} &=& \frac{\mathcal{R}'(u_1)}{\sqrt{q}}\cos\left(2\pi u_2\right)
-\end{eqnarray}
-and the second:
-\begin{eqnarray}
-\frac{\partial x}{\partial u_2} &=& \frac{\mathcal{R}(u_1)}{\sqrt{q}}\cos\left(2\pi u_2\right)
-\end{eqnarray}
+\section{Sinusoidal Example}
+Phase transition
+
+\section{``Transit'' Example}
 

### 2014-06-08 Documents/Paper/ms.tex
Starting to describe the problem
Hierarchical

@@
 \section{Introduction}
-Many inference problems have the following structure.
+Many inference problems have the following structure. There is some unknown
+number, $N$, of objects in a region. Each of the $N$ objects has a property
+$x$, which may be a single scalar value (for example, a mass), or a set of
+values (e.g. a position and a mass).
+
+We would like to assign a prior to the objects' properties $\{x_i\}_{i=1}^N$.
+Since $N$ may be large, it is usually easier to assign an ``interim prior''
+conditional on some hyperparameters $\alpha$.
 

### 2014-06-09 Documents/Paper/ms.tex
hierarchical again

@@
 Since $N$ may be large, it is usually easier to assign an ``interim prior''
-conditional on some hyperparameters $\alpha$.
+conditional on some hyperparameters $\alpha$, and then assign a prior to
+$\alpha$. This kind of model is usually called {\it hierarchical}.
 

### 2014-06-09 Documents/Paper/ms.tex
Factorise the prior
Paragraph on our motivation for using NS

@@
 
+The prior for $N$, $\alpha$, and $\{x_i\}$ is usually factorised
+in the following way:
+
+\begin{eqnarray}
+p(N, \alpha, \{x_i\}) &=& p(N) p(\alpha | N) p(\{x_i\} | \alpha, N) \\
+&=& p(N) p(\alpha) \prod_{i=1}^N p(\{x_i\} | \alpha).
+\end{eqnarray}
+
+Here we have assumed the priors for $N$ and $\alpha$ are independent, and
+the interim prior for $\{x_i\}$ is iid and does not depend on $N$.
+
+In other studies, it has been common to do separate runs of Nested Sampling
+with different values of $N$. Then, the posterior for $N$ can be calculated
+based on the estimates of the evidence or marginal likelihood. Our motivation
+for using Nested Sampling is different. With reversible jump MCMC it is possible
+to obtain the posterior for $N$ with a single run. However, the exploration
+of the posterior may be very difficult, and sometimes the problem may even
+contain a phase transition. We use Nested Sampling to overcome these
+difficulties.
+
 \section{Sinusoidal Example}

### 2014-06-09 Documents/Paper/ms.tex
Acknowledgements
DNest paper in bibliography
Some acknowledgements

@@
 
+\section*{Acknowledgements}
+This work is supported by a Marsden Fast-Start grant
+from the Royal Society of New Zealand. I would like to thank
+Anna Pancoast (UCSB), David Hogg (NYU), Daniel Foreman-Mackey (NYU),
+Courtney Donovan (Auckland), Tom Loredo (), John Skilling ()
+
+\begin{thebibliography}{}
+\bibitem[\protect\citeauthoryear{Brewer, P{\'a}rtay,
+\& Cs{\'a}nyi}{2011b}]{dnest} Brewer B.~J., P{\'a}rtay L.~B., Cs{\'a}nyi G., 2011,
+Statistics and Computing, 21, 4, 649-656. arXiv:0912.2380
+
+\end{thebibliography}
+
 \end{document}

### 2014-06-09 Documents/Paper/ms.tex
More acknowledgements

@@
 This work is supported by a Marsden Fast-Start grant
-from the Royal Society of New Zealand. I would like to thank
+from the Royal Society of New Zealand. I would like to thank the following
+people for valuable conversations and inspiration:
 Anna Pancoast (UCSB), David Hogg (NYU), Daniel Foreman-Mackey (NYU),
-Courtney Donovan (Auckland), Tom Loredo (), John Skilling ()
+Courtney Donovan (Auckland), Tom Loredo (),
+John Skilling (MaxEnt Data Consultants), and Daniela Huppenkothen ().
 

### 2014-06-09 Documents/Paper/ms.tex
Note to self
Sinusoid
More references I'll need eventually

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Consider having a section on what kind of ME constraint would lead  %
+% to the same prior as a hierarchical model.                          %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Sinusoidal Example}
 Phase transition
+
+The model for the signal is
+\begin{eqnarray}
+y(t) &=& \sum_{i=1}^N A_i \sin \left(\frac{2\pi t}{T_i} + \phi_i\right)
+\end{eqnarray}
+where there are $N$ sinusoids in the signal, the
+amplitudes are $\{A_i\}$, the periods are $\{T_i\}$, and the phases are
+$\{\phi_i\}$.
 
@@
 
+\bibitem[Brewer et al.(2011)]{2011MNRAS.412.2521B} Brewer, B.~J., Lewis,
+G.~F., Belokurov, V., Irwin, M.~J., Bridges, T.~J., Evans, N.~W.\ 2011.\
+Modelling of the complex CASSOWARY/SLUGS gravitational lenses.\ Monthly
+Notices of the Royal Astronomical Society 412, 2521-2529
+
+\bibitem[\protect\citeauthoryear{Green}{1995}]{rjmcmc}
+Green, P.~J., 1995, Reversible Jump Markov Chain Monte Carlo Computation and Bayesian Model Determination, Biometrika 82 (4): 711–732.
+
+\bibitem[Neal(2001)]{neal} Neal, R.~M., 2001, 
+Annealed importance sampling, Statistics and Computing, vol. 11, pp. 125-139.
+
+\bibitem[\protect\citeauthoryear{Skilling}{1998}]{massinf}
+Skilling J., 1998, Massive Inference and Maximum Entropy, in Maximum Entropy
+and Bayesian Methods, Kluwer Academic Publishers, Dordrecht/Boston/London p.14
+
+\bibitem[\protect\citeauthoryear{Skilling}{2006}]{skilling} Skilling, J., 2006, Nested Sampling for General Bayesian Computation, Bayesian Analysis 4, pp. 833-860.
+
 \end{thebibliography}

### 2014-06-10 Examples/SineWaves/generate_data.py
Generate sinewave data with phase transition

@@ -1,0 +1,41 @@
+from pylab import *
+
+"""
+Original example with two phase transations, but they do not interfere
+with the posterior.
+"""
+seed(123)
+
+t = linspace(0., 100., 1001)
+y = sin(2.*pi*t/30.) + sin(2.*pi*t/2. + 1.)
+sig = 1.
+Y = y + sig*randn(y.size)
+
+data = empty((t.size, 3))
+data[:,0], data[:,1], data[:,2] = t, Y, sig
+
+savetxt('fake_data.txt', data)
+
+plot(t, y, 'r')
+plot(t, Y, 'bo')
+show()
+
+
+"""
+This one has a phase transition at temperature = 1 :-D
+"""
+
+t = linspace(0., 100., 1001)
+y = sin(2.*pi*t/30.) + 0.3*sin(2.*pi*t/2. + 1.)
+sig = 1.
+Y = y + sig*randn(y.size)
+
+data = empty((t.size, 3))
+data[:,0], data[:,1], data[:,2] = t, Y, sig
+
+savetxt('fake_data.txt', data)
+
+plot(t, y, 'r')
+plot(t, Y, 'bo')
+show()
+

### 2014-06-10 Examples/SineWaves/generate_data.py
Beautify plot

@@ -35,7 +35,11 @@ data[:,0], data[:,1], data[:,2] = t, Y, sig
 
 savetxt('fake_data.txt', data)
 
-plot(t, y, 'r')
-plot(t, Y, 'bo')
+plot(t, y, 'r', linewidth=2, label='Noise-free signal')
+plot(t, Y, 'bo', alpha=0.4, label='Noisy measurements')
+ylim([-5., 5.])
+legend(numpoints=1)
+xlabel('Time')
+ylabel('Signal')
 show()
 

### 2014-06-10 Examples/SineWaves/generate_data.py
Save figure

@@ -41,5 +41,6 @@ ylim([-5., 5.])
 legend(numpoints=1)
 xlabel('Time')
 ylabel('Signal')
+savefig('sinewave_data.pdf', bbox_inches='tight')
 show()
 

### 2014-06-10 Examples/SineWaves/generate_data.py
Don't make first plot

@@ -2,7 +2,8 @@ from pylab import *
 
 """
 Original example with two phase transations, but they do not interfere
-with the posterior.
+with the posterior. This is here for no reason other than to make sure
+the second example is identical to what Courtney has been using.
 """
 seed(123)
 
@@ -10,16 +11,6 @@ t = linspace(0., 100., 1001)
 y = sin(2.*pi*t/30.) + sin(2.*pi*t/2. + 1.)
 sig = 1.
 Y = y + sig*randn(y.size)
-
-data = empty((t.size, 3))
-data[:,0], data[:,1], data[:,2] = t, Y, sig
-
-savetxt('fake_data.txt', data)
-
-plot(t, y, 'r')
-plot(t, Y, 'bo')
-show()
-
 
 """
 This one has a phase transition at temperature = 1 :-D

### 2014-06-10 Examples/SineWaves/generate_data.py
Font stuff

@@ -26,6 +26,8 @@ data[:,0], data[:,1], data[:,2] = t, Y, sig
 
 savetxt('fake_data.txt', data)
 
+rc("font", size=14, family="serif", serif="Computer Sans")
+rc("text", usetex=True)
 plot(t, y, 'r', linewidth=2, label='Noise-free signal')
 plot(t, Y, 'bo', alpha=0.4, label='Noisy measurements')
 ylim([-5., 5.])

### 2014-06-10 Documents/Paper/ms.tex
Sampling distribution
Sinewave data figure

@@
 
+The sampling distribution (probability distribution for the data given the
+parameters) is a normal (gaussian) distribution with mean zero and standard
+deviation $\sigma$, applied independently to each data point.
+
+The simulated data is shown in Figure~\ref{fig:sinewave_data}.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{sinewave_data.pdf}
+\caption{The simulated data for the sinusoidal example.
+\label{fig:sinewave_data}}
+\end{center}
+\end{figure}
+
 \section{``Transit'' Example}

### 2014-06-11 Documents/Paper/ms.tex
Paragraphs how I like them
Parskip package
Gaussian likelihood

--- 2014-06-09 Documents/Paper/ms.tex
@@
 \usepackage[left=3cm,top=3cm,right=3cm]{geometry}
+\usepackage{parskip}
 
@@
 \renewcommand{\textfraction}{0.1}
-\parindent=0cm
 
@@
 \section{Introduction}
+\setlength{\parindent}{0cm}
+\setlength{\parskip}{3mm}
 Many inference problems have the following structure. There is some unknown
@@
 \begin{eqnarray}
-y(t) &=& \sum_{i=1}^N A_i \sin \left(\frac{2\pi t}{T_i} + \phi_i\right)
+\mu(t) &=& \sum_{i=1}^N A_i \sin \left(\frac{2\pi t}{T_i} + \phi_i\right)
 \end{eqnarray}
@@
 $\{\phi_i\}$.
+
+The sampling distribution (probability distribution for the data given the
+parameters) is a normal (gaussian) distribution with mean zero and standard
+deviation $\sigma$, applied independently to each data point:
+\begin{eqnarray}
+y_i | N, \{A_i\}, \{T_i\}, \{\phi_i\} \sim
+\mathcal{N}\left(\mu(t_i), \sigma^2\right).
+\end{eqnarray}
+The simulated data is shown in Figure~\ref{fig:sinewave_data}.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{sinewave_data.pdf}
+\caption{The simulated data for the sinusoidal example.
+\label{fig:sinewave_data}}
+\end{center}
+\end{figure}
 

### 2014-06-11 Documents/Paper/ms.tex
Talking about the oscillations

@@
 \end{eqnarray}
-The simulated data is shown in Figure~\ref{fig:sinewave_data}.
+We simulated some data based on the assumption $N=2$.
+The simulated data is shown in Figure~\ref{fig:sinewave_data}, and shows a
+large, low period oscillation with a smaller, much faster oscillation
+superimposed. The noise level is such that the fast oscillation is difficult
+to detect.
 
@@
 
-\section{``Transit'' Example}
+%\section{``Transit'' Example}
 

### 2014-06-11 Examples/SineWaves/Data.cpp
Copied DNest3 template

@@ -1,0 +1,12 @@
+#include "Data.h"
+
+Data::Data()
+{
+
+}
+
+void Data::load(const char* filename)
+{
+
+}
+

### 2014-06-11 Examples/SineWaves/Data.h

@@ -1,0 +1,20 @@
+#ifndef _Data_
+#define _Data_
+
+class Data
+{
+	private:
+
+	public:
+		Data();
+		void load(const char* filename);
+
+	// Singleton
+	private:
+		static Data instance;
+	public:
+		static Data& get_instance() { return instance; }
+};
+
+#endif
+

### 2014-06-11 Examples/SineWaves/MyModel.cpp

@@ -1,0 +1,38 @@
+#include "MyModel.h"
+#include "RandomNumberGenerator.h"
+#include "Utils.h"
+#include <cmath>
+
+using namespace std;
+using namespace DNest3;
+
+MyModel::MyModel()
+{
+
+}
+
+void MyModel::fromPrior()
+{
+
+}
+
+double MyModel::perturb()
+{
+	return 0.;
+}
+
+double MyModel::logLikelihood() const
+{
+	return 0.;
+}
+
+void MyModel::print(std::ostream& out) const
+{
+
+}
+
+string MyModel::description() const
+{
+	return string("");
+}
+

### 2014-06-11 Examples/SineWaves/MyModel.h

@@ -1,0 +1,50 @@
+/*
+* Copyright (c) 2009, 2010, 2011, 2012 Brendon J. Brewer.
+*
+* This file is part of DNest3.
+*
+* DNest3 is free software: you can redistribute it and/or modify
+* it under the terms of the GNU General Public License as published by
+* the Free Software Foundation, either version 3 of the License, or
+* (at your option) any later version.
+*
+* DNest3 is distributed in the hope that it will be useful,
+* but WITHOUT ANY WARRANTY; without even the implied warranty of
+* MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the
+* GNU General Public License for more details.
+*
+* You should have received a copy of the GNU General Public License
+* along with DNest3. If not, see <http://www.gnu.org/licenses/>.
+*/
+
+#ifndef _MyModel_
+#define _MyModel_
+
+#include "Model.h"
+#include <vector>
+
+class MyModel:public DNest3::Model
+{
+	private:
+
+	public:
+		MyModel();
+
+		// Generate the point from the prior
+		void fromPrior();
+
+		// Metropolis-Hastings proposals
+		double perturb();
+
+		// Likelihood function
+		double logLikelihood() const;
+
+		// Print to stream
+		void print(std::ostream& out) const;
+
+		// Return string with column information
+		std::string description() const;
+};
+
+#endif
+

### 2014-06-11 Examples/SineWaves/MyDistribution.cpp
A distribution for RJObject

@@ -1,0 +1,59 @@
+#include "MyDistribution.h"
+#include "RandomNumberGenerator.h"
+#include "Utils.h"
+#include <cmath>
+
+using namespace DNest3;
+
+MyDistribution::MyDistribution(double x_min, double x_max,
+					double mu_min, double mu_max)
+:x_min(x_min)
+,x_max(x_max)
+,mu_min(mu_min)
+,mu_max(mu_max)
+{
+
+}
+
+void MyDistribution::fromPrior()
+{
+	mu = exp(log(mu_min) + log(mu_max/mu_min)*randomU());
+}
+
+double MyDistribution::perturb_parameters()
+{
+	double logH = 0.;
+
+	mu = log(mu);
+	mu += log(mu_max/mu_min)*pow(10., 1.5 - 6.*randomU())*randn();
+	mu = mod(mu - log(mu_min), log(mu_max/mu_min)) + log(mu_min);
+	mu = exp(mu);
+
+	return logH;
+}
+
+double MyDistribution::log_pdf(const std::vector<double>& vec) const
+{
+	if(vec[0] < x_min || vec[0] > x_max || vec[1] < 0.)
+		return -1E300;
+
+	return -log(mu) - vec[1]/mu;
+}
+
+void MyDistribution::from_uniform(std::vector<double>& vec) const
+{
+	vec[0] = x_min + (x_max - x_min)*vec[0];
+	vec[1] = -mu*log(1. - vec[1]);
+}
+
+void MyDistribution::to_uniform(std::vector<double>& vec) const
+{
+	vec[0] = (vec[0] - x_min)/(x_max - x_min);
+	vec[1] = 1. - exp(-vec[1]/mu);
+}
+
+void MyDistribution::print(std::ostream& out) const
+{
+	out<<mu<<' ';
+}
+

### 2014-06-11 Examples/SineWaves/MyDistribution.h

@@ -1,0 +1,34 @@
+#ifndef _MyDistribution_
+#define _MyDistribution_
+
+#include <Distributions/Distribution.h>
+
+// Based on ClassicMassInf1D from RJObject
+
+class MyDistribution:public Distribution
+{
+	private:
+		// Limits
+		double x_min, x_max;
+		double mu_min, mu_max;
+
+		// Mean of exponential distribution for masses
+		double mu;
+
+		double perturb_parameters();
+
+	public:
+		MyDistribution(double x_min, double x_max,
+					double mu_min, double mu_max);
+
+		void fromPrior();
+
+		double log_pdf(const std::vector<double>& vec) const;
+		void from_uniform(std::vector<double>& vec) const;
+		void to_uniform(std::vector<double>& vec) const;
+
+		void print(std::ostream& out) const;
+};
+
+#endif
+

### 2014-06-11 Examples/SineWaves/MyDistribution.h
Note to self

@@ -4,7 +4,8 @@
 #include <Distributions/Distribution.h>
 
 // Based on ClassicMassInf1D from RJObject
-
+// Think of "position x" as log-period
+// and mass as amplitude
 class MyDistribution:public Distribution
 {
 	private:
@@ -12,7 +13,7 @@ class MyDistribution:public Distribution
 		double x_min, x_max;
 		double mu_min, mu_max;
 
-		// Mean of exponential distribution for masses
+		// Mean of exponential distribution for amplitudes
 		double mu;
 
 		double perturb_parameters();

### 2014-06-11 Examples/SineWaves/MyDistribution.cpp
Added phase

@@ -32,9 +32,14 @@ double MyDistribution::perturb_parameters()
 	return logH;
 }
 
+// vec[0] = "position" (log-period)
+// vec[1] = amplitude
+// vec[2] = phase
+
 double MyDistribution::log_pdf(const std::vector<double>& vec) const
 {
-	if(vec[0] < x_min || vec[0] > x_max || vec[1] < 0.)
+	if(vec[0] < x_min || vec[0] > x_max || vec[1] < 0. ||
+			vec[2] < 0. || vec[2] > 2.*M_PI)
 		return -1E300;
 
 	return -log(mu) - vec[1]/mu;
@@ -44,12 +49,14 @@ void MyDistribution::from_uniform(std::vector<double>& vec) const
 {
 	vec[0] = x_min + (x_max - x_min)*vec[0];
 	vec[1] = -mu*log(1. - vec[1]);
+	vec[2] = 2.*M_PI*vec[2];
 }
 
 void MyDistribution::to_uniform(std::vector<double>& vec) const
 {
 	vec[0] = (vec[0] - x_min)/(x_max - x_min);
 	vec[1] = 1. - exp(-vec[1]/mu);
+	vec[2] = vec[2]/(2.*M_PI);
 }
 
 void MyDistribution::print(std::ostream& out) const

### 2014-06-11 Examples/SineWaves/MyModel.cpp
An RJObject for the frequencies

@@ -7,6 +7,7 @@ using namespace std;
 using namespace DNest3;
 
 MyModel::MyModel()
+:objects(3, 10, false, MyDistribution(-10., 10., 1E-3, 1E3))
 {
 
 }

### 2014-06-11 Examples/SineWaves/MyModel.h

@@ -1,31 +1,15 @@
-/*
-* Copyright (c) 2009, 2010, 2011, 2012 Brendon J. Brewer.
-*
-* This file is part of DNest3.
-*
-* DNest3 is free software: you can redistribute it and/or modify
-* it under the terms of the GNU General Public License as published by
-* the Free Software Foundation, either version 3 of the License, or
-* (at your option) any later version.
-*
-* DNest3 is distributed in the hope that it will be useful,
-* but WITHOUT ANY WARRANTY; without even the implied warranty of
-* MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the
-* GNU General Public License for more details.
-*
-* You should have received a copy of the GNU General Public License
-* along with DNest3. If not, see <http://www.gnu.org/licenses/>.
-*/
-
 #ifndef _MyModel_
 #define _MyModel_
 
 #include "Model.h"
 #include <vector>
+#include <RJObject.h>
+#include "MyDistribution.h"
 
 class MyModel:public DNest3::Model
 {
 	private:
+		RJObject<MyDistribution> objects;
 
 	public:
 		MyModel();

### 2014-06-11 Examples/SineWaves/MyModel.cpp
Proposals etc

@@ -14,12 +14,25 @@ MyModel::MyModel()
 
 void MyModel::fromPrior()
 {
-
+	objects.fromPrior();
+	sigma = exp(log(1E-3) + log(1E6)*randomU());
 }
 
 double MyModel::perturb()
 {
-	return 0.;
+	double logH = 0.;
+
+	if(randomU() <= 0.75)
+		logH += objects.perturb();
+	else
+	{
+		sigma = log(sigma);
+		sigma += log(1E6)*randh();
+		sigma = mod(sigma - log(1E-3), log(1E6)) + log(1E-3);
+		sigma = exp(sigma);
+	}
+
+	return logH;
 }
 
 double MyModel::logLikelihood() const
@@ -29,11 +42,12 @@ double MyModel::logLikelihood() const
 
 void MyModel::print(std::ostream& out) const
 {
-
+	objects.print(out); out<<' ';
+	out<<sigma<<' ';
 }
 
 string MyModel::description() const
 {
-	return string("");
+	return string("objects, sigma");
 }
 

### 2014-06-11 Examples/SineWaves/MyModel.h

@@ -10,6 +10,7 @@ class MyModel:public DNest3::Model
 {
 	private:
 		RJObject<MyDistribution> objects;
+		double sigma; // Noise standard deviation
 
 	public:
 		MyModel();

### 2014-06-11 Examples/SineWaves/Data.cpp
Load data

@@ -1,4 +1,8 @@
 #include "Data.h"
+#include <fstream>
+#include <iostream>
+
+using namespace std;
 
 Data::Data()
 {
@@ -7,6 +11,24 @@ Data::Data()
 
 void Data::load(const char* filename)
 {
+	fstream fin(filename, ios::in);
+	if(!fin)
+	{
+		cerr<<"# Error. Couldn't open file "<<filename<<"."<<endl;
+		return;
+	}
 
+	// Empty the vectors
+	t.clear();
+	y.clear();
+
+	double temp1, temp2;
+	while(fin>>temp1 && fin>>temp2)
+	{
+		t.push_back(temp1);
+		y.push_back(temp2);
+	}
+	cout<<"# Loaded "<<t.size()<<" data points from file "<<filename<<endl;
+	fin.close();
 }
 

### 2014-06-11 Examples/SineWaves/Data.h

@@ -1,13 +1,20 @@
 #ifndef _Data_
 #define _Data_
 
+#include <vector>
+
 class Data
 {
 	private:
+		std::vector<double> t, y;
 
 	public:
 		Data();
 		void load(const char* filename);
+
+		// Getters
+		const std::vector<double>& get_t() const { return t; }
+		const std::vector<double>& get_y() const { return y; }
 
 	// Singleton
 	private:

### 2014-06-11 Examples/SineWaves/Data.cpp
Load data set into static instance

@@ -4,6 +4,8 @@
 
 using namespace std;
 
+Data Data::instance;
+
 Data::Data()
 {
 

### 2014-06-11 Examples/SineWaves/main.cpp

@@ -1,19 +1,19 @@
 #include <iostream>
 #include "Start.h"
 #include "MyModel.h"
+#include "Data.h"
 
 using namespace std;
 using namespace DNest3;
 
 int main(int argc, char** argv)
 {
-	#ifndef DNest3_No_Boost
-	MTSampler<MyModel> sampler = setup_mt<MyModel>(argc, argv);
-	#else
-	Sampler<MyModel> sampler = setup<MyModel>(argc, argv);
-	#endif
+	// Load the data
+	Data::get_instance().load("fake_data.txt");
 
+	MTSampler<MyModel> sampler = setup_mt<MyModel>(argc, argv);
 	sampler.run();
+
 	return 0;
 }
 

### 2014-06-11 Examples/SineWaves/Data.cpp
Full stop

@@ -30,7 +30,8 @@ void Data::load(const char* filename)
 		t.push_back(temp1);
 		y.push_back(temp2);
 	}
-	cout<<"# Loaded "<<t.size()<<" data points from file "<<filename<<endl;
+	cout<<"# Loaded "<<t.size()<<" data points from file "
+			<<filename<<"."<<endl;
 	fin.close();
 }
 

### 2014-06-11 Examples/SineWaves/MyModel.cpp
Calculate mock data

@@ -1,6 +1,7 @@
 #include "MyModel.h"
 #include "RandomNumberGenerator.h"
 #include "Utils.h"
+#include "Data.h"
 #include <cmath>
 
 using namespace std;
@@ -8,6 +9,7 @@ using namespace DNest3;
 
 MyModel::MyModel()
 :objects(3, 10, false, MyDistribution(-10., 10., 1E-3, 1E3))
+,mu(Data::get_instance().get_t().size())
 {
 
 }
@@ -16,6 +18,29 @@ void MyModel::fromPrior()
 {
 	objects.fromPrior();
 	sigma = exp(log(1E-3) + log(1E6)*randomU());
+	calculate_mu();
+}
+
+void MyModel::calculate_mu()
+{
+	// Get the times from the data
+	const vector<double>& t = Data::get_instance().get_t();
+
+	// Get the components
+	const vector< vector<double> >& components = objects.get_components();
+
+	// Zero the signal
+	mu.assign(mu.size(), 0.);
+
+	double T, A, phi;
+	for(size_t j=0; j<components.size(); j++)
+	{
+		T = exp(components[j][0]);
+		A = components[j][1];
+		phi = components[j][2];
+		for(size_t i=0; i<t.size(); i++)
+			mu[i] += A*sin(2.*M_PI*t[i]/T + phi);
+	}
 }
 
 double MyModel::perturb()
@@ -23,7 +48,10 @@ double MyModel::perturb()
 	double logH = 0.;
 
 	if(randomU() <= 0.75)
+	{
 		logH += objects.perturb();
+		calculate_mu();
+	}
 	else
 	{
 		sigma = log(sigma);
@@ -42,8 +70,10 @@ double MyModel::logLikelihood() const
 
 void MyModel::print(std::ostream& out) const
 {
-	objects.print(out); out<<' ';
+	for(size_t i=0; i<mu.size(); i++)
+		out<<mu[i]<<' ';
 	out<<sigma<<' ';
+	objects.print(out); out<<' ';
 }
 
 string MyModel::description() const

### 2014-06-11 Examples/SineWaves/MyModel.h

@@ -12,6 +12,10 @@ class MyModel:public DNest3::Model
 		RJObject<MyDistribution> objects;
 		double sigma; // Noise standard deviation
 
+		// The signal
+		std::vector<double> mu;
+		void calculate_mu();
+
 	public:
 		MyModel();
 

### 2014-06-11 Examples/SineWaves/MyModel.cpp
Likelihood

@@ -65,7 +65,15 @@ double MyModel::perturb()
 
 double MyModel::logLikelihood() const
 {
-	return 0.;
+	// Get the data
+	const vector<double>& y = Data::get_instance().get_y();
+
+	double logL = 0.;
+	double var = sigma*sigma;
+	for(size_t i=0; i<y.size(); i++)
+		logL += -0.5*log(2.*M_PI*var) - 0.5*pow(y[i] - mu[i], 2)/var;
+
+	return logL;
 }
 
 void MyModel::print(std::ostream& out) const

### 2014-06-11 Examples/SineWaves/display.py
display script

@@ -1,0 +1,14 @@
+from pylab import *
+
+data = loadtxt('fake_data.txt')
+posterior_sample = atleast_2d(loadtxt('posterior_sample.txt'))
+
+ion()
+for i in xrange(0, posterior_sample.shape[0]):
+  hold(False)
+  plot(data[:,0], data[:,1], 'b.')
+  hold(True)
+  plot(data[:,0], posterior_sample[i, 0:data.shape[0]], 'r')
+  draw()
+ioff()
+show()

### 2014-06-11 Examples/SineWaves/showresults.py

@@ -1,26 +1,4 @@
-# Copyright (c) 2009, 2010, 2011, 2012 Brendon J. Brewer.
-#
-# This file is part of DNest3.
-#
-# DNest3 is free software: you can redistribute it and/or modify
-# it under the terms of the GNU General Public License as published by
-# the Free Software Foundation, either version 3 of the License, or
-# (at your option) any later version.
-#
-# DNest3 is distributed in the hope that it will be useful,
-# but WITHOUT ANY WARRANTY; without even the implied warranty of
-# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the
-# GNU General Public License for more details.
-#
-# You should have received a copy of the GNU General Public License
-# along with DNest3. If not, see <http://www.gnu.org/licenses/>.
-
-# Import postprocess from two directories up
-# http://stackoverflow.com/a/9806045
-import os
-parentdir = os.path.dirname(os.path.dirname(os.path.abspath(__file__)))
-parentdir = os.path.dirname(parentdir)
-os.sys.path.insert(0, parentdir) 
 import postprocess
 postprocess.postprocess()
+import display
 

### 2014-06-11 Examples/SineWaves/generate_data.py
Two columns

@@ -21,8 +21,8 @@ y = sin(2.*pi*t/30.) + 0.3*sin(2.*pi*t/2. + 1.)
 sig = 1.
 Y = y + sig*randn(y.size)
 
-data = empty((t.size, 3))
-data[:,0], data[:,1], data[:,2] = t, Y, sig
+data = empty((t.size, 2))
+data[:,0], data[:,1] = t, Y
 
 savetxt('fake_data.txt', data)
 

### 2014-06-14 Documents/Paper/ms.tex
Starfield reference
Words about sinusoid

@@
 \section{Sinusoidal Example}
-Phase transition
+In this section we demonstrate a seemingly simple example which exhibits
+a phase transition, making a standard MCMC approach difficult. Suppose
+a signal is composed of $N$ sinusoids, of different frequencies and amplitudes.
+This kind of model has many applications\citep[see e.g.][]{bretthorst}.
 
@@
 \begin{thebibliography}{}
+
+\bibitem[Brewer et al.(2013)]{2013AJ....146....7B} Brewer, B.~J., 
+Foreman-Mackey, D., Hogg, D.~W.\ 2013.\ Probabilistic Catalogs for Crowded 
+Stellar Fields.\ The Astronomical Journal 146, 7.
+
 \bibitem[\protect\citeauthoryear{Brewer, P{\'a}rtay,

### 2014-06-14 Documents/Paper/ms.tex
More talk about sinusoids

@@
 a phase transition, making a standard MCMC approach difficult. Suppose
-a signal is composed of $N$ sinusoids, of different frequencies and amplitudes.
+a signal is composed of $N$ sinusoids, of different periods, amplitudes,
+and phases. The signal is observed at various times $\{t_i\}_{i=1}^N$ with
+noise, and we want to use the resulting data to infer the number of sinusoids
+$N$, along with the periods $\{T_i\}_{i=1}^N$, amplitudes $\{A_i\}_{i=1}^N$,
+and phases $\{\phi_i\}_{i=1}^N$.
 This kind of model has many applications\citep[see e.g.][]{bretthorst}.

### 2014-06-15 Documents/Paper/ms.tex
More writing about sinusoid data
Talking about the proposals

@@
 \usepackage{parskip}
+\usepackage{amsmath}
 
@@
 
-\section{Sinusoidal Example}
+\section{Metropolis proposals for the general problem}
+We will now define a set of proposal distributions that could be used to
+sample the prior distribution
+$p(N, \alpha, \{x_i\}) = p(N) p(\alpha) \prod_{i=1}^N p(\{x_i\} | \alpha)$.
+These proposals will need to satisfy detailed balance in order to be valid
+when used inside DNS.
+
+\section{Sinusoidal example}
 In this section we demonstrate a seemingly simple example which exhibits
@@
 
+The true values of the parameters are $N=2$, $\mathbf{A} = \{1, 0.3\}$,
+$\mathbf{T}=\{30, 2\}$, and $\boldsymbol{\phi} = \{0, 1\}$. The signal was
+observed at $n=1001$ points equally spaced between $t=0$ and $t=100$. The data
+is displayed in Figure~\ref{fig:sinewave_data}.
+
 \begin{figure}
@@
 
-%\section{``Transit'' Example}
+%\section{``Transit'' example}
 
@@
 Anna Pancoast (UCSB), David Hogg (NYU), Daniel Foreman-Mackey (NYU),
-Courtney Donovan (Auckland), Tom Loredo (),
+Courtney Donovan (Auckland), Tom Loredo (), Iain Murray (Edinburgh),
 John Skilling (MaxEnt Data Consultants), and Daniela Huppenkothen ().

### 2014-06-15 Documents/Paper/ms.tex
Need heavy tailed proposals

@@
 These proposals will need to satisfy detailed balance in order to be valid
-when used inside DNS.
+when used inside DNS. The proposals will also be very heavy tailed. When
+DNS is exploring a distribution close to the prior, large proposals will
+generally be needed. When DNS is exploring a distribution constrained to
+very high values of the likelihood function, much smaller proposals will be
+appropriate. Rather than attempting expensive tuning (since the number of
+distributions involved may number in the hundreds or even thousands), we will
+apply the heavy tailed proposal distributions and simply accept that there will
+be some waste.
 

### 2014-06-19 Documents/Paper/ms.tex
Other ideas for examples

--- 2014-06-14 Documents/Paper/ms.tex
@@
 \section{Sinusoidal Example}
-In this section we demonstrate a seemingly simple example which exhibits
-a phase transition, making a standard MCMC approach difficult. Suppose
-a signal is composed of $N$ sinusoids, of different periods, amplitudes,
-and phases. The signal is observed at various times $\{t_i\}_{i=1}^N$ with
-noise, and we want to use the resulting data to infer the number of sinusoids
-$N$, along with the periods $\{T_i\}_{i=1}^N$, amplitudes $\{A_i\}_{i=1}^N$,
-and phases $\{\phi_i\}_{i=1}^N$.
-This kind of model has many applications\citep[see e.g.][]{bretthorst}.
+Phase transition
 
@@
 %\section{``Transit'' Example}
+%\section{``Asteroseismology'' Example}
+%\section{``Galaxy Field'' Example}
 
@@
 \begin{thebibliography}{}
-
-\bibitem[Brewer et al.(2013)]{2013AJ....146....7B} Brewer, B.~J., 
-Foreman-Mackey, D., Hogg, D.~W.\ 2013.\ Probabilistic Catalogs for Crowded 
-Stellar Fields.\ The Astronomical Journal 146, 7.
-
 \bibitem[\protect\citeauthoryear{Brewer, P{\'a}rtay,

### 2014-06-19 Documents/Paper/ms.tex
Talking about the simulated sinewave data

--- 2014-06-15 Documents/Paper/ms.tex
@@
 \begin{eqnarray}
-\mu(t) &=& \sum_{i=1}^N A_i \sin \left(\frac{2\pi t}{T_i} + \phi_i\right)
+y(t) &=& \sum_{i=1}^N A_i \sin \left(\frac{2\pi t}{T_i} + \phi_i\right)
 \end{eqnarray}
@@
 \begin{eqnarray}
-y_i | N, \{A_i\}, \{T_i\}, \{\phi_i\} \sim
+Y_i | N, \{A_i\}, \{T_i\}, \{\phi_i\} \sim
 \mathcal{N}\left(\mu(t_i), \sigma^2\right).
@@
 superimposed. The noise level is such that the fast oscillation is difficult
-to detect.
-
-The true values of the parameters are $N=2$, $\mathbf{A} = \{1, 0.3\}$,
+to detect. The true values of the parameters were
+$N=2$, $\mathbf{A} = \{1, 0.3\}$,
 $\mathbf{T}=\{30, 2\}$, and $\boldsymbol{\phi} = \{0, 1\}$. The signal was
-observed at $n=1001$ points equally spaced between $t=0$ and $t=100$. The data
+observed at $n=1001$ points equally spaced between $t=0$ and $t=100$.
+The true form of the signal is:
+\begin{eqnarray}
+y(t) = \sin\left(\frac{2\pi t}{30}\right) +
+0.3 \sin\left(\frac{2\pi t}{2} + 1\right)
+\end{eqnarray}
+and the noise standard deviation was $\sigma = 1$. The data $\{Y_i\}_{i=1}^n$
 is displayed in Figure~\ref{fig:sinewave_data}.
@@
 \includegraphics[scale=0.5]{sinewave_data.pdf}
-\caption{The simulated data for the sinusoidal example.
+\caption{The simulated data for the sinusoidal example. The solid line shows
+the true signal and the points are the measurements, simulated based on a
+noise standard deviation of $\sigma = 1$.
 \label{fig:sinewave_data}}
@@
 
+%\section{``Transit'' Example}
+%\section{``Asteroseismology'' Example}
+%\section{``Galaxy Field'' Example}
 %\section{``Transit'' example}

### 2014-06-21 Documents/Paper/ms.tex
Writing about proposals

@@
 be some waste.
+
+Most of the proposals involve changing one (or a subset) of the parameters
+and/or hyperparameters while keeping the others fixed.
+
+\subsection{Proposals that modify $N$}
+The first kind of proposal we consider are proposals that change the
+dimension of the model, i.e. proposals that change the value of $N$. By
+necessity, these will also change the parameters $\{x_i\}_{i=1}^N$ because
+the number of parameters will be changed.
 

### 2014-06-21 Documents/Paper/ms.tex
Proposal for N

@@
 necessity, these will also change the parameters $\{x_i\}_{i=1}^N$ because
-the number of parameters will be changed.
+the number of parameters will be changed. The proposals used here are
+traditionally called {\it birth and death} proposals.
+
+We will assume that the prior for $N$ is a discrete uniform distribution
+between 0 and some $N_{\rm max}$, inclusive. The proposal starts by choosing
+a new value $N'$ according to
+\begin{eqnarray}
+N' &=& \textnormal{mod}\left(N + \delta_N, N_{\rm max} + 1\right)
+\end{eqnarray}
+where $\delta_N$ is drawn from a heavy tailed
+distribution.
 

### 2014-06-21 Documents/Paper/ms.tex
What happens to parameters when N changes
Slight wording chaneg
The extent of the heavy tailed distribution

@@
 p(N, \alpha, \{x_i\}) &=& p(N) p(\alpha | N) p(\{x_i\} | \alpha, N) \\
-&=& p(N) p(\alpha) \prod_{i=1}^N p(\{x_i\} | \alpha).
+&=& p(N) p(\alpha) \prod_{i=1}^N p(x_i | \alpha).
 \end{eqnarray}
@@
 where $\delta_N$ is drawn from a heavy tailed
-distribution.
+distribution. The most probable values of $\delta_N$ are $\pm 1$, but values
+of order $N_{\rm max}$ also have some probability, to allow fast exploration
+when the target distribution is similar to the prior.
+
+If $N' > N$, the extra parameters needed, $\{x_i\}_{i=N+1}^{N'}$,
+are drawn from their prior conditional on the current value of the
+hyperparameters, i.e. the interim prior $p(x | \alpha)$.
+
+If $N' < N$, then $N - N'$ objects must be removed from the model. All
+of the objects have the same probability of being selected for removal.
 

### 2014-06-21 Documents/Paper/ms.tex
Subsections for the different proposals
Removed braces

@@
 
+% Consider terminology: "objects" vs. "components"
+
 \section{Metropolis proposals for the general problem}
@@
 sample the prior distribution
-$p(N, \alpha, \{x_i\}) = p(N) p(\alpha) \prod_{i=1}^N p(\{x_i\} | \alpha)$.
+$p(N, \alpha, \{x_i\}) = p(N) p(\alpha) \prod_{i=1}^N p(x_i | \alpha)$.
 These proposals will need to satisfy detailed balance in order to be valid
@@
 of the objects have the same probability of being selected for removal.
+
+\subsection{Proposals that modify objects}
+
+\subsection{Proposals that change the hyperparameters, keeping the objects fixed}
+
+\subsection{Proposals that change the hyperparameters and all of the objects}
 

### 2014-06-22 Documents/Paper/ms.tex
CDF etc

@@
 \subsection{Proposals that modify objects}
+Let $F(x | \alpha)$ be the cumulative distribution corresponding to the interim
+prior $p(x | \alpha)$, and $F^{-1}(x | \alpha)$ its inverse.
+Then $F(x_i)$ (for all $i$) has a uniform prior between 0 and 1. A proposal
+for an object involves transforming its parameters to $[0, 1]$ using $F^{-1}$,
+make the proposal, and then transform back.
 

### 2014-06-23 Documents/Paper/ms.tex
Equations
Not necessarily the CDF

--- 2014-06-21 Documents/Paper/ms.tex
@@
 \subsection{Proposals that modify objects}
+We now consider a proposal distribution that modifies one or more of the
+objects $\{x_i\}$, while keeping the number of objects $N$, as well as the
+hyperparameters $\alpha$, fixed.
+
+Let $F(x; \alpha)$ be a function that takes an object $x$ and transforms it
+to a value $u$ that has a uniform distribution between 0 and 1, given $\alpha$.
+If the object $x$ consists of a single scalar value, $F$ is the cumulative
+distribution of the interim prior. Denote the inverse of $F$ by $G$.
+
+A proposal
+for an object involves transforming its parameters to $[0, 1]$ using $F$,
+making the proposal in terms of $u$, and then transforming back.
+Specifically, the proposal chooses
+a new value $x_i'$ from the current value $x_i$ as follows:
+\begin{eqnarray}
+u_i &=& F(x_i; \alpha)\\
+u_i' &=& \textnormal{mod}\left(u_i + \delta_u, 1\right)\\
+x_i' &=& F^{-1}(u_i'; \alpha).
+\end{eqnarray}
 

### 2014-06-23 Documents/Paper/ms.tex
Notation
Number to change uses heavy tailed dist

@@
 u_i' &=& \textnormal{mod}\left(u_i + \delta_u, 1\right)\\
-x_i' &=& F^{-1}(u_i'; \alpha).
+x_i' &=& G(u_i'; \alpha).
 \end{eqnarray}
+where $\delta_u$ is drawn from a heavy-tailed distribution.
+
+Choosing just one object to change is most appropriate when the DNS
+distribution is very constrained. When it is close to the prior, bolder
+proposals that change more than one object at a time are possible. Hence,
+we choose the number of objects to change from a heavy tailed distribution
+wherer the most probable value is 1 but there is also a nontrivial probability
+of proposing to change $\sim N$ objects at once.
 

### 2014-06-23 Documents/Paper/ms.tex
Proposal to move alpha

@@
 \subsection{Proposals that change the hyperparameters, keeping the objects fixed}
+Another kind of proposal that we will consider is a proposal that keeps all of
+the objects fixed in place (i.e. leaves $\{x_i\}$) unchanged, but changes
+the hyperparameter(s) from their current value $\alpha$
+to a new value $\alpha'$. The proposal for the hyperparameter(s) is chosen
+to satisfy detailed balance with respect to $p(\alpha)$.
+
+The Metropolis acceptance probability must include the following factor:
+\begin{eqnarray}
+\frac{\prod_{i=1}^N p(x_i | \alpha')}{\prod_{i=1}^N p(x_i | \alpha)}
+\end{eqnarray}
 
@@
 %\section{``Galaxy Field'' Example}
-%\section{``Transit'' example}
 

### 2014-06-23 Documents/Paper/ms.tex
Efficiency
Inefficiency

@@
 
+Since this proposal leaves the objects $\{x_i\}$ fixed, it will usually not
+affect the value of the likelihood, and therefore the likelihood will not need
+to be recomputed.
+
 \subsection{Proposals that change the hyperparameters and all of the objects}
+The above proposals allow for changes to $N$, $\alpha$, and $\{x_i\}$,
+and are therefore sufficient to allow for ``correct'' exploration of the
+prior distribution, and the constrained prior distributions used in DNS.
+However, they do not necessarily allow for {\it efficient} exploration, even
+of the prior itself. The main reason for this is the inability for large
+changes to be made to the hyperparameters $\alpha$.
 

### 2014-06-23 Documents/Paper/ms.tex
Wording change

@@
 
-The Metropolis acceptance probability must include the following factor:
+The overall Metropolis acceptance probability
+for this kind of move, if sampling the prior (or the constrained prior of DNS)
+must include the following factor:
 \begin{eqnarray}
@@
 \end{eqnarray}
-
 Since this proposal leaves the objects $\{x_i\}$ fixed, it will usually not

### 2014-06-23 Documents/Paper/ms.tex
Labels and more talking

@@
 
-\subsection{Proposals that modify $N$}
+\subsection{Proposals that modify $N$}\label{sec:proposal1}
 The first kind of proposal we consider are proposals that change the
@@
 
-\subsection{Proposals that modify objects}
+\subsection{Proposals that modify objects}\label{sec:proposal2}
 We now consider a proposal distribution that modifies one or more of the
@@
 
-\subsection{Proposals that change the hyperparameters, keeping the objects fixed}
+\subsection{Proposals that change the hyperparameters,
+keeping the objects fixed}\label{sec:proposal3}
 Another kind of proposal that we will consider is a proposal that keeps all of
@@
 \frac{\prod_{i=1}^N p(x_i | \alpha')}{\prod_{i=1}^N p(x_i | \alpha)}
+\label{eqn:acceptance_prob}
 \end{eqnarray}
@@
 
-\subsection{Proposals that change the hyperparameters and all of the objects}
+\subsection{Proposals that change the hyperparameters
+and all of the objects}\label{sec:proposal4}
 The above proposals allow for changes to $N$, $\alpha$, and $\{x_i\}$,
@@
 of the prior itself. The main reason for this is the inability for large
-changes to be made to the hyperparameters $\alpha$.
+changes to be made to the hyperparameters $\alpha$. If the proposed change
+from $\alpha$ to $\alpha'$ is large, the ratio in
+Equation~\ref{eqn:acceptance_prob} is likely to be very small, and the move
+will probably be rejected.
+
+Therefore, we allow an additional move that changes $\alpha$ to a new
+value $\alpha'$, but rather than leaving the objects $\{x_i\}$ fixed, we
+drag them so they represent the distribution $p(x | \alpha')$ rather than
+$p(x|\alpha)$. We do this by making use of the transformation functions
+$F(x; \alpha)$ and $G(x; \alpha)$ defined in Section~\ref{sec:proposal2}.
 

### 2014-06-23 Documents/Paper/ms.tex
Dragging equations
:= for assignment
Updated affiliations

@@
 \begin{eqnarray}
-u_i &=& F(x_i; \alpha)\\
-u_i' &=& \textnormal{mod}\left(u_i + \delta_u, 1\right)\\
-x_i' &=& G(u_i'; \alpha).
+u_i &:=& F(x_i; \alpha)\\
+u_i' &:=& \textnormal{mod}\left(u_i + \delta_u, 1\right)\\
+x_i' &:=& G(u_i'; \alpha).
 \end{eqnarray}
@@
 $F(x; \alpha)$ and $G(x; \alpha)$ defined in Section~\ref{sec:proposal2}.
+
+The ``dragging'' process works as follows, and must be carried out on
+each object:
+\begin{eqnarray}
+u_i &:=& F(x_i; \alpha)\\
+x_i' &:=& G(x_i; \alpha')
+\end{eqnarray}
 
@@
 Anna Pancoast (UCSB), David Hogg (NYU), Daniel Foreman-Mackey (NYU),
-Courtney Donovan (Auckland), Tom Loredo (), Iain Murray (Edinburgh),
-John Skilling (MaxEnt Data Consultants), and Daniela Huppenkothen ().
+Courtney Donovan (Auckland), Tom Loredo (Cornell), Iain Murray (Edinburgh),
+John Skilling (MaxEnt Data Consultants), and Daniela Huppenkothen
+(Amsterdam, NYU).
 

### 2014-06-23 Documents/Paper/pgm.py
A pgm

@@ -1,0 +1,29 @@
+"""
+PGM for the general problem using daft (http://daft-pgm.org/)
+"""
+
+from matplotlib import rc
+rc("font", family="serif", size=12)
+rc("text", usetex=True)
+
+import daft
+
+pgm = daft.PGM([4., 4.], origin=[-1.5, -1.5])
+
+# Create the nodes
+pgm.add_node(daft.Node('alpha', r'$\alpha$', 0., 2.))
+pgm.add_node(daft.Node('x', r'$x_i$', 0., 0.))
+pgm.add_node(daft.Node('data', r'$\mathcal{D}$', 2., 0., observed=True))
+
+# Add the edges
+pgm.add_edge('alpha', 'x')
+pgm.add_edge('x', 'data')
+
+# Add the plates
+pgm.add_plate(daft.Plate([-1., -1., 2., 2.], label=r'Objects $i=1, ..., N$'))
+
+pgm.render()
+pgm.figure.savefig("pgm.eps")
+pgm.figure.savefig("pgm.pdf")
+
+

### 2014-06-23 Documents/Paper/ms.tex
Embolden symbols

@@
 \renewcommand{\textfraction}{0.1}
+
+\newcommand{\hyperparams}{\boldsymbol{\alpha}}
+\newcommand{\xx}{\mathbf{x}}
 
@@
 number, $N$, of objects in a region. Each of the $N$ objects has a property
-$x$, which may be a single scalar value (for example, a mass), or a set of
+$\xx$, which may be a single scalar value (for example, a mass), or a set of
 values (e.g. a position and a mass).
 
-We would like to assign a prior to the objects' properties $\{x_i\}_{i=1}^N$.
+We would like to assign a prior to the objects' properties $\{\xx_i\}_{i=1}^N$.
 Since $N$ may be large, it is usually easier to assign an ``interim prior''
-conditional on some hyperparameters $\alpha$, and then assign a prior to
-$\alpha$. This kind of model is usually called {\it hierarchical}.
+conditional on some hyperparameters $\hyperparams$, and then assign a prior to
+$\hyperparams$. This kind of model is usually called {\it hierarchical}.
 
-The prior for $N$, $\alpha$, and $\{x_i\}$ is usually factorised
+The prior for $N$, $\hyperparams$, and $\{\xx_i\}$ is usually factorised
 in the following way:
@@
 \begin{eqnarray}
-p(N, \alpha, \{x_i\}) &=& p(N) p(\alpha | N) p(\{x_i\} | \alpha, N) \\
-&=& p(N) p(\alpha) \prod_{i=1}^N p(x_i | \alpha).
+p(N, \hyperparams, \{\xx_i\}) &=& p(N) p(\hyperparams | N) p(\{\xx_i\} | \hyperparams, N) \\
+&=& p(N) p(\hyperparams) \prod_{i=1}^N p(\xx_i | \hyperparams).
 \end{eqnarray}
 
-Here we have assumed the priors for $N$ and $\alpha$ are independent, and
-the interim prior for $\{x_i\}$ is iid and does not depend on $N$.
+Here we have assumed the priors for $N$ and $\hyperparams$ are independent, and
+the interim prior for $\{\xx_i\}$ is iid and does not depend on $N$.
 
@@
 sample the prior distribution
-$p(N, \alpha, \{x_i\}) = p(N) p(\alpha) \prod_{i=1}^N p(x_i | \alpha)$.
+$p(N, \hyperparams, \{\xx_i\}) = p(N) p(\hyperparams) \prod_{i=1}^N p(\xx_i | \hyperparams)$.
 These proposals will need to satisfy detailed balance in order to be valid
@@
 dimension of the model, i.e. proposals that change the value of $N$. By
-necessity, these will also change the parameters $\{x_i\}_{i=1}^N$ because
+necessity, these will also change the parameters $\{\xx_i\}_{i=1}^N$ because
 the number of parameters will be changed. The proposals used here are
@@
 
-If $N' > N$, the extra parameters needed, $\{x_i\}_{i=N+1}^{N'}$,
+If $N' > N$, the extra parameters needed, $\{\xx_i\}_{i=N+1}^{N'}$,
 are drawn from their prior conditional on the current value of the
-hyperparameters, i.e. the interim prior $p(x | \alpha)$.
+hyperparameters, i.e. the interim prior $p(x | \hyperparams)$.
 
@@
 We now consider a proposal distribution that modifies one or more of the
-objects $\{x_i\}$, while keeping the number of objects $N$, as well as the
-hyperparameters $\alpha$, fixed.
+objects $\{\xx_i\}$, while keeping the number of objects $N$, as well as the
+hyperparameters $\hyperparams$, fixed.
 
-Let $F(x; \alpha)$ be a function that takes an object $x$ and transforms it
-to a value $u$ that has a uniform distribution between 0 and 1, given $\alpha$.
+Let $F(x; \hyperparams)$ be a function that takes an object $x$ and transforms it
+to a value $u$ that has a uniform distribution between 0 and 1, given $\hyperparams$.
 If the object $x$ consists of a single scalar value, $F$ is the cumulative
@@
 Specifically, the proposal chooses
-a new value $x_i'$ from the current value $x_i$ as follows:
+a new value $\xx_i'$ from the current value $\xx_i$ as follows:
 \begin{eqnarray}
-u_i &:=& F(x_i; \alpha)\\
+u_i &:=& F(\xx_i; \hyperparams)\\
 u_i' &:=& \textnormal{mod}\left(u_i + \delta_u, 1\right)\\
-x_i' &:=& G(u_i'; \alpha).
+\xx_i' &:=& G(u_i'; \hyperparams).
 \end{eqnarray}
@@
 Another kind of proposal that we will consider is a proposal that keeps all of
-the objects fixed in place (i.e. leaves $\{x_i\}$) unchanged, but changes
-the hyperparameter(s) from their current value $\alpha$
-to a new value $\alpha'$. The proposal for the hyperparameter(s) is chosen
-to satisfy detailed balance with respect to $p(\alpha)$.
+the objects fixed in place (i.e. leaves $\{\xx_i\}$) unchanged, but changes
+the hyperparameter(s) from their current value $\hyperparams$
+to a new value $\hyperparams'$. The proposal for the hyperparameter(s) is chosen
+to satisfy detailed balance with respect to $p(\hyperparams)$.
 
@@
 \begin{eqnarray}
-\frac{\prod_{i=1}^N p(x_i | \alpha')}{\prod_{i=1}^N p(x_i | \alpha)}
+\frac{\prod_{i=1}^N p(\xx_i | \hyperparams')}{\prod_{i=1}^N p(\xx_i | \hyperparams)}
 \label{eqn:acceptance_prob}
 \end{eqnarray}
-Since this proposal leaves the objects $\{x_i\}$ fixed, it will usually not
+Since this proposal leaves the objects $\{\xx_i\}$ fixed, it will usually not
 affect the value of the likelihood, and therefore the likelihood will not need
@@
 and all of the objects}\label{sec:proposal4}
-The above proposals allow for changes to $N$, $\alpha$, and $\{x_i\}$,
+The above proposals allow for changes to $N$, $\hyperparams$, and $\{\xx_i\}$,
 and are therefore sufficient to allow for ``correct'' exploration of the
@@
 of the prior itself. The main reason for this is the inability for large
-changes to be made to the hyperparameters $\alpha$. If the proposed change
-from $\alpha$ to $\alpha'$ is large, the ratio in
+changes to be made to the hyperparameters $\hyperparams$. If the proposed change
+from $\hyperparams$ to $\hyperparams'$ is large, the ratio in
 Equation~\ref{eqn:acceptance_prob} is likely to be very small, and the move
@@
 
-Therefore, we allow an additional move that changes $\alpha$ to a new
-value $\alpha'$, but rather than leaving the objects $\{x_i\}$ fixed, we
-drag them so they represent the distribution $p(x | \alpha')$ rather than
-$p(x|\alpha)$. We do this by making use of the transformation functions
-$F(x; \alpha)$ and $G(x; \alpha)$ defined in Section~\ref{sec:proposal2}.
+Therefore, we allow an additional move that changes $\hyperparams$ to a new
+value $\hyperparams'$, but rather than leaving the objects $\{\xx_i\}$ fixed, we
+drag them so they represent the distribution $p(x | \hyperparams')$ rather than
+$p(x|\hyperparams)$. We do this by making use of the transformation functions
+$F(x; \hyperparams)$ and $G(x; \hyperparams)$ defined in Section~\ref{sec:proposal2}.
 
@@
 \begin{eqnarray}
-u_i &:=& F(x_i; \alpha)\\
-x_i' &:=& G(x_i; \alpha')
+u_i &:=& F(\xx_i; \hyperparams)\\
+\xx_i' &:=& G(\xx_i; \hyperparams')
 \end{eqnarray}

### 2014-06-23 Documents/Paper/ms.tex
More bold

@@
 are drawn from their prior conditional on the current value of the
-hyperparameters, i.e. the interim prior $p(x | \hyperparams)$.
+hyperparameters, i.e. the interim prior $p(\xx | \hyperparams)$.
 
@@
 value $\hyperparams'$, but rather than leaving the objects $\{\xx_i\}$ fixed, we
-drag them so they represent the distribution $p(x | \hyperparams')$ rather than
-$p(x|\hyperparams)$. We do this by making use of the transformation functions
+drag them so they represent the distribution $p(\xx | \hyperparams')$ rather than
+$p(\xx|\hyperparams)$. We do this by making use of the transformation functions
 $F(x; \hyperparams)$ and $G(x; \hyperparams)$ defined in Section~\ref{sec:proposal2}.

### 2014-06-23 Documents/Paper/pgm.py
Bold x (not sure how to make the alpha bold)

@@ -12,7 +12,7 @@ pgm = daft.PGM([4., 4.], origin=[-1.5, -1.5])
 
 # Create the nodes
 pgm.add_node(daft.Node('alpha', r'$\alpha$', 0., 2.))
-pgm.add_node(daft.Node('x', r'$x_i$', 0., 0.))
+pgm.add_node(daft.Node('x', r'$\mathbf{x}_i$', 0., 0.))
 pgm.add_node(daft.Node('data', r'$\mathcal{D}$', 2., 0., observed=True))
 
 # Add the edges

### 2014-06-23 Documents/Paper/pgm.py
Added N node

@@ -8,16 +8,19 @@ rc("text", usetex=True)
 
 import daft
 
-pgm = daft.PGM([4., 4.], origin=[-1.5, -1.5])
+pgm = daft.PGM([5., 4.], origin=[-2.5, -1.5])
 
 # Create the nodes
 pgm.add_node(daft.Node('alpha', r'$\alpha$', 0., 2.))
 pgm.add_node(daft.Node('x', r'$\mathbf{x}_i$', 0., 0.))
 pgm.add_node(daft.Node('data', r'$\mathcal{D}$', 2., 0., observed=True))
+pgm.add_node(daft.Node('N', r'$N$', -2., 0.))
+pgm.add_node(daft.Node('hidden', r'', -1., 0., scale=0, plot_params={'alpha':0}))
 
 # Add the edges
 pgm.add_edge('alpha', 'x')
 pgm.add_edge('x', 'data')
+pgm.add_edge('N', 'hidden')
 
 # Add the plates
 pgm.add_plate(daft.Plate([-1., -1., 2., 2.], label=r'Objects $i=1, ..., N$'))

### 2014-06-23 Documents/Paper/pgm.py
Bold alpha

@@ -2,16 +2,18 @@
 PGM for the general problem using daft (http://daft-pgm.org/)
 """
 
+import matplotlib
 from matplotlib import rc
 rc("font", family="serif", size=12)
 rc("text", usetex=True)
+matplotlib.rcParams['text.latex.preamble']=[r"\usepackage{amsmath}"]
 
 import daft
 
 pgm = daft.PGM([5., 4.], origin=[-2.5, -1.5])
 
 # Create the nodes
-pgm.add_node(daft.Node('alpha', r'$\alpha$', 0., 2.))
+pgm.add_node(daft.Node('alpha', r'$\boldsymbol{\alpha}$', 0., 2.))
 pgm.add_node(daft.Node('x', r'$\mathbf{x}_i$', 0., 0.))
 pgm.add_node(daft.Node('data', r'$\mathcal{D}$', 2., 0., observed=True))
 pgm.add_node(daft.Node('N', r'$N$', -2., 0.))

### 2014-06-26 RJObject.h
Some changes to perturb_components, should improve things slightly

@@ -44,7 +44,7 @@ class RJObject
 
 		// Helper methods -- these do one thing at a time
 		double perturb_num_components(double scale);
-		double perturb_components(double chance, double scale);
+		double perturb_components(double chance);
 
 		// Helper methods -- add or remove single component
 		double add_component();

### 2014-06-27 Examples/SineWaves/make_plots.py
Plot posterior for N

@@ -1,0 +1,14 @@
+from pylab import *
+
+rc("font", size=14, family="serif", serif="Computer Sans")
+rc("text", usetex=True)
+
+posterior_sample = loadtxt('posterior_sample.txt')
+
+hist(posterior_sample[:,1005], 50)
+xlabel('Number of Sinusoids $N$', fontsize=14)
+ylabel('Number of posterior samples', fontsize=14)
+xlim([-0.5, 10.5])
+savefig('N_result.pdf', bbox_inches='tight')
+show()
+

### 2014-06-27 Documents/Paper/ms.tex
pgm figure in paper
Results figure
Got a title

@@
 
-\title{RJObject Paper}
+\title{Inference for trans-dimensional Bayesian models with Diffusive Nested
+Sampling}
 \author{Brendon J. Brewer}
@@
 the interim prior for $\{\xx_i\}$ is iid and does not depend on $N$.
+
+\begin{figure}
+\begin{center}
+\includegraphics{pgm.pdf}
+\caption{\it A probabilistic graphical model (PGM) depicting the kind
+of model discussed in this paper. Produced using daft ({\tt daft-pgm.org}).
+\label{fig:pgm}}
+\end{center}
+\end{figure}
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{N_result.pdf}
+\caption{The inference for $N$, the number of sinusoids, based on the
+sinusoid data. The true value was $N=2$.
+\label{fig:N_result}}
+\end{center}
+\end{figure}
+
 %\section{``Transit'' Example}

### 2014-06-27 Documents/Paper/ms.tex
New references

@@
 
+\bibitem[Bretthorst(1988)]{bretthorst} Bretthorst, G.~Larry.\  1988.\
+Bayesian Spectrum Analysis and Parameter Estimation.\ In Lecture Notes in
+Statistics, 48, Springer-Verlag, New York, New York.
+
+\bibitem[Brewer et al.(2007)]{2007ApJ...654..551B} Brewer, B.~J., Bedding, 
+T.~R., Kjeldsen, H., Stello, D.\ 2007.\ Bayesian Inference from 
+Observations of Solar-like Oscillations.\ The Astrophysical Journal 654, 
+551-557.
+
 \bibitem[Brewer et al.(2013)]{2013AJ....146....7B} Brewer, B.~J., 

### 2014-06-27 Documents/Paper/ms.tex
Heavy tailed proposal details

@@
 \section{Metropolis proposals for the general problem}
-We will now define a set of proposal distributions that could be used to
+We will now define a set of proposal distributions that can be used to
 sample the prior distribution
@@
 These proposals will need to satisfy detailed balance in order to be valid
-when used inside DNS. The proposals will also be very heavy tailed. When
+when used inside DNS. When the proposal satisfies detailed balance with respect
+to the prior, the DNS algorithm can incorporate hard likelihood constraints by
+rejecting any proposal whose likelihood does not satisfy the constraint.
+The overall proposal is a mixture of all of the proposals listed here.
+
+When using DNS,
+the proposals typically should be very heavy tailed. When
 DNS is exploring a distribution close to the prior, large proposals will
-generally be needed. When DNS is exploring a distribution constrained to
+generally be needed to explore the prior efficiently.
+When DNS is exploring a distribution constrained to
 very high values of the likelihood function, much smaller proposals will be
@@
 apply the heavy tailed proposal distributions and simply accept that there will
-be some waste.
+be some waste. If a parameter $u$ has a uniform prior between 0 and 1, then
+a heavy-tailed proposal is defined by:
+\begin{eqnarray}
+u' &=& \textnormal{mod}\left(u + 10^{1.5 - 6a}b, 1\right)
+\end{eqnarray}
+where $a \sim \textnormal{Uniform}(0, 1)$ and $b\sim \mathcal{N}(0,1)$. When
+$a$ is close to zero, the size of the proposal is of order 10, which (because
+of the mod function) effectively randomises $u'$ from the prior. If $a$ is
+close to 1, the scale of the proposal is roughly $10^{-4}$ times the prior
+width.
 
 Most of the proposals involve changing one (or a subset) of the parameters
-and/or hyperparameters while keeping the others fixed.
+and/or hyperparameters while keeping the others fixed, and are heavy-t
 

### 2014-06-27 Documents/Paper/ms.tex
More writing about proposal

@@
 \end{eqnarray}
-where $a \sim \textnormal{Uniform}(0, 1)$ and $b\sim \mathcal{N}(0,1)$. When
+where $a \sim \textnormal{Uniform}(0, 1)$ and $b\sim \mathcal{N}(0,1)$.
+This is like a standard gaussian/normal random walk proposal, but with a
+non-fixed step size, resulting in a scale mixture of normals.
+When
 $a$ is close to zero, the size of the proposal is of order 10, which (because
@@
 close to 1, the scale of the proposal is roughly $10^{-4}$ times the prior
-width.
+width. The coefficient of $a$ was set to 6 because smaller proposals are
+usually not necessary. The value 1.5 in the exponent was found to optimize the
+performance of this proposal when sampling from $U(0, 1)$ and $\mathcal{N}(0,1)$
+priors.
 
 Most of the proposals involve changing one (or a subset) of the parameters
-and/or hyperparameters while keeping the others fixed, and are heavy-t
+and/or hyperparameters while keeping the others fixed.
 

### 2014-06-27 Documents/Paper/ms.tex
Nested sampling etc

@@
 the interim prior for $\{\xx_i\}$ is iid and does not depend on $N$.
+Given data $\mathcal{D}$, we usually want to generate samples from the
+posterior distribution given by:
+\begin{eqnarray}
+p(N, \hyperparams, \{\xx_i\} | \mathcal{D}) \propto
+p(N, \hyperparams, \{\xx_i\})
+p(\mathcal{D} | N, \hyperparams, \{\xx_i\})\label{eq:bayes}
+\end{eqnarray}
+The marginal likelihood, or evidence, is also an important quantity, and is
+the normalisation constant of the posterior given in Equation~\ref{eq:bayes}:
+\begin{eqnarray}
+\mathcal{Z} &=& p(\mathcal{D})\\
+&=& \sum_{i=0}^{N_{\rm max}} \int
+p(N, \hyperparams, \{\xx_i\})p(\mathcal{D} | N, \hyperparams, \{\xx_i\})
+\, d^N \xx_i \, d\hyperparams \label{eq:evidence}
+\end{eqnarray}
 
@@
 In other studies, it has been common to do separate runs of Nested Sampling
-with different values of $N$. Then, the posterior for $N$ can be calculated
-based on the estimates of the evidence or marginal likelihood. Our motivation
-for using Nested Sampling is different. With reversible jump MCMC it is possible
-to obtain the posterior for $N$ with a single run. However, the exploration
-of the posterior may be very difficult, and sometimes the problem may even
-contain a phase transition. We use Nested Sampling to overcome these
-difficulties.
+$N$ fixed at various trial values. Then, the posterior for $N$ can be calculated
+based on the estimates of the evidence obtained as a function of $N$. However,
+in this paper, we use trans-dimensional MCMC moves to infer $N$. In other words,
+$N$ is treated as just another parameter, and the marginal likelihood defined
+in Equation~\ref{eq:evidence}, which we can compute, also includes the sum over
+the unknown parameter $N$. Our motivation
+for using Nested Sampling is not primarily to calculate $\mathcal{Z}$.
+The motivation for using Nested Sampling, rather than just sampling the
+posterior using trans-dimensional MCMC, is that the posterior often contains
+difficult features, such as strong dependencies or multiple modes, that cause
+problems with mixing. Diffusive Nested Sampling
+(DNS) replaces the posterior distribution with an alternative target
+distribution composed of a mixture of the prior distribution with other, more
+constrained distributions. This is similar to the idea of simulated annealing,
+where the target distribution is modified to something ``easier'', however,
+the sequence of distributions used in Nested Sampling avoids some of the
+problems with simulated annealing: in particular, there is no need to choose
+a temperature schedule, and the method does not fail when phase transitions
+are present.
 

### 2014-06-27 Documents/Paper/ms.tex
capitalization
Comment about bretthorst

@@
 
-\title{Inference for trans-dimensional Bayesian models with Diffusive Nested
+\title{Inference for Trans-dimensional Bayesian Models with Diffusive Nested
 Sampling}
@@
 and phases $\{\phi_i\}_{i=1}^N$.
-This kind of model has many applications\citep[see e.g.][]{bretthorst}.
+This kind of model has many applications, and has been solved analytically
+under certain sets of assumptions \citep[see e.g.][]{bretthorst}.
 

### 2014-06-28 Examples/Template/MyDistribution.cpp
Gutted template

@@ -1,0 +1,44 @@
+#include "MyDistribution.h"
+#include "RandomNumberGenerator.h"
+#include "Utils.h"
+#include <cmath>
+
+using namespace DNest3;
+
+MyDistribution::MyDistribution()
+{
+
+}
+
+void MyDistribution::fromPrior()
+{
+
+}
+
+double MyDistribution::perturb_parameters()
+{
+	double logH = 0.;
+
+	return logH;
+}
+
+double MyDistribution::log_pdf(const std::vector<double>& vec) const
+{
+	return 0.;
+}
+
+void MyDistribution::from_uniform(std::vector<double>& vec) const
+{
+
+}
+
+void MyDistribution::to_uniform(std::vector<double>& vec) const
+{
+
+}
+
+void MyDistribution::print(std::ostream& out) const
+{
+
+}
+

### 2014-06-28 Examples/Template/MyDistribution.h

@@ -1,0 +1,25 @@
+#ifndef _MyDistribution_
+#define _MyDistribution_
+
+#include <Distributions/Distribution.h>
+
+class MyDistribution:public Distribution
+{
+	private:
+
+		double perturb_parameters();
+
+	public:
+		MyDistribution();
+
+		void fromPrior();
+
+		double log_pdf(const std::vector<double>& vec) const;
+		void from_uniform(std::vector<double>& vec) const;
+		void to_uniform(std::vector<double>& vec) const;
+
+		void print(std::ostream& out) const;
+};
+
+#endif
+

### 2014-06-28 Examples/Template/MyModel.cpp

@@ -1,0 +1,41 @@
+#include "MyModel.h"
+#include "RandomNumberGenerator.h"
+#include "Utils.h"
+#include <cmath>
+
+using namespace std;
+using namespace DNest3;
+
+MyModel::MyModel()
+:objects(3, 10, false, MyDistribution())
+{
+
+}
+
+void MyModel::fromPrior()
+{
+	objects.fromPrior();
+}
+
+double MyModel::perturb()
+{
+	double logH = 0.;
+
+	return logH;
+}
+
+double MyModel::logLikelihood() const
+{
+	return 0.;
+}
+
+void MyModel::print(std::ostream& out) const
+{
+	objects.print(out); out<<' ';
+}
+
+string MyModel::description() const
+{
+	return string("objects");
+}
+

### 2014-06-28 Documents/Paper/ms.tex
Don't go to sampling so soon

@@
 the interim prior for $\{\xx_i\}$ is iid and does not depend on $N$.
-Given data $\mathcal{D}$, we usually want to generate samples from the
+Given data $\mathcal{D}$, we usually want to calculate properties of the
 posterior distribution given by:
@@
 \end{eqnarray}
+The general method used to calculate properties of the posterior distribution
+is to generate samples from it using Markov Chain Monte Carlo (MCMC).
+
 The marginal likelihood, or evidence, is also an important quantity, and is

### 2014-06-28 Documents/Paper/ms.tex
More info in caption

@@
 of model discussed in this paper. Produced using daft ({\tt daft-pgm.org}).
+The number, $N$, of objects in the model is an unknown parameter, as are the
+properties of the objects. The data depends on the properties of the objects.
+Note that it is possible (and common) to have other parameters in the model
+that point directly to the data. However, we have omitted such parameters
+for simplicity.
 \label{fig:pgm}}

### 2014-06-28 Documents/Paper/ms.tex
Adding extra details

@@
 
-\section{Introduction}
+\section{Introduction}\label{sec:introduction}
 \setlength{\parindent}{0cm}
@@
 
-The model for the signal is
+In the notation of Section~\ref{sec:introduction}, the parameters of the
+each object are
+\begin{eqnarray}
+\xx_i &=& \{A_i, T_i, \phi_i\}.
+\end{eqnarray}
+
+The model for the shape of the (noise-free) signal is
 \begin{eqnarray}
@@
 $\{\phi_i\}$.
-
 The sampling distribution (probability distribution for the data given the
@@
 \begin{eqnarray}
-Y_i | N, \{A_i\}, \{T_i\}, \{\phi_i\} \sim
+Y_i | N, \{A_i\}, \{T_i\}, \{\phi_i\}, \sigma \sim
 \mathcal{N}\left(\mu(t_i), \sigma^2\right).
 \end{eqnarray}
-We simulated some data based on the assumption $N=2$.
+This also depends on an additional noise standard deviation parameter $\sigma$
+which will also be inferred.
+
+To demonstrate the techniques,
+we simulated some data based on the assumption $N=2$, i.e. the true signal was
+composed of two sinusoids.
 The simulated data is shown in Figure~\ref{fig:sinewave_data}, and shows a

### 2014-06-28 Documents/Paper/ms.tex
Optimisation strategies

@@
 
+\section{Optimisation Techniques}\label{sec:optimisation}
+In both of the examples discussed in this paper, the likelihood evaluation
+involved computing a mock noise-free signal from the current value of the
+model parameters. The mathematical form of the mock signal was a sum over
+all $N$ components present. Computing the mock signal is often the most
+expensive step in the evaluation of the likelihood.
+
+However, many of the proposal distributions used involve changing a subset
+of the parameters, while keeping others fixed. Considerable speedups can be
+achieved by, when appropriate, updating the mock signal to reflect the proposed
+change to the model, rather than computing the entire mock signal from scratch.
+For example, if the proposal is to add two new objects to the model, the
+mock signal can simply be updated by adding the effect of the two new objects
+to the model.
+
+In general, it is possible to compute the number of components that have been
+affected by the proposal. If this is greater than or equal to $N$, we compute
+the mock signal from scratch, otherwise we subtract the effect of those
+components that have been removed and add in the effect of those that have been
+added.
+
 \section*{Acknowledgements}

### 2014-06-28 Documents/Paper/ms.tex
More sections and details

@@
 
+\subsection{Prior Distributions}
 In the notation of Section~\ref{sec:introduction}, the parameters of the
@@
 \end{eqnarray}
+For the interim prior $p(\xx_i | \hyperparams)$, we introduced a single
+hyperparameter $\mu$, such that $A_i$ given $\mu$ has an exponential prior
+with mean $\mu$. For the periods, we assigned a log-uniform distribution for
+the periods between fixed boundaries, and a uniform distribution for the phases
+between 0 and $2\pi$. In other words, our prior is only hierarchical for the
+amplitudes $\{A_i\}$.
 
@@
 This also depends on an additional noise standard deviation parameter $\sigma$
-which will also be inferred.
+which will also be inferred. We used a log-uniform prior for $\sigma$ between
+$10^{-3}$ and $10^{3}$.
 
+\subsection{Simulated Dataset}
 To demonstrate the techniques,

### 2014-07-03 Examples/GalaxyField/MyDistribution.h
Copy of SineWaves, to start work on GalaxyField

@@ -1,0 +1,35 @@
+#ifndef _MyDistribution_
+#define _MyDistribution_
+
+#include <Distributions/Distribution.h>
+
+// Based on ClassicMassInf1D from RJObject
+// Think of "position x" as log-period
+// and mass as amplitude
+class MyDistribution:public Distribution
+{
+	private:
+		// Limits
+		double x_min, x_max;
+		double mu_min, mu_max;
+
+		// Mean of exponential distribution for amplitudes
+		double mu;
+
+		double perturb_parameters();
+
+	public:
+		MyDistribution(double x_min, double x_max,
+					double mu_min, double mu_max);
+
+		void fromPrior();
+
+		double log_pdf(const std::vector<double>& vec) const;
+		void from_uniform(std::vector<double>& vec) const;
+		void to_uniform(std::vector<double>& vec) const;
+
+		void print(std::ostream& out) const;
+};
+
+#endif
+

### 2014-07-03 Examples/GalaxyField/MyModel.cpp

@@ -1,0 +1,91 @@
+#include "MyModel.h"
+#include "RandomNumberGenerator.h"
+#include "Utils.h"
+#include "Data.h"
+#include <cmath>
+
+using namespace std;
+using namespace DNest3;
+
+MyModel::MyModel()
+:objects(3, 10, false, MyDistribution(-10., 10., 1E-3, 1E3))
+,mu(Data::get_instance().get_t().size())
+{
+
+}
+
+void MyModel::fromPrior()
+{
+	objects.fromPrior();
+	sigma = exp(log(1E-3) + log(1E6)*randomU());
+	calculate_mu();
+}
+
+void MyModel::calculate_mu()
+{
+	// Get the times from the data
+	const vector<double>& t = Data::get_instance().get_t();
+
+	// Get the components
+	const vector< vector<double> >& components = objects.get_components();
+
+	// Zero the signal
+	mu.assign(mu.size(), 0.);
+
+	double T, A, phi;
+	for(size_t j=0; j<components.size(); j++)
+	{
+		T = exp(components[j][0]);
+		A = components[j][1];
+		phi = components[j][2];
+		for(size_t i=0; i<t.size(); i++)
+			mu[i] += A*sin(2.*M_PI*t[i]/T + phi);
+	}
+}
+
+double MyModel::perturb()
+{
+	double logH = 0.;
+
+	if(randomU() <= 0.75)
+	{
+		logH += objects.perturb();
+		calculate_mu();
+	}
+	else
+	{
+		sigma = log(sigma);
+		sigma += log(1E6)*randh();
+		sigma = mod(sigma - log(1E-3), log(1E6)) + log(1E-3);
+		sigma = exp(sigma);
+	}
+
+	return logH;
+}
+
+double MyModel::logLikelihood() const
+{
+	// Get the data
+	const vector<double>& y = Data::get_instance().get_y();
+
+	double logL = 0.;
+	double var = sigma*sigma;
+	for(size_t i=0; i<y.size(); i++)
+		logL += -0.5*log(2.*M_PI*var) - 0.5*pow(y[i] - mu[i], 2)/var;
+
+	return logL;
+}
+
+void MyModel::print(std::ostream& out) const
+{
+	for(size_t i=0; i<mu.size(); i++)
+		out<<mu[i]<<' ';
+	out<<sigma<<' ';
+	objects.print(out); out<<' ';
+}
+
+string MyModel::description() const
+{
+	return string("objects, sigma");
+}
+

### 2014-07-03 Examples/GalaxyField/MyModel.h

@@ -1,0 +1,39 @@
+#ifndef _MyModel_
+#define _MyModel_
+
+#include "Model.h"
+#include <vector>
+#include <RJObject.h>
+#include "MyDistribution.h"
+
+class MyModel:public DNest3::Model
+{
+	private:
+		RJObject<MyDistribution> objects;
+		double sigma; // Noise standard deviation
+
+		// The signal
+		std::vector<double> mu;
+		void calculate_mu();
+
+	public:
+		MyModel();
+
+		// Generate the point from the prior
+		void fromPrior();
+
+		// Metropolis-Hastings proposals
+		double perturb();
+
+		// Likelihood function
+		double logLikelihood() const;
+
+		// Print to stream
+		void print(std::ostream& out) const;
+
+		// Return string with column information
+		std::string description() const;
+};
+
+#endif
+

### 2014-07-03 Examples/GalaxyField/showresults.py

@@ -1,0 +1,3 @@
+import postprocess
+postprocess.postprocess()
+

### 2014-07-04 Examples/GalaxyField/Data.cpp
Data class copied from Lensing2, but without PSF stuff

@@ -1,6 +1,7 @@
 #include "Data.h"
-#include <fstream>
 #include <iostream>
+#include <fstream>
+#include <cmath>
 
 using namespace std;
 
@@ -11,27 +12,75 @@ Data::Data()
 
 }
 
-void Data::load(const char* filename)
+void Data::load(const char* metadata_file, const char* image_file,
+			const char* sigma_file, const char* psf_file)
 {
-	fstream fin(filename, ios::in);
+	/*
+	* First, read in the metadata
+	*/
+	int psf_size;
+	fstream fin(metadata_file, ios::in);
 	if(!fin)
-	{
-		cerr<<"# Error. Couldn't open file "<<filename<<"."<<endl;
-		return;
-	}
-
-	// Empty the vectors
-	t.clear();
-	y.clear();
-
-	double temp1, temp2;
-	while(fin>>temp1 && fin>>temp2)
-	{
-		t.push_back(temp1);
-		y.push_back(temp2);
-	}
-	cout<<"# Loaded "<<t.size()<<" data points from file "
-			<<filename<<"."<<endl;
+		cerr<<"# ERROR: couldn't open file "<<metadata_file<<"."<<endl;
+	fin>>ni>>nj;
+	fin>>x_min>>x_max>>y_min>>y_max;
 	fin.close();
+
+	// Make sure maximum > minimum
+	if(x_max <= x_min || y_max <= y_min)
+		cerr<<"# ERROR: strange input in "<<metadata_file<<"."<<endl;
+
+	// Compute pixel widths
+	dx = (x_max - x_min)/nj;
+	dy = (y_max - y_min)/ni;
+
+	// Check that pixels are square
+	if(abs(log(dx/dy)) >= 1E-3)
+		cerr<<"# ERROR: pixels aren't square."<<endl;
+
+	compute_ray_grid();
+
+	/*
+	* Now, load the image
+	*/
+	fin.open(image_file, ios::in);
+	if(!fin)
+		cerr<<"# ERROR: couldn't open file "<<image_file<<"."<<endl;
+	image.assign(ni, vector<double>(nj));
+	for(size_t i=0; i<image.size(); i++)
+		for(size_t j=0; j<image[i].size(); j++)
+			fin>>image[i][j];
+	fin.close();
+
+	/*
+	* Load the sigma map
+	*/
+	fin.open(sigma_file, ios::in);
+	if(!fin)
[…]

### 2014-07-04 Examples/GalaxyField/Data.h

@@ -6,15 +6,47 @@
 class Data
 {
 	private:
-		std::vector<double> t, y;
+		// Number of pixels
+		int ni, nj;
+
+		// Coordinates of image boundaries
+		double x_min, x_max, y_min, y_max;
+
+		// Pixel widths
+		double dx, dy;
+
+		// Coordinates of pixel centers
+		std::vector< std::vector<double> > x_rays;
+		std::vector< std::vector<double> > y_rays;
+
+		// The pixels
+		std::vector< std::vector<double> > image;
+
+		// Sigma map
+		std::vector< std::vector<double> > sigma;
+
+		void compute_ray_grid();
 
 	public:
 		Data();
-		void load(const char* filename);
+		void load(const char* metadata_file, const char* image_file,
+				const char* sigma_file, const char* psf_file);
 
 		// Getters
-		const std::vector<double>& get_t() const { return t; }
-		const std::vector<double>& get_y() const { return y; }
+		int get_ni() const { return ni; }
+		int get_nj() const { return nj; }
+		double get_x_min() const { return x_min; }
+		double get_x_max() const { return x_max; }
+		double get_y_min() const { return y_min; }
+		double get_y_max() const { return y_max; }
+		const std::vector< std::vector<double> >& get_x_rays() const
+			{ return x_rays; }
+		const std::vector< std::vector<double> >& get_y_rays() const
+			{ return y_rays; }
+		const std::vector< std::vector<double> >& get_image() const
+			{ return image; }
+		const std::vector< std::vector<double> >& get_sigma() const
+			{ return sigma; }
 
 	// Singleton
 	private:

### 2014-07-04 Examples/GalaxyField/MyDistribution.cpp
Starting work on hyperparameters

@@ -6,61 +6,48 @@
 using namespace DNest3;
 
 MyDistribution::MyDistribution(double x_min, double x_max,
-					double mu_min, double mu_max)
+					double y_min, double y_max,
+					double fluxlim_min, double fluxlim_max)
 :x_min(x_min)
 ,x_max(x_max)
-,mu_min(mu_min)
-,mu_max(mu_max)
+,y_min(y_min)
+,y_max(y_max)
+,fluxlim_min(fluxlim_min)
+,fluxlim_max(fluxlim_max)
 {
 
 }
 
 void MyDistribution::fromPrior()
 {
-	mu = exp(log(mu_min) + log(mu_max/mu_min)*randomU());
+
 }
 
 double MyDistribution::perturb_parameters()
 {
 	double logH = 0.;
 
-	mu = log(mu);
-	mu += log(mu_max/mu_min)*pow(10., 1.5 - 6.*randomU())*randn();
-	mu = mod(mu - log(mu_min), log(mu_max/mu_min)) + log(mu_min);
-	mu = exp(mu);
-
 	return logH;
 }
 
-// vec[0] = "position" (log-period)
-// vec[1] = amplitude
-// vec[2] = phase
-
+// x, y, flux
 double MyDistribution::log_pdf(const std::vector<double>& vec) const
 {
-	if(vec[0] < x_min || vec[0] > x_max || vec[1] < 0. ||
-			vec[2] < 0. || vec[2] > 2.*M_PI)
-		return -1E300;
-
-	return -log(mu) - vec[1]/mu;
+	return 0.;
 }
 
 void MyDistribution::from_uniform(std::vector<double>& vec) const
 {
-	vec[0] = x_min + (x_max - x_min)*vec[0];
-	vec[1] = -mu*log(1. - vec[1]);
-	vec[2] = 2.*M_PI*vec[2];
+
 }
 
 void MyDistribution::to_uniform(std::vector<double>& vec) const
 {
-	vec[0] = (vec[0] - x_min)/(x_max - x_min);
-	vec[1] = 1. - exp(-vec[1]/mu);
-	vec[2] = vec[2]/(2.*M_PI);
+
 }
 
 void MyDistribution::print(std::ostream& out) const
 {
-	out<<mu<<' ';
+
 }
 

### 2014-07-04 Examples/GalaxyField/MyDistribution.h

@@ -3,24 +3,25 @@
 
 #include <Distributions/Distribution.h>
 
-// Based on ClassicMassInf1D from RJObject
-// Think of "position x" as log-period
-// and mass as amplitude
+// Hyperparameters setting interim prior for galaxy properties
 class MyDistribution:public Distribution
 {
 	private:
 		// Limits
-		double x_min, x_max;
-		double mu_min, mu_max;
+		double x_min, x_max, y_min, y_max;
+		double fluxlim_min, fluxlim_max;
 
-		// Mean of exponential distribution for amplitudes
-		double mu;
+		// Lower limit for Pareto dist
+		double fluxlim;
+		// Slope for Pareto dist
+		double alpha;
 
 		double perturb_parameters();
 
 	public:
 		MyDistribution(double x_min, double x_max,
-					double mu_min, double mu_max);
+					double y_min, double y_max,
+					double fluxlim_min, double fluxlim_max);
 
 		void fromPrior();
 

### 2014-07-04 Examples/GalaxyField/MyDistribution.h
Pareto stuff

@@ -13,8 +13,8 @@ class MyDistribution:public Distribution
 
 		// Lower limit for Pareto dist
 		double fluxlim;
-		// Slope for Pareto dist
-		double alpha;
+		// 1/slope for Pareto dist
+		double gamma;
 
 		double perturb_parameters();
 

### 2014-07-05 Examples/GalaxyField/MyDistribution.cpp
Proposals for two hyperparameters

@@ -20,12 +20,29 @@ MyDistribution::MyDistribution(double x_min, double x_max,
 
 void MyDistribution::fromPrior()
 {
-
+	fluxlim = exp(log(fluxlim_min) + log(fluxlim_max/fluxlim_min)*rand());
+	gamma = 2.*randomU();
 }
 
 double MyDistribution::perturb_parameters()
 {
 	double logH = 0.;
+
+	int which = randInt(2);
+
+	if(which == 0)
+	{
+		fluxlim = log(fluxlim);
+		fluxlim += log(fluxlim_max/fluxlim_min)*randh();
+		fluxlim = mod(fluxlim - log(fluxlim_min),
+			log(fluxlim_max/fluxlim_min)) + log(fluxlim_min);
+		fluxlim = exp(fluxlim);
+	}
+	else
+	{
+		gamma += 2.*randh();
+		gamma = mod(gamma, 2.);
+	}
 
 	return logH;
 }
@@ -48,6 +65,6 @@ void MyDistribution::to_uniform(std::vector<double>& vec) const
 
 void MyDistribution::print(std::ostream& out) const
 {
-
+	out<<fluxlim<<' '<<gamma<<' ';
 }
 

### 2014-07-09 Examples/GalaxyField/Frames/make_movie.sh
Good old movie making scripts

@@ -1,0 +1,5 @@
+#!/bin/bash
+avconv -r 10 -f image2 -i '%04d.png' -b:v 4096k movie.avi
+
+
+

### 2014-07-09 Examples/GalaxyField/display.py

@@ -1,5 +1,11 @@
 from pylab import *
 
+import os
+
+saveFrames = False # For making movies
+if saveFrames:
+	os.system('rm Frames/*.png')
+
 posterior_sample = atleast_2d(loadtxt('posterior_sample.txt'))
 data = loadtxt('Data/test_image.txt')
 sig = loadtxt('Data/test_sigma.txt')
@@ -16,6 +22,10 @@ for i in xrange(0, posterior_sample.shape[0]):
 	imshow((img - data)/sigma)
 	title('Standardised Residuals')
 	draw()
+
+	if saveFrames:
+		savefig('Frames/' + '%0.4d'%(i+1) + '.png', bbox_inches='tight')
+		print('Frames/' + '%0.4d'%(i+1) + '.png')
 
 ioff()
 show()

### 2014-07-09 Documents/Paper/ms.tex
sentence
Pareto
Pareto stuff
GalaxyField hyperparameters

@@
 \maketitle
-\abstract{This is the abstract.}
+\abstract{Blah...
+We demonstrate the techniques on a seemingly simple problem of detecting
+sinusoidal signals in noise, which exhibits a phase transition.}
 
@@
 %\section{``Asteroseismology'' Example}
-%\section{``Galaxy Field'' Example}
+\section{``Galaxy Field'' Example}
+
+We model each galaxy as a mixture of two concentric elliptical galaxies, with
+the following eight parameters. Firstly, two
+parameters $x_c$ and $y_c$ describe the central position of the galaxy within
+the image. The flux $f$ describes the total integral of the galaxy's
+intensity profile. The axis ratio $q$ describes the ellipticity of the galaxy
+and $\theta$ its orientation angle with respect to horizontal. Finally
+we include a ``radius ratio'' $u$ describing the ratio of the radius of the
+smaller gaussian with respect to that of the bigger gaussian, and a parameter
+$v$ describing the fraction of the total flux in the outer gaussian
+(therefore the fraction of the light in the smaller gaussian is $1-f$).
+
+For the interim prior for the total fluxes of the galaxies, and for the
+widths of the galaxies, we use the Pareto
+distribution. The probability density function for a variable $x$ given
+hyperparameters $x_{\rm min}$ and $a$ is
+\begin{eqnarray}
+p(x | x_{\rm min}, a) &=&
+\left\{
+\begin{array}{lr}
+\frac{a x_{\rm min}^{a}}{x^{a + 1}}, & x \geq x_{\rm min}\\
+0, & \textnormal{otherwise.}
+\end{array}
+\right.
+\end{eqnarray}
+Since we are using this for both the fluxes and the widths of the galaxies,
+there will be four hyperparameters in total for the fluxes and widths.
+
+% Pareto in terms of log(x)
+% p(x) ~ x^(-(alpha + 1))
+% Let u = log(x), du = 1/x dx
+% p(u) = e^(-alpha*u)
 

### 2014-07-09 Documents/Paper/ms.tex
Fixed slight mistake

@@
 $v$ describing the fraction of the total flux in the outer gaussian
-(therefore the fraction of the light in the smaller gaussian is $1-f$).
+(therefore the fraction of the light in the smaller gaussian is $1-v$).
 

### 2014-07-09 Examples/GalaxyField/make_plots.py
Plot GalaxyField data

@@ -1,0 +1,27 @@
+from pylab import *
+
+rc("font", size=14, family="serif", serif="Computer Sans")
+rc("text", usetex=True)
+
+# Piecewise linear stretch
+def stretch(x):
+	y = x.copy()
+	y = (y - y.min())/(y.max() - y.min())
+	y[y > 0.1] = 0.1 + 0.05*(y[y > 0.1] - 0.1)
+	return y
+
+data = loadtxt('Data/test_image.txt')
+
+img = stretch(data)
+
+imshow(img, cmap='gray')
+xlabel('$x$')
+ylabel('$y$')
+gca().set_xticks([-0.5, 99.5, 199.5])
+gca().set_yticks([-0.5, 99.5, 199.5])
+gca().set_xticklabels(['-1', '0', '1'])
+gca().set_yticklabels(['1', '0', '-1'])
+title('GalaxyField Data')
+savefig('image.pdf', bbox_inches='tight')
+show()
+

### 2014-07-09 Examples/GalaxyField/make_plots.py
GalaxyField data figure

@@ -14,7 +14,7 @@ data = loadtxt('Data/test_image.txt')
 
 img = stretch(data)
 
-imshow(img, cmap='gray')
+imshow(img, cmap='gray', interpolation='nearest')
 xlabel('$x$')
 ylabel('$y$')
 gca().set_xticks([-0.5, 99.5, 199.5])
@@ -22,6 +22,6 @@ gca().set_yticks([-0.5, 99.5, 199.5])
 gca().set_xticklabels(['-1', '0', '1'])
 gca().set_yticklabels(['1', '0', '-1'])
 title('GalaxyField Data')
-savefig('image.pdf', bbox_inches='tight')
+savefig('galaxyfield_data.pdf', bbox_inches='tight')
 show()
 

### 2014-07-10 Examples/GalaxyField/display.py
Use stretch

@@ -2,6 +2,13 @@ from pylab import *
 
 import os
 
+# Piecewise linear stretch
+def stretch(x):
+	y = x.copy()
+	y = (y - y.min())/(y.max() - y.min())
+	y[y > 0.1] = 0.1 + 0.05*(y[y > 0.1] - 0.1)
+	return y
+
 saveFrames = False # For making movies
 if saveFrames:
 	os.system('rm Frames/*.png')
@@ -15,12 +22,22 @@ hold(False)
 for i in xrange(0, posterior_sample.shape[0]):
 	img = posterior_sample[i, 0:200**2].reshape((200, 200))
 	subplot(1, 2, 1)
-	imshow(img)
+	imshow(stretch(img))
 	title('Model {i}'.format(i=i))
+	gca().set_xticks([-0.5, 99.5, 199.5])
+	gca().set_yticks([-0.5, 99.5, 199.5])
+	gca().set_xticklabels(['-1', '0', '1'])
+	gca().set_yticklabels(['1', '0', '-1'])
+
 	subplot(1, 2, 2)
 	sigma = sqrt(sig**2 + posterior_sample[i,-2]**2)
 	imshow((img - data)/sigma)
 	title('Standardised Residuals')
+	gca().set_xticks([-0.5, 99.5, 199.5])
+	gca().set_yticks([-0.5, 99.5, 199.5])
+	gca().set_xticklabels(['-1', '0', '1'])
+	gca().set_yticklabels(['1', '0', '-1'])
+
 	draw()
 
 	if saveFrames:

### 2014-07-10 Examples/GalaxyField/Frames/make_movie.sh
frame rate 5

@@ -1,5 +1,5 @@
 #!/bin/bash
-avconv -r 10 -f image2 -i '%04d.png' -b:v 4096k movie.avi
+avconv -r 5 -f image2 -i '%04d.png' -b:v 4096k movie.avi
 
 
 

### 2014-07-17 Documents/Paper/ms.tex
GalaxyField data figure
An abstract

@@
 \maketitle
-\abstract{Blah...
-We demonstrate the techniques on a seemingly simple problem of detecting
-sinusoidal signals in noise, which exhibits a phase transition.}
+\abstract{Many inference problems require us to infer the number $N$ of
+objects in some region, along with their properties $\{\xx_i\}_{i=1}^N$,
+from a dataset $\mathcal{D}$. These problems are typically solved using one
+of the following two methods: i)
+by executing a Monte Carlo algorithm (such as Nested Sampling) once for each
+possible value of $N$,
+and calculating the marginal likelihood or evidence as a function of $N$; or
+ii) by doing a single run that allows the model dimension $N$ to change (such as
+Markov Chain Monte Carlo with birth/death moves), and obtaining the posterior
+for $N$ directly. In this paper we present a general approach to this problem
+that uses trans-dimensional MCMC embedded {\it within} a Nested Sampling
+algorithm, allowing us to explore the posterior distribution and calculate
+the marginal likelihood (summed over $N$) even if the problem contains a phase
+transition. We present two example problems, finding sinusoidal signals in
+noisy data, and finding and measuring galaxies in a noisy astronomical image.}
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.7]{galaxyfield_data.pdf}
+\caption{The simulated data for the GalaxyField example.
+\label{fig:galaxyfield_data}}
+\end{center}
+\end{figure}
+
 % Pareto in terms of log(x)

### 2014-07-22 Documents/Paper/ms.tex
Slight tweak in abstract

@@
 the marginal likelihood (summed over $N$) even if the problem contains a phase
-transition. We present two example problems, finding sinusoidal signals in
+transition or other difficult features such as multimodality.
+We present two example problems, finding sinusoidal signals in
 noisy data, and finding and measuring galaxies in a noisy astronomical image.}

### 2014-07-23 Examples/SineWaves/postprocess.py
So I can tweak the figures

@@ -1,0 +1,226 @@
+# Copyright (c) 2009, 2010, 2011, 2012 Brendon J. Brewer.
+#
+# This file is part of DNest3.
+#
+# DNest3 is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+#
+# DNest3 is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with DNest3. If not, see <http://www.gnu.org/licenses/>.
+
+import copy
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+matplotlib.rc("font", family="serif", size=12)
+matplotlib.rc("text", usetex=True)
+
+def logsumexp(values):
+	biggest = np.max(values)
+	x = values - biggest
+	result = np.log(np.sum(np.exp(x))) + biggest
+	return result
+
+def logdiffexp(x1, x2):
+	biggest = x1
+	xx1 = x1 - biggest
+	xx2 = x2 - biggest
+	result = np.log(np.exp(xx1) - np.exp(xx2)) + biggest
+	return result
+
+def postprocess(temperature=1., numResampleLogX=1, plot=True, loaded=[], \
+			cut=0., save=True, zoom_in=True):
+	if len(loaded) == 0:
+		levels = np.atleast_2d(np.loadtxt("levels.txt"))
+		sample_info = np.atleast_2d(np.loadtxt("sample_info.txt"))
+		sample = np.atleast_2d(np.loadtxt("sample.txt"))
+		#if(sample.shape[0] == 1):
+		#	sample = sample.T
+	else:
+		levels, sample_info, sample = loaded[0], loaded[1], loaded[2]
+
+	sample = sample[int(cut*sample.shape[0]):, :]
+	sample_info = sample_info[int(cut*sample_info.shape[0]):, :]
+
+	if sample.shape[0] != sample_info.shape[0]:
+		print('# Size mismatch. Truncating...')
[…]

### 2014-07-23 Examples/SineWaves/postprocess.py
Nicer plot for paper

@@ -158,16 +158,17 @@ def postprocess(temperature=1., numResampleLogX=1, plot=True, loaded=[], \
 			plt.figure(3)
 			if z == 0:
 				plt.subplot(2,1,1)
-				plt.plot(logx_samples[:,z], sample_info[:,1], 'b.', label='Samples')
+				plt.plot(logx_samples[:,z], sample_info[:,1], 'b.', markersize=1, label='Samples')
 				plt.hold(True)
 				plt.plot(levels[1:,0], levels[1:,1], 'r.', label='Levels')
 				plt.legend(numpoints=1, loc='lower left')
-				plt.ylabel('log(L)')
-				plt.title(str(z+1) + "/" + str(numResampleLogX) + ", log(Z) = " + str(logz_estimates[z][0]))
+				plt.title('Likelihood Curve')
+				plt.ylabel(r'$\log(L)$')
+
 				# Use all plotted logl values to set ylim
 				combined_logl = np.hstack([sample_info[:,1], levels[1:, 1]])
 				combined_logl = np.sort(combined_logl)
-				lower = combined_logl[int(0.05*combined_logl.size)]
+				lower = combined_logl[int(0.07*combined_logl.size)]
 				upper = combined_logl[-1]
 				diff = upper - lower
 				lower -= 0.05*diff
@@ -182,12 +183,14 @@ def postprocess(temperature=1., numResampleLogX=1, plot=True, loaded=[], \
 		if plot:
 			plt.subplot(2,1,2)
 			plt.hold(False)
-			plt.plot(logx_samples[:,z], P_samples[:,z], 'b.')
+			plt.plot(logx_samples[:,z], P_samples[:,z], 'b.', markersize=1)
 			plt.ylabel('Posterior Weights')
-			plt.xlabel('log(X)')
+			plt.xlabel(r'$\log(X)$')
 			plt.xlim(xlim)
 			if numResampleLogX > 1:
 				plt.draw()
+
+			plt.savefig('sinewaves_likelihood.pdf')
 
 	P_samples = np.mean(P_samples, 1)
 	P_samples = P_samples/np.sum(P_samples)

### 2014-07-23 Examples/SineWaves/postprocess.py
New figure

@@ -190,7 +190,7 @@ def postprocess(temperature=1., numResampleLogX=1, plot=True, loaded=[], \
 			if numResampleLogX > 1:
 				plt.draw()
 
-			plt.savefig('sinewaves_likelihood.pdf')
+			plt.savefig('sinewaves_likelihood.pdf', bbox_inches='tight')
 
 	P_samples = np.mean(P_samples, 1)
 	P_samples = P_samples/np.sum(P_samples)

### 2014-07-27 Documents/Paper/ms.tex
New figure
Some edits

@@
 \maketitle
-\abstract{Many inference problems require us to infer the number $N$ of
+\abstract{Many inference problems involve inferring the number $N$ of
 objects in some region, along with their properties $\{\xx_i\}_{i=1}^N$,
@@
 $\xx$, which may be a single scalar value (for example, a mass), or a set of
-values (e.g. a position and a mass).
+values (e.g. a two-dimensional position and a mass). These problems are often
+challenging due to the unknown (and potentially large) dimensionality.
 
-We would like to assign a prior to the objects' properties $\{\xx_i\}_{i=1}^N$.
+To do the inference, we must assign a prior to the objects' properties
+$\{\xx_i\}_{i=1}^N$.
 Since $N$ may be large, it is usually easier to assign an ``interim prior''
@@
 $\hyperparams$. This kind of model is usually called {\it hierarchical}.
-
 The prior for $N$, $\hyperparams$, and $\{\xx_i\}$ is usually factorised
@@
 The general method used to calculate properties of the posterior distribution
-is to generate samples from it using Markov Chain Monte Carlo (MCMC).
+is to generate samples from it using Markov Chain Monte Carlo (MCMC). The
+structure of the model is depicted graphically in Figure~\ref{fig:pgm}.
 
@@
 
-In other studies, it has been common to do separate runs of Nested Sampling
-$N$ fixed at various trial values. Then, the posterior for $N$ can be calculated
+In other studies, it has been common to do separate runs of Nested Sampling,
+with $N$ fixed at various trial values.
+Then, the posterior for $N$ can be calculated
 based on the estimates of the evidence obtained as a function of $N$. However,
@@
 distribution composed of a mixture of the prior distribution with other, more
-constrained distributions. This is similar to the idea of simulated annealing,
-where the target distribution is modified to something ``easier'', however,
+constrained distributions, facilitating mixing between the multiple modes
+and along degeneracy curves.
+This is similar to the idea of simulated annealing,
+where the target distribution is modified from the posterior to something
+``easier'', however,
 the sequence of distributions used in Nested Sampling avoids some of the
@@
 \begin{center}
+\includegraphics[scale=0.5]{sinewaves_likelihood.pdf}
+\caption{The inference for $N$, the number of sinusoids, based on the
+sinusoid data. The true value was $N=2$.
+\label{fig:sinewaves_likelihood}}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
 \includegraphics[scale=0.5]{N_result.pdf}

### 2014-07-28 Documents/Paper/ms.tex
A word changed
Paragraphs talking about sinewaves results

@@
 be some waste. If a parameter $u$ has a uniform prior between 0 and 1, then
-a heavy-tailed proposal is defined by:
+a particular choice of heavy-tailed proposal is defined by:
 \begin{eqnarray}
@@
 
+\subsection{Results}
+We ran DNS on the simulated data shown in Figure~\ref{fig:sinewave_data}.
+The shape of the likelihood function, with respect to the prior mass (the
+usual likelihood curve associated with Nested Sampling), us shown in
+Figure~\ref{fig:sinewaves_likelihood}. Two phase transitions can be seen. The
+first, at $\log(X) \approx -10$ nats,
+separates ``noise-only'' models (where the
+entire dataset is accounted for by the noise term) from ``one-sinusoid'' models
+where $N=1$. At $\log(X) \approx -28$ nats, a second phase transition separates
+$N=1$ models from $N=2$ models.
+
+If we were to do this analysis by trying to explore the posterior directly,
+rather than by using Nested Sampling, it would be difficult to jump between
+the $N=1$ solution and the $N=2$ solution, because the posterior distribution
+would be composed of a mixture of a small-volume but high-likelihood spike
+and a large-volume but (relatively) low likelihood
+slab. Note that this phase transition still exists if we condition on $N=2$,
+so is not caused by the trans-dimensional model.
+
 \begin{figure}

### 2014-07-28 Documents/Paper/ms.tex
Caption

@@
 \includegraphics[scale=0.5]{sinewaves_likelihood.pdf}
-\caption{The inference for $N$, the number of sinusoids, based on the
-sinusoid data. The true value was $N=2$.
+\caption{{\bf Top panel: }
+The shape of the likelihood function with respect to prior mass.
+{\bf Bottom panel: }The posterior weights of the saved samples. The existence
+of a phase transition causes this to display two separate peaks, which
+would be difficult to mix between if we simply attempted to sample the posterior
+distribution.
 \label{fig:sinewaves_likelihood}}

### 2014-07-28 Documents/Paper/ms.tex
Some refs

@@
 \section{``Galaxy Field'' Example}
+Source detection is an important problem in astrophysics. Given some data,
+usually one or more noisy, blurred, images of the sky, we would like to know
+how many objects $N$ (such as stars or galaxies) are in the image, and their
+properties (such as flux, size, orientation). Various ad-hoc approaches exist,
+covering a wide spectrum between ad-hoc approaches and principled inference
+approaches
+\citep[e.g.][]{irwin, sextractor, dolphot, 2003MNRAS.338..765H, starfield}.
+
+In this section we apply
+the DNS approach to a toy version of the problem of detecting and quantifying
+extended objects such as galaxies in a noisy image.
 
@@
 \begin{thebibliography}{}
+\bibitem[Bertin and
+Arnouts(1996)]{sextractor} Bertin, E., Arnouts, S.\ 1996.\ SExtractor: Software for source extraction.\ Astronomy and Astrophysics Supplement Series 117, 393-404.
 
@@
 
-\bibitem[Brewer et al.(2013)]{2013AJ....146....7B} Brewer, B.~J., 
+\bibitem[Brewer et al.(2013)]{starfield} Brewer, B.~J., 
 Foreman-Mackey, D., Hogg, D.~W.\ 2013.\ Probabilistic Catalogs for Crowded 
@@
 
+\bibitem[Dolphin(2000)]{dolphot} Dolphin, A.~E.\ 2000, PASP, 
+112, 1383
+
 \bibitem[\protect\citeauthoryear{Green}{1995}]{rjmcmc}
 Green, P.~J., 1995, Reversible Jump Markov Chain Monte Carlo Computation and Bayesian Model Determination, Biometrika 82 (4): 711–732.
+
+\bibitem[Hobson 
+\& McLachlan(2003)]{2003MNRAS.338..765H} Hobson, M.~P., \& McLachlan, C.\ 2003, \mnras, 338, 765
+
+\bibitem[Irwin(1985)]{irwin} Irwin, M.~J.\ 1985, \mnras, 214,
+575
 

### 2014-07-28 Documents/Paper/ms.tex
No MNRAS control sequence
Feroz paper
More talk

@@
 This kind of model has many applications, and has been solved analytically
-under certain sets of assumptions \citep[see e.g.][]{bretthorst}.
+under certain sets of assumptions \citep[see e.g.][]{bretthorst}. Similar
+models have been used in many different fields
+\citep[e.g.][]{}. Note that this problem is also very closely related to the
+problem of detecting exoplanet signals in radial velocity data, which has
+attracted a lot of interest
+\citep[e.g.][]{2007ApJ...654..551B, gregory, fengji, 2011MNRAS.415.3462F}.
 
@@
 
+\bibitem[Feroz et al.(2011)]{2011MNRAS.415.3462F} Feroz, F., Balan, S.~T.,
+Hobson, M.~P.\ 2011.\ Detecting extrasolar planets from stellar radial
+velocities using Bayesian evidence.\ Monthly Notices of the Royal
+Astronomical Society 415, 3462-3472.
+
 \bibitem[\protect\citeauthoryear{Green}{1995}]{rjmcmc}
@@
 
+\bibitem[Gregory(2011)]{gregory} Gregory, P.~C.\ 2011.\ 
+Bayesian exoplanet tests of a new method for MCMC sampling in highly 
+correlated model parameter spaces.\ Monthly Notices of the Royal 
+Astronomical Society 410, 94-110.
+
 \bibitem[Hobson 
-\& McLachlan(2003)]{2003MNRAS.338..765H} Hobson, M.~P., \& McLachlan, C.\ 2003, \mnras, 338, 765
+\& McLachlan(2003)]{2003MNRAS.338..765H} Hobson, M.~P., \& McLachlan, C.\ 2003, MNRAS, 338, 765
 
-\bibitem[Irwin(1985)]{irwin} Irwin, M.~J.\ 1985, \mnras, 214,
+\bibitem[Hou et al.(2014)]{fengji} Hou, F., Goodman, J., Hogg, 
+D.~W.\ 2014.\ The Probabilities of Orbital-Companion Models for Stellar 
+Radial Velocity Data.\ ArXiv e-prints arXiv:1401.6128.
+
+\bibitem[Irwin(1985)]{irwin} Irwin, M.~J.\ 1985, MNRAS, 214,
 575

### 2014-07-28 Documents/Paper/ms.tex
More citations

@@
 models have been used in many different fields
-\citep[e.g.][]{}. Note that this problem is also very closely related to the
+\citep[e.g.][]{2005PhRvD..72b2001U, 2007ApJ...654..551B, 2003AIPC..659....3B}.
+Note that this problem is also very closely related to the
 problem of detecting exoplanet signals in radial velocity data, which has
-attracted a lot of interest
+attracted a lot of research attention in recent years
 \citep[e.g.][]{2007ApJ...654..551B, gregory, fengji, 2011MNRAS.415.3462F}.
@@
 
+\bibitem[Bretthorst(2003)]{2003AIPC..659....3B} Bretthorst, G.~L.\ 2003.\ 
+Frequency Estimation, Multiple Stationary Nonsinusoidal Resonances With 
+Trend.\ Bayesian Inference and Maximum Entropy Methods in Science and 
+Engineering 659, 3-22.
+
 \bibitem[Brewer et al.(2007)]{2007ApJ...654..551B} Brewer, B.~J., Bedding, 
@@
 
+\bibitem[Umst{\"a}tter et al.(2005)]{2005PhRvD..72b2001U} Umst{\"a}tter, 
+R., Christensen, N., Hendry, M., Meyer, R., Simha, V., Veitch, J., 
+Vigeland, S., Woan, G.\ 2005.\ Bayesian modeling of source confusion in 
+LISA data.\ Physical Review D 72, 022001.
+
 \end{thebibliography}

### 2014-07-28 Documents/Paper/ms.tex
Slightly narrower margins
New figure

@@
 \usepackage{natbib}
-\usepackage[left=3cm,top=3cm,right=3cm]{geometry}
+\usepackage[left=2.5cm,top=2.5cm,right=2.5cm]{geometry}
 \usepackage{parskip}
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{galaxyfield_likelihood.pdf}
+\caption{{\bf Top panel: }
+The shape of the likelihood function with respect to prior mass for the
+GalaxyField data.
+{\bf Bottom panel: }The posterior weights of the saved samples. Note the phase
+transition at $\log(X) \approx -100$ nats. While this doesn't affect the
+posterior distribution (unlike the sinewave example), it would affect calculation
+of the marginal likelihood if annealing were used.
+\label{fig:sinewaves_likelihood}}
+\end{center}
+\end{figure}
+
 % Pareto in terms of log(x)

### 2014-07-29 RJObjectImpl.h
Don't force 12 sig figs

@@ -218,7 +218,6 @@ double RJObject<Distribution>::remove_component()
 template<class Distribution>
 void RJObject<Distribution>::print(std::ostream& out) const
 {
-	out<<std::setprecision(12);
 	out<<num_dimensions<<' '<<max_num_components<<' ';
 	dist.print(out); out<<' ';
 	out<<num_components<<' ';

### 2014-07-31 Documents/Paper/ms.tex
Equation for galaxy

@@
 extended objects such as galaxies in a noisy image.
-
-We model each galaxy as a mixture of two concentric elliptical galaxies, with
+We model each galaxy as a mixture of two concentric elliptical components, with
 the following eight parameters. Firstly, two
@@
 intensity profile. The axis ratio $q$ describes the ellipticity of the galaxy
-and $\theta$ its orientation angle with respect to horizontal. Finally
+and $\theta$ its orientation angle with respect to horizontal. The radius of the
+bigger gaussian is a parameter $a$.
+Finally
 we include a ``radius ratio'' $u$ describing the ratio of the radius of the
 smaller gaussian with respect to that of the bigger gaussian, and a parameter
-$v$ describing the fraction of the total flux in the outer gaussian
-(therefore the fraction of the light in the smaller gaussian is $1-v$).
+$v$ describing the fraction of the total flux in the smaller gaussian
+(therefore the fraction of the light in the bigger gaussian is $1-v$).
+In a coordinate system aligned with the major axis of the ellipse, the
+surface brightness profile of a ``galaxy'' is given by
+\begin{eqnarray}
+\rho(x, y) &=&
+\frac{f(1-v)}{2\pi a^2}\exp\left[-\frac{1}{2a^2}\left(qx^2 + y^2/q\right)\right]
++\frac{fv}{2\pi (ua)^2}\exp\left[-\frac{1}{2(ua)^2}\left(qx^2 + y^2/q\right)\right]
+\end{eqnarray}
 

### 2014-08-01 Distributions/Distribution.cpp
Virtual destructor

@@ -7,6 +7,11 @@ Distribution::Distribution()
 
 }
 
+Distribution::~Distribution()
+{
+
+}
+
 double Distribution::perturb1(const vector< vector<double> >& components,
 				vector< vector<double> >& u_components)
 {

### 2014-08-01 Distributions/Distribution.h

@@ -26,6 +26,8 @@ class Distribution
 	public:
 		Distribution();
 
+		virtual ~Distribution();
+
 		// Generate parameters from the prior
 		virtual void fromPrior() = 0;
 

### 2014-08-03 Documents/Paper/ms.tex
a --> w
URL, caching

@@
 and $\theta$ its orientation angle with respect to horizontal. The radius of the
-bigger gaussian is a parameter $a$.
+bigger gaussian is a parameter $w$.
 Finally
@@
 \rho(x, y) &=&
-\frac{f(1-v)}{2\pi a^2}\exp\left[-\frac{1}{2a^2}\left(qx^2 + y^2/q\right)\right]
-+\frac{fv}{2\pi (ua)^2}\exp\left[-\frac{1}{2(ua)^2}\left(qx^2 + y^2/q\right)\right]
+\frac{f(1-v)}{2\pi w^2}\exp\left[-\frac{1}{2w^2}\left(qx^2 + y^2/q\right)\right]\\
+&&+\frac{fv}{2\pi (uw)^2}\exp\left[-\frac{1}{2(uw)^2}\left(qx^2 + y^2/q\right)\right].
 \end{eqnarray}
@@
 components that have been removed and add in the effect of those that have been
-added.
+added. In the C++ implementation of this work, available at
+{\tt https://www.github.com/eggplantbren/RJObject/}, when a proposal takes
+place, the difference between the old and new model is cached and is available
+to the user, so that the likelihood can be updated rather than computed
+from scratch.
 

### 2014-08-04 Documents/Talk/ms.tex
Talk template

\documentclass{beamer}
\usetheme[pageofpages=of,% String used between the current page and the
                         % total page count.
          bullet=circle,% Use circles instead of squares for bullets.
          titleline=true,% Show a line below the frame title.
          alternativetitlepage=true,% Use the fancy title page.
       %   titlepagelogo=logo-polito,% Logo for the first page.
       %   watermark=watermark-polito,% Watermark used in every page.
       %   watermarkheight=100px,% Height of the watermark.
       %   watermarkheightmult=4,% The watermark image is 4 times bigger
                                % than watermarkheight.
          ]{Torino}

\author{Marco Barisione}
\title{Torino, a pretty theme for \LaTeX{} Beamer}
\institute{Politecnico di Torino}
\date{September 18, 2007}

\begin{document}
\include{content}
\end{document}

### 2014-08-04 Documents/Talk/content.tex
Started writing slides

% DO NOT COMPILE THIS FILE DIRECTLY!
% This is included by the other .tex files.

\begin{frame}[t,plain]
\titlepage
\end{frame}

\begin{frame}[t]{Context}
\begin{itemize}
\item Many problems have the following structure:
  \begin{itemize}
  \item There are $N$ objects in a region, and we don't know the value of $N$
  \item Each object has properties $\mathbf{x}_i$
  \item We have some data $\mathcal{D}$ which we want to use to infer both $N$
        and $\{\mathbf{x}_i\}$.
  \end{itemize}
\end{itemize}
\end{frame}

### 2014-08-05 Documents/Talk/content.tex
Sinewave data figure
Galaxy figure

@@
 \end{frame}
+
+\begin{frame}[t]{Motivating Problem I}
+\begin{itemize}
+\item How many sinusoids are in this data?
+\end{itemize}
+\begin{center}
+\includegraphics[scale=0.35]{sinewave_data.pdf}
+\end{center}
+\begin{itemize}
+\item Also, what are their periods, amplitudes and phases?
+\end{itemize}
+\end{frame}
+
+\begin{frame}[t]{Motivating Problem II}
+\begin{itemize}
+\item How many galaxies are in this data?
+\end{itemize}
+\begin{center}
+\includegraphics[scale=0.35]{../Paper/galaxyfield_data.pdf}
+\end{center}
+\begin{itemize}
+\item Also, what are their positions, sizes, orientations, etc?
+\end{itemize}
+\end{frame}
+
+\begin{frame}[t]{Bayesian Inference 101}
+Probabilities describe degrees of plausibility
+
+Uncertain quantities described by probability distributions
+
+\end{frame}

### 2014-08-06 Documents/Talk/content.tex
Bayes 101

@@
 \titlepage
-\end{frame}
-
-\begin{frame}[t]{Context}
-\begin{itemize}
-\item Many problems have the following structure:
-  \begin{itemize}
-  \item There are $N$ objects in a region, and we don't know the value of $N$
-  \item Each object has properties $\mathbf{x}_i$
-  \item We have some data $\mathcal{D}$ which we want to use to infer both $N$
-        and $\{\mathbf{x}_i\}$.
-  \end{itemize}
-\end{itemize}
 \end{frame}
@@
 
-\begin{frame}[t]{Bayesian Inference 101}
-Probabilities describe degrees of plausibility
+\begin{frame}[t]{Context}
+\begin{itemize}
+\item Many problems have the following structure:
+  \begin{itemize}
+  \item There are $N$ objects in a region, and we don't know the value of $N$
+  \item Each object has properties $\mathbf{x}_i$
+  \item We have some data $\mathcal{D}$ which we want to use to infer both $N$
+        and $\{\mathbf{x}_i\}$.
+  \end{itemize}
+\end{itemize}
+\end{frame}
 
-Uncertain quantities described by probability distributions
+\begin{frame}[t]{Bayesian Inference 101: I}
+Bayesian Inference is a unified framework for solving inference problems.
+We need the following ingredients:
+
+\begin{itemize}
+\item A ``hypothesis space'' describing the set of possible answers to our
+question (``parameter space'' in fitting is the same concept).
+\item A ``prior distribution'' $\pi\left(\theta\right)$ describing how plausible
+each of the possible solutions is, not taking into account the data.
+\end{itemize}
 
 \end{frame}
+
+\begin{frame}[t]{Bayesian Inference 101: II}
+
+The data helps us by changing our prior distribution into the ``posterior''
+distribution, given by
+\begin{eqnarray}
+\frac{\pi(\theta)\mathcal{L}(\theta)}{\mathcal{Z}}
+\end{eqnarray}
+\end{frame}

### 2014-08-09 Documents/Talk/content.tex
Slide placeholders
Various work on the introductory slides

--- 2014-08-05 Documents/Talk/content.tex
@@
 
+\begin{frame}[t]{Context}
+\begin{itemize}
+\item Many problems have the following structure:
+  \begin{itemize}
+  \item There are $N$ objects in a region, and we don't know the value of $N$
+  \item Each object has properties $\mathbf{x}_i$
+  \item We have some data $\mathcal{D}$ which we want to use to infer both $N$
+        and $\{\mathbf{x}_i\}$.
+  \end{itemize}
+\end{itemize}
+\end{frame}
+
 \begin{frame}[t]{Bayesian Inference 101}
-Probabilities describe degrees of plausibility
+Bayesian Inference is a unified framework for solving inference problems.
+We need the following ingredients:
 
-Uncertain quantities described by probability distributions
+\begin{itemize}
+\setlength{\itemsep}{20pt}
+\item A {\bf hypothesis space} describing the set of possible answers to our
+question (``parameter space'' in fitting is the same concept).
+\item A {\bf prior distribution} $p(\theta)$ describing how plausible
+each of the possible solutions is, not taking into account the data.
+\item A {\bf sampling distribution} $p(D | \theta)$ describing our knowledge
+about the connection between the parameters and the data.
+\end{itemize}
 
 \end{frame}
+
+\begin{frame}[t]{Bayesian Inference 101}
+The data helps us by changing our prior distribution to the {\bf posterior
+distribution}, given by
+\begin{eqnarray}
+p(\theta | D) &=& \frac{p(\theta) p(D|\theta)}{p(D)}
+\end{eqnarray}
+where the denominator is the normalisation constant, usually called either
+the {\bf marginal likelihood} or the {\bf evidence}.
+\begin{eqnarray}
+p(D) &=& \int p(\theta)p(D|\theta) \, d\theta.
+\end{eqnarray}
+
+\end{frame}
+
+\begin{frame}[t]{Bayesian Computation}
+To explore the posterior distribution, the Metropolis algorithm is a standard
+technique.
+
+\end{frame}
+
+\begin{frame}[t]{Birth and Death}
+For problems of unknown dimensionality, the hypothesis space is the union
+of several fixed-dimension hypothesis spaces.
+
+\end{frame}

### 2014-08-09 Documents/Talk/plots.py
Bayes plots

@@ -1,0 +1,23 @@
+from pylab import *
+
+rc("font", size=16, family="serif", serif="Computer Sans")
+rc("text", usetex=True)
+
+x = linspace(0., 100., 1001)
+
+prior = 1./(x + 10.)**2
+prior = prior/trapz(prior, x=x)
+lik = 0.08*exp(-0.5*(x - 60)**2/10**2)
+post = prior*lik
+post = post/trapz(post, x=x)
+
+plot(x, prior, 'b', linewidth=3, label='Prior')
+plot(x, lik, 'c--', linewidth=2, label='Likelihood')
+plot(x, post, 'r', linewidth=3, label='Posterior')
+xlabel(r'$\theta$')
+ylabel('Probability Density')
+legend(loc='upper right')
+savefig('bayes.pdf', bbox_inches='tight')
+show()
+
+

### 2014-08-09 Documents/Talk/plots.py
Show 'best fit'

@@ -14,9 +14,10 @@ post = post/trapz(post, x=x)
 plot(x, prior, 'b', linewidth=3, label='Prior')
 plot(x, lik, 'c--', linewidth=2, label='Likelihood')
 plot(x, post, 'r', linewidth=3, label='Posterior')
+plot(60, 0.08, 'c*', markersize=20, label='``Best fit\'\'')
 xlabel(r'$\theta$')
 ylabel('Probability Density')
-legend(loc='upper right')
+legend(loc='upper right', numpoints=1)
 savefig('bayes.pdf', bbox_inches='tight')
 show()
 

### 2014-08-09 Documents/Talk/content.tex
Include bayes figure
Tidying the introductory slides I have so far

@@
 \item How many sinusoids are in this data?
+\item Also, what are their periods, amplitudes and phases?
 \end{itemize}
@@
 \end{center}
-\begin{itemize}
-\item Also, what are their periods, amplitudes and phases?
-\end{itemize}
 \end{frame}
@@
 \item How many galaxies are in this data?
+\item Also, what are their positions, sizes, orientations, etc?
 \end{itemize}
@@
 \end{center}
-\begin{itemize}
-\item Also, what are their positions, sizes, orientations, etc?
-\end{itemize}
 \end{frame}
 
-\begin{frame}[t]{Context}
+\begin{frame}[t]{Similarities}
 \begin{itemize}
-\item Many problems have the following structure:
+\item Many problems (such as the previous two) have this structure:
+\vspace{20pt}
   \begin{itemize}
+  \setlength{\itemsep}{20pt}
   \item There are $N$ objects in a region, and we don't know the value of $N$
@@
   \item We have some data $\mathcal{D}$ which we want to use to infer both $N$
-        and $\{\mathbf{x}_i\}$.
+        and $\{\mathbf{x}_i\}_{i=1}^N$.
   \end{itemize}
@@
 \item A {\bf sampling distribution} $p(D | \theta)$ describing our knowledge
-about the connection between the parameters and the data.
+about the connection between the parameters and the data. When $D$ is known
+this is a function of $\theta$ called the {\bf likelihood}.
 \end{itemize}
@@
 
+\begin{frame}[t]{Posterior Distribution vs. Maximum Likelihood}
+The practical difference between these two concepts is greater in higher
+dimensional problems.
+\begin{center}
+\includegraphics[scale=0.4]{bayes.pdf}
+\end{center}
+\end{frame}
+
 \begin{frame}[t]{Bayesian Computation}

### 2014-08-09 Examples/GalaxyField/postprocess.py
Resample more

@@ -205,7 +205,7 @@ def postprocess(temperature=1., numResampleLogX=1, plot=True, loaded=[], \
 	print("Effective sample size = " + str(ESS))
 
 	# Resample to uniform weight
-	N = int(ESS)
+	N = 300#int(ESS)
 	posterior_sample = np.zeros((N, sample.shape[1]))
 	w = P_samples
 	w = w/np.max(w)

### 2014-08-09 Examples/GalaxyField/make_plots.py
Number of galaxies plot

@@ -25,3 +25,12 @@ title('GalaxyField Data')
 savefig('galaxyfield_data.pdf', bbox_inches='tight')
 show()
 
+posterior_sample = loadtxt('posterior_sample.txt')
+
+hist(posterior_sample[:,40010], 50)
+xlabel('Number of Galaxies $N$', fontsize=14)
+ylabel('Number of posterior samples', fontsize=14)
+xlim([-0.5, 100.5])
+savefig('N_galaxy_result.pdf', bbox_inches='tight')
+show()
+

### 2014-08-10 Documents/Talk/content.tex
Metropolis

@@
 To explore the posterior distribution, the Metropolis algorithm is a standard
-technique.
+technique:
+
+\begin{itemize}
+\item Start at some point $\theta$ in the hypothesis space.
+\item Loop\\
+$\{$
+  \begin{itemize}
+  \item Generate {\bf proposal} from some distribution $q(\theta' | \theta)$
+  (e.g. slightly perturb the current position).
+  \item With probability$^*$ $\alpha = \min\left(1, \frac{p(\theta')p(D|\theta')}{p(\theta)p(D|\theta)}\right)$, accept the proposal (i.e. replace $\theta$ with $\theta'$).
+  \item Otherwise, stay in the same place.
+  \end{itemize}
+$\}$
+\end{itemize}
 

### 2014-08-10 Documents/Talk/content.tex
Marginalisation figure

@@
 
-\begin{frame}[t]{Bayesian Computation}
+\begin{frame}[t]{Computation: Markov Chain Monte Carlo}
 To explore the posterior distribution, the Metropolis algorithm is a standard
@@
 \end{itemize}
+\end{frame}
 
+\begin{frame}[t]{What sampling gives us}
+\begin{columns}[T]
+\begin{column}{0.4\textwidth}
+  \begin{itemize}
+  \setlength{\itemsep}{20pt}
+  \item {\bf Marginalisation} becomes trivial
+  \item We can quantify all uncertainties we might be interested in
+  \end{itemize}
+\end{column}
+\hfill
+\begin{column}{0.55\textwidth}
+  \includegraphics[scale=0.25]{marginalisation.pdf}
+\end{column}
+
+\end{columns}
 \end{frame}

### 2014-08-11 Documents/Talk/plots.py
Plot showing challenges

@@ -21,4 +21,22 @@ legend(loc='upper right', numpoints=1)
 savefig('bayes.pdf', bbox_inches='tight')
 show()
 
+x = linspace(-1, 1, 201)
+[x, y] = meshgrid(x, x)
+y = y[::-1, :]
+
+rsq1 = (x + 0.2)**2 + (y - 0.5)**2
+rsq2 = (x - 0.4)**2 + (y - 0.7)**2
+rsq3 = (y + 1 - x**2)**2
+f1 = exp(-0.5*rsq1/0.05**2)
+f2 = exp(-0.5*rsq2/0.03**2)
+f3 = exp(-0.5*rsq3/0.02**2)
+f4 = exp(-0.5*rsq2/0.3**2)
+imshow(-f1 - f2 - f3 - f4, cmap='gray')
+xlabel('$x$')
+ylabel('$y$')
+gca().set_xticks([])
+gca().set_yticks([])
+savefig('challenges.pdf', bbox_inches='tight')
+show()
 

### 2014-08-11 Documents/Talk/content.tex
Tinkering with layout
Birth death sentence
Dimensions picture
Challenges
Moved slide
Challenges figure slide
NS slide

@@
 \end{center}
-\end{frame}
-
-\begin{frame}[t]{Similarities}
-\begin{itemize}
-\item Many problems (such as the previous two) have this structure:
-\vspace{20pt}
-  \begin{itemize}
-  \setlength{\itemsep}{20pt}
-  \item There are $N$ objects in a region, and we don't know the value of $N$
-  \item Each object has properties $\mathbf{x}_i$
-  \item We have some data $\mathcal{D}$ which we want to use to infer both $N$
-        and $\{\mathbf{x}_i\}_{i=1}^N$.
-  \end{itemize}
-\end{itemize}
 \end{frame}
@@
 \begin{columns}[T]
-\begin{column}{0.4\textwidth}
+\begin{column}{0.35\textwidth}
+  \vspace{30pt}
   \begin{itemize}
@@
 \hfill
-\begin{column}{0.55\textwidth}
+\begin{column}{0.5\textwidth}
+  \hspace{-30pt}
   \includegraphics[scale=0.25]{marginalisation.pdf}
@@
 
+\begin{frame}[t]{Structure of the two example problems}
+  \begin{itemize}
+  \setlength{\itemsep}{10pt}
+  \item There are $N$ objects in a region, and we don't know the value of $N$
+  \item Each object has properties $\mathbf{x}_i$
+  \item We have some data $\mathcal{D}$ which we want to use to infer both $N$
+        and $\{\mathbf{x}_i\}_{i=1}^N$.
+  \end{itemize}
+\begin{center}
+\includegraphics[scale=0.5]{../Paper/pgm.pdf}
+\end{center}
+\end{frame}
+
 \begin{frame}[t]{Birth and Death}
 For problems of unknown dimensionality, the hypothesis space is the union
-of several fixed-dimension hypothesis spaces.
+of several fixed-dimension hypothesis spaces. Can add {\bf birth and death}
+proposals that try to increase or decrease the number of objects in the model.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.7]{drawing.pdf}
+\end{center}
+\end{figure}
+\end{frame}
+
+\begin{frame}[t]{What's the issue?}
+\begin{itemize}
+\setlength{\itemsep}{20pt}
+\item Birth-death MCMC seems to have solved our problem. Or has it?
+\item What about these challenges?
+\vspace{20pt}
+  \begin{itemize}
+  \setlength{\itemsep}{20pt}
+  \item Multiple modes
+  \item Strong degeneracies
+  \item Phase transitions
+  \end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}[t]{Challenging features}
+\begin{center}
+\includegraphics[scale=0.4]{challenges.pdf}
+\end{center}
+\end{frame}
+
+\begin{frame}[t]{Nested Sampling}
+In 2004, John Skilling introduced {\bf Nested Sampling}, a technique
+that starts with the prior distribution, and shrinks volume at a controlled
+rate. Maps big space into easy 1-D problem.
+
+Resolves phase transition problem, and helps a little with multiple modes \&
+strong degeneracies. Figure from Skilling (2006).
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.2]{ns.png}
+\end{center}
+\end{figure}
 

### 2014-08-11 Documents/Talk/plots.py
DNest figure

@@ -40,3 +40,13 @@ gca().set_yticks([])
 savefig('challenges.pdf', bbox_inches='tight')
 show()
 
+sample = loadtxt('sample.txt')
+plot(sample[:,0], sample[:,1], 'k.', markersize=1)
+axis('scaled')
+xlabel('$x$')
+ylabel('$y$')
+gca().set_xticks([])
+gca().set_yticks([])
+savefig('dnest.png', bbox_inches='tight')
+show()
+

### 2014-08-11 Documents/Talk/content.tex
DNest figure
Include dnest figure
Software

@@
 \end{frame}
+
+\begin{frame}[t]{Diffusive Nested Sampling}
+\begin{columns}[T]
+\begin{column}{0.35\textwidth}
+  \vspace{30pt}
+  \begin{itemize}
+  \setlength{\itemsep}{20pt}
+  \item Replace the posterior distribution with a {\bf mixture of constrained priors}
+  \item Can cross valleys
+  \end{itemize}
+\end{column}
+\hfill
+\begin{column}{0.5\textwidth}
+  \hspace{50pt}
+  \includegraphics[scale=0.4]{dnest.png}
+\end{column}
+\end{columns}
+\end{frame}
+
+\begin{frame}[t]{Diffusive Nested Sampling}
+\begin{itemize}
+\setlength{\itemsep}{20pt}
+\item I've implemented the sinewave problem and the galaxy problem
+\item The common features (what proposals to use?) are delegated to
+a separate library called {\tt RJObject}
+\item Uses birth-death MCMC, but exploring the DNS distribution, rather than
+the (harder) posterior.
+\end{itemize}
+\end{frame}
+
+\begin{frame}[t]{Software}
+\begin{itemize}
+\setlength{\itemsep}{20pt}
+\item Diffusive Nested Sampling in C++
+{\tt github.com/eggplantbren/DNest3}
+\item Birth-death models in C++
+{\tt github.com/eggplantbren/RJObject}
+\item These are free software (GPLd).
+\end{itemize}
+\end{frame}

### 2014-08-11 Documents/Talk/content.tex
More slides

@@
   \item Can cross valleys
+  \item Posterior is available via re-weighting
   \end{itemize}
@@
 
+\begin{frame}[t]{Motivating Problem I}
+\begin{itemize}
+\item How many sinusoids are in this data?
+\item Also, what are their periods, amplitudes and phases?
+\end{itemize}
+\begin{center}
+\includegraphics[scale=0.35]{sinewave_data.pdf}
+\end{center}
+\end{frame}
+
+\begin{frame}[t]{Results from the sinusoid data}
+The $N=1$ and $N=2$ solutions are separated by a phase transition.
+\begin{center}
+\includegraphics[scale=0.45]{../Paper/N_result.pdf}
+\end{center}
+\end{frame}
+
+\begin{frame}[t]{Sinusoid data movie}
+
+\end{frame}
+
+\begin{frame}[t]{Results from the galaxy data}
+\begin{center}
+\includegraphics[scale=0.45]{../Paper/N_galaxy_result.pdf}
+\end{center}
+\end{frame}
+
+\begin{frame}[t]{Galaxy data movie}
+{\tt https://www.youtube.com/watch?v=s1udKctW2HY}
+\end{frame}
+
 \begin{frame}[t]{Software}
@@
 \item These are free software (GPLd).
+\item The learning curve is steep (sorry) but I'm open to collaborations.
 \end{itemize}

### 2014-08-17 Documents/Paper/ms.tex
GalaxyField N figure
The two phases
Paragraph

@@
 
-In other studies, it has been common to do separate runs of Nested Sampling,
-with $N$ fixed at various trial values.
+In other studies, it has been common to do separate runs of Nested Sampling
+(or another method),
+with $N$ fixed at various trial values \citep[e.g.][]{fengji}.
 Then, the posterior for $N$ can be calculated
@@
 
+Since DNS is effectively the Metropolis-Hastings algorithm applied to a
+distribution other than the posterior, proposal distributions are needed.
+In Section~\ref{sec:proposals}, we outline a set of proposal distributions used
+to explore the hypothesis space of possible values for $N$, $\hyperparams$,
+and $\{\xx_i\}$.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section{Metropolis proposals for the general problem}
+\section{Metropolis proposals for the general problem}\label{sec:proposals}
 We will now define a set of proposal distributions that can be used to
@@
 The shape of the likelihood function with respect to prior mass for the
-GalaxyField data.
-{\bf Bottom panel: }The posterior weights of the saved samples. Note the phase
-transition at $\log(X) \approx -100$ nats. While this doesn't affect the
+GalaxyField data. Note the phase
+transition at $\log(X) \approx -100$ nats. This separates models with just
+the few brightest galaxies from models with many faint galaxies as well.
+While this doesn't affect the
 posterior distribution (unlike the sinewave example), it would affect calculation
 of the marginal likelihood if annealing were used.
+{\bf Bottom panel: }The posterior weights of the saved samples.
 \label{fig:sinewaves_likelihood}}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{N_galaxy_result.pdf}
+\caption{The inference for $N$, the number of galaxies, based on the
+GalaxyField data. The true value was $N=47$.
+\label{fig:N_galaxy_result}}
 \end{center}

### 2014-08-27 Documents/Paper/ms.tex
Going through paper, added Stephens reference

@@
 \setlength{\parskip}{3mm}
-Many inference problems have the following structure. There is some unknown
-number, $N$, of objects in a region. Each of the $N$ objects has a property
+Consider the following class of inference problems. There is some unknown
+number, $N$, of objects in a (physical or metaphorical) region. Each of the $N$ objects has a property
 $\xx$, which may be a single scalar value (for example, a mass), or a set of
 values (e.g. a two-dimensional position and a mass). These problems are often
-challenging due to the unknown (and potentially large) dimensionality.
+challenging to solve due to the unknown (and potentially large) dimensionality.
 
-To do the inference, we must assign a prior to the objects' properties
+To carry out the inference, we must assign a prior to the objects' properties
 $\{\xx_i\}_{i=1}^N$.
@@
 Given data $\mathcal{D}$, we usually want to calculate properties of the
-posterior distribution given by:
+posterior distribution, given by:
 \begin{eqnarray}
@@
 \end{eqnarray}
-The general method used to calculate properties of the posterior distribution
-is to generate samples from it using Markov Chain Monte Carlo (MCMC). The
-structure of the model is depicted graphically in Figure~\ref{fig:pgm}.
+The structure of the model is depicted graphically in Figure~\ref{fig:pgm}.
+
+The computational method used to calculate properties of the posterior distribution
+is to generate samples from it using Markov Chain Monte Carlo (MCMC). Since
+$N$ is unknown, an MCMC sampler needs to be able to jump between candidate
+solutions with different numbers of objects, using either birth-death MCMC
+\citep{birthdeath} or the reversible jump framework of \citet{rjmcmc}.
 
@@
 
+\bibitem[Stephens(2000)]{birthdeath} Stephens, M., 2000, ``Bayesian analysis of mixture models with an unknown number of components-an alternative to reversible jump methods.'', Annals of Statistics (2000), 40-74.
+
 \bibitem[Umst{\"a}tter et al.(2005)]{2005PhRvD..72b2001U} Umst{\"a}tter, 

### 2014-08-27 Documents/Paper/ms.tex
More references and writing

@@
 
-In other studies, it has been common to do separate runs of Nested Sampling
+Some authors have addressed these kinds of problems by
+doing separate runs of Nested Sampling
 (or another method),
-with $N$ fixed at various trial values \citep[e.g.][]{fengji}.
+with $N$ fixed at various trial values \citep[e.g.][]{fengji, feroz}.
 Then, the posterior for $N$ can be calculated
@@
 distribution composed of a mixture of the prior distribution with other, more
-constrained distributions, facilitating mixing between the multiple modes
+constrained versions of the prior, facilitating mixing between the multiple modes
 and along degeneracy curves.
-This is similar to the idea of simulated annealing,
+This is similar to the idea of annealed importance sampling
+\citep{neal} or parallel tempering \citep{pt},
 where the target distribution is modified from the posterior to something
-``easier'', however,
+``easier''. However,
 the sequence of distributions used in Nested Sampling avoids some of the
-problems with simulated annealing: in particular, there is no need to choose
-a temperature schedule, and the method does not fail when phase transitions
+problems with annealing-based methods. In particular, there is no need to choose
+a temperature schedule (a potentially large set of tuning parameters),
+and the method does not fail when phase transitions
 are present.
@@
 Since DNS is effectively the Metropolis-Hastings algorithm applied to a
-distribution other than the posterior, proposal distributions are needed.
+distribution other than the posterior, proposal distributions are needed
+which define how a walker moves around the hypothesis space.
 In Section~\ref{sec:proposals}, we outline a set of proposal distributions used
@@
 {\bf Bottom panel: }The posterior weights of the saved samples.
-\label{fig:sinewaves_likelihood}}
+\label{fig:galaxyfield_likelihood}}
 \end{center}
@@
 
+\bibitem[Feroz and Hobson(2014)]{feroz} Feroz, F., Hobson, 
+M.~P.\ 2014.\ Bayesian analysis of radial velocity data of GJ667C with 
+correlated noise: evidence for only two planets.\ Monthly Notices of the 
+Royal Astronomical Society 437, 3540-3549.
+
 \bibitem[\protect\citeauthoryear{Green}{1995}]{rjmcmc}
@@
 Astronomical Society 410, 94-110.
+
+\bibitem[Hansmann(1997)]{pt} Hansmann, Ulrich HE., 1997, ``Parallel tempering algorithm for conformational studies of biological molecules.'', Chemical Physics Letters 281, no. 1 (1997): 140-150.
 

### 2014-08-27 Documents/Paper/ms.tex
Talking about heavy tailed proposals

@@
 \begin{eqnarray}
-u' &=& \textnormal{mod}\left(u + 10^{1.5 - 6a}b, 1\right)
+u' &:=& \textnormal{mod}\left(u + 10^{1.5 - 6a}b, 1\right)\label{eqn:heavytailed}
 \end{eqnarray}
@@
 
-Most of the proposals involve changing one (or a subset) of the parameters
-and/or hyperparameters while keeping the others fixed.
+Most of the proposals discussed in this paper involve changing one
+(or a subset) of the parameters and/or hyperparameters, while keeping the
+others fixed. Whenever a heavy-tailed proposal distribution
+is required, we use the proposal from Equation~\ref{eqn:heavytailed},
+or an analogous one, e.g. a discrete version when the proposal is to change $N$.
 

### 2014-08-27 Documents/Paper/ms.tex
Editing in N proposal section

@@
 where $\delta_N$ is drawn from a heavy tailed
-distribution. The most probable values of $\delta_N$ are $\pm 1$, but values
+distribution which is a discrete analogue of Equation~\ref{eqn:heavytailed}.
+The most probable values of $\delta_N$ are $\pm 1$, but values
 of order $N_{\rm max}$ also have some probability, to allow fast exploration
-when the target distribution is similar to the prior.
+when the target distribution is similar to the prior. If $N' = N$, $N'$ is
+regenerated as $N+1$ with probability 0.5, and $N-1$ with probability 0.5,
+and then wrapped back into the range $\{0,...,N_{\rm max}\}$ if necessary.
 
-If $N' > N$, the extra parameters needed, $\{\xx_i\}_{i=N+1}^{N'}$,
+If $N' > N$, i.e. the proposal is to add objects to the model,
+the extra parameters needed, $\{\xx_i\}_{i=N+1}^{N'}$,
 are drawn from their prior conditional on the current value of the
@@
 
-If $N' < N$, then $N - N'$ objects must be removed from the model. All
+If $N' < N$, i.e. the proposal is to remove objects from the model,
+then $N - N'$ objects must be selected for removal. All
 of the objects have the same probability of being selected for removal.

### 2014-08-27 Documents/Paper/ms.tex
More editing

@@
 hyperparameters, i.e. the interim prior $p(\xx | \hyperparams)$.
-
 If $N' < N$, i.e. the proposal is to remove objects from the model,
@@
 \end{eqnarray}
-where $\delta_u$ is drawn from a heavy-tailed distribution.
+where $\delta_u$ is drawn from a heavy-tailed distribution as in
+Equation~\ref{eqn:heavytailed}.
 
@@
 to a new value $\hyperparams'$. The proposal for the hyperparameter(s) is chosen
-to satisfy detailed balance with respect to $p(\hyperparams)$.
+to satisfy detailed balance with respect to $p(\hyperparams)$ and should be
+heavy-tailed.
 
@@
 affect the value of the likelihood, and therefore the likelihood will not need
-to be recomputed.
+to be recomputed. Therefore this proposal is most useful for mixing the values
+of the hyperparameters $\hyperparams$ when the target distribution is highly
+constrained.
 
@@
 Therefore, we allow an additional move that changes $\hyperparams$ to a new
-value $\hyperparams'$, but rather than leaving the objects $\{\xx_i\}$ fixed, we
-drag them so they represent the distribution $p(\xx | \hyperparams')$ rather than
+value $\hyperparams'$, but rather than leaving the objects $\{\xx_i\}$ fixed,
+they are ``dragged''
+so they represent the distribution $p(\xx | \hyperparams')$ rather than
 $p(\xx|\hyperparams)$. We do this by making use of the transformation functions
-$F(x; \hyperparams)$ and $G(x; \hyperparams)$ defined in Section~\ref{sec:proposal2}.
+$F(x; \hyperparams)$ and $G(x; \hyperparams)$ defined in
+Section~\ref{sec:proposal2}.
 
@@
 \end{eqnarray}
+In other words, the objects' parameters are transformed to $[0,1]$ using the
+current value of the hyperparameters, and then transformed back using the
+proposed value of the hyperparameters so they represent the new interim prior
+rather than the old one.
 

### 2014-08-27 Documents/Paper/ms.tex
More editing

@@
 Then, the posterior for $N$ can be calculated
-based on the estimates of the evidence obtained as a function of $N$. However,
-in this paper, we use trans-dimensional MCMC moves to infer $N$. In other words,
-$N$ is treated as just another parameter, and the marginal likelihood defined
-in Equation~\ref{eq:evidence}, which we can compute, also includes the sum over
-the unknown parameter $N$. Our motivation
-for using Nested Sampling is not primarily to calculate $\mathcal{Z}$.
+based on the estimates of the evidence obtained as a function of $N$.
+This approach does not generalise well to large $N$, as it would
+require a large number of parallel runs.
+
+In this paper, we introduce an approach that uses trans-dimensional MCMC moves
+to infer $N$, but embeds this process in the Nested Sampling framework to
+overcome difficulties that would be encountered by sampling the posterior
+distribution directly. Our motivation
+for using Nested Sampling is not primarily to calculate the marginal
+likelihood $\mathcal{Z}$ (including summing over $N$, as in
+Equation~\ref{eq:evidence}),
+although $\mathcal{Z}$ is readily available from the output.
 The motivation for using Nested Sampling, rather than just sampling the
@@
 and $\{\xx_i\}$.
+
+A C++ implementation of the ideas described in this paper is available online
+(under the terms of the GNU General Public Licence) at
+{\tt https://github.com/eggplantbren/RJObject}.
 

### 2014-08-27 Documents/Paper/ms.tex
Writing in the Sinewave section

@@
 
-\subsection{Prior Distributions}
-In the notation of Section~\ref{sec:introduction}, the parameters of the
-each object are
-\begin{eqnarray}
-\xx_i &=& \{A_i, T_i, \phi_i\}.
-\end{eqnarray}
-For the interim prior $p(\xx_i | \hyperparams)$, we introduced a single
-hyperparameter $\mu$, such that $A_i$ given $\mu$ has an exponential prior
-with mean $\mu$. For the periods, we assigned a log-uniform distribution for
-the periods between fixed boundaries, and a uniform distribution for the phases
-between 0 and $2\pi$. In other words, our prior is only hierarchical for the
-amplitudes $\{A_i\}$.
-
-The model for the shape of the (noise-free) signal is
-\begin{eqnarray}
-y(t) &=& \sum_{i=1}^N A_i \sin \left(\frac{2\pi t}{T_i} + \phi_i\right)
-\end{eqnarray}
-where there are $N$ sinusoids in the signal, the
-amplitudes are $\{A_i\}$, the periods are $\{T_i\}$, and the phases are
-$\{\phi_i\}$.
-The sampling distribution (probability distribution for the data given the
-parameters) is a normal (gaussian) distribution with mean zero and standard
-deviation $\sigma$, applied independently to each data point:
-\begin{eqnarray}
-Y_i | N, \{A_i\}, \{T_i\}, \{\phi_i\}, \sigma \sim
-\mathcal{N}\left(\mu(t_i), \sigma^2\right).
-\end{eqnarray}
-This also depends on an additional noise standard deviation parameter $\sigma$
-which will also be inferred. We used a log-uniform prior for $\sigma$ between
-$10^{-3}$ and $10^{3}$.
-
 \subsection{Simulated Dataset}
@@
 \end{eqnarray}
-and the noise standard deviation was $\sigma = 1$. The data $\{Y_i\}_{i=1}^n$
-is displayed in Figure~\ref{fig:sinewave_data}.
+and the noise standard deviation was $\sigma = 1$.
 
@@
 
+\subsection{Prior Distributions}
+To infer $N$, the number of sinusoids, as well as all the periods, amplitudes,
+and phases, we need to define prior distributions.
+In the notation of Section~\ref{sec:introduction}, the parameters of the
+each object are
+\begin{eqnarray}
+\xx_i &=& \{A_i, T_i, \phi_i\}.
+\end{eqnarray}
+For the interim prior $p(\xx_i | \hyperparams)$, we introduced a single
+hyperparameter $\mu$, such that $A_i$ given $\mu$ has an exponential prior
+with mean $\mu$. For the periods, we assigned a log-uniform distribution for
+the periods between fixed boundaries, and a uniform distribution for the phases
+between 0 and $2\pi$. In other words, our prior is only hierarchical for the
+amplitudes $\{A_i\}$. The prior for $\mu$ was a log-uniform prior, with
+probability density $\propto 1/\mu$, where $\mu \in [10^{-3}, 10^3]$.
+
+The model for the shape of the (noise-free) signal is
+\begin{eqnarray}
+y(t) &=& \sum_{i=1}^N A_i \sin \left(\frac{2\pi t}{T_i} + \phi_i\right)
+\end{eqnarray}
+where there are $N$ sinusoids in the signal, the
+amplitudes are $\{A_i\}$, the periods are $\{T_i\}$, and the phases are
+$\{\phi_i\}$.
+The sampling distribution (probability distribution for the data given the
+parameters) was a normal (gaussian) distribution with mean zero and standard
+deviation $\sigma$, applied independently to each data point:
+\begin{eqnarray}
+Y_i | N, \{A_i\}, \{T_i\}, \{\phi_i\}, \sigma \sim
+\mathcal{N}\left(\mu(t_i), \sigma^2\right).
+\end{eqnarray}
+This also depends on an additional noise standard deviation parameter $\sigma$
+which will also be inferred. We used a log-uniform prior for $\sigma$ between
+$10^{-3}$ and $10^{3}$.
+
 \subsection{Results}
 We ran DNS on the simulated data shown in Figure~\ref{fig:sinewave_data}.
-The shape of the likelihood function, with respect to the prior mass (the
-usual likelihood curve associated with Nested Sampling), us shown in
-Figure~\ref{fig:sinewaves_likelihood}. Two phase transitions can be seen. The
-first, at $\log(X) \approx -10$ nats,
+The shape of the likelihood function with respect to the prior mass (the
+usual likelihood curve associated with Nested Sampling) is shown in
+Figure~\ref{fig:sinewaves_likelihood}. Two phase transitions can be seen,
+where this curve is concave up. The
+first phase transition, at $\log(X) \approx -10$ nats,
 separates ``noise-only'' models (where the
@@
 where $N=1$. At $\log(X) \approx -28$ nats, a second phase transition separates
-$N=1$ models from $N=2$ models.
+$N=1$ models from $N=2$ models. This phase transition causes a double-peaked
+feature in the posterior weights (lower panel of
+Figure~\ref{fig:sinewaves_likelihood}).
 
@@
 
+Although DNS explores a distribution other than the posterior (a mixture of
+constrained priors is used instead), the saved samples can be assigned
+importance weights so we can represent the posterior distribution. The
+marginal posterior distribution for $N$ is shown in Figure~\ref{fig:N_result}.
+The most probable value is $N=2$ which is also the true value, and there is
+a small probability for $N=1$. The decreasing probabilities for $N > 2$ are
+due to the well-known natural ``Occam's Razor'' effect that can occur in
+Bayesian Inference \citep[e.g.][Chapter 28]{mackay}.
+
 \begin{figure}
@@
 \caption{The inference for $N$, the number of sinusoids, based on the
-sinusoid data. The true value was $N=2$.
+sinusoid data. The true value was $N=2$, which is also the most probable value
+given the data. Of course, this is not always the case.
 \label{fig:N_result}}
@@
 
+\bibitem[MacKay(2003)]{mackay} MacKay, David J.~C., 2003, ``Information theory, inference, and learning algorithms''. Vol. 7.
+Cambridge: Cambridge university press, 2003.
+
 \bibitem[Neal(2001)]{neal} Neal, R.~M., 2001, 

### 2014-08-28 Documents/Paper/ms.tex
Working on galaxyfield section

@@
 
-\section{Sinusoidal example}
+\section{Sinusoidal example}\label{sec:sinewaves}
 In this section we demonstrate a seemingly simple example which exhibits
@@
 how many objects $N$ (such as stars or galaxies) are in the image, and their
-properties (such as flux, size, orientation). Various ad-hoc approaches exist,
+properties (such as flux, size, orientation). Various approaches exist,
 covering a wide spectrum between ad-hoc approaches and principled inference
@@
 \citep[e.g.][]{irwin, sextractor, dolphot, 2003MNRAS.338..765H, starfield}.
+The more principled techniques tend to be much more computationally
+intensive, so should only be applied to interesting or
+especially challenging subsets of astronomical imaging.
 
@@
 extended objects such as galaxies in a noisy image.
-We model each galaxy as a mixture of two concentric elliptical components, with
+We model each galaxy as a mixture of two concentric elliptical gaussian
+components, with
 the following eight parameters. Firstly, two
@@
 \end{eqnarray}
+We generated a simulated $200 \times 200$ pixel noisy image
+(Figure~\ref{fig:galaxyfield_data}) containing 47 ``galaxies'',
+in order to test the algorithm.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.7]{galaxyfield_data.pdf}
+\caption{The simulated image for the GalaxyField example. The image is
+$200\times 200$ pixels in size and contains $N=47$ ``galaxies''.
+\label{fig:galaxyfield_data}}
+\end{center}
+\end{figure}
 
 For the interim prior for the total fluxes of the galaxies, and for the
-widths of the galaxies, we use the Pareto
+widths of the galaxies, we used the Pareto
 distribution. The probability density function for a variable $x$ given
@@
 Since we are using this for both the fluxes and the widths of the galaxies,
-there will be four hyperparameters in total for the fluxes and widths.
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.7]{galaxyfield_data.pdf}
-\caption{The simulated data for the GalaxyField example.
-\label{fig:galaxyfield_data}}
-\end{center}
-\end{figure}
+there will be four hyperparameters in total: two lower limits (one for the
+fluxes and one for the widths) and two slopes. As in Section~\ref{sec:sinewaves}
+we allowed the noise standard deviation $\sigma$ to be a free parameter with
+a log-uniform prior between $10^{-3}$ and $10^3$.
 
@@
 components that have been removed and add in the effect of those that have been
-added. In the C++ implementation of this work, available at
-{\tt https://www.github.com/eggplantbren/RJObject/}, when a proposal takes
+added. In the C++ implementation of this work, when a proposal takes
 place, the difference between the old and new model is cached and is available

### 2014-08-28 Documents/Paper/ms.tex
label switching

@@
 challenging to solve due to the unknown (and potentially large) dimensionality.
+In addition, they also have the so-called ``label-switching degeneracy'' issue
+\citep{label_switching}, where the meaning of the model is invariant to
+switching the identities of the objects.
 
@@
 
+\bibitem[Jasra et al(2005)]{label_switching} Jasra, A., Holmes, C.~C, and
+Stephens, D.~A., Markov Chain Monte Carlo Methods and the Label Switching
+Problem in Bayesian Mixture Modeling, Statistical Science
+Vol. 20, No. 1, pp. 50-67
+
 \bibitem[MacKay(2003)]{mackay} MacKay, David J.~C., 2003, ``Information theory, inference, and learning algorithms''. Vol. 7.

### 2014-08-28 Documents/Paper/ms.tex
Working on results

@@
 
+The marginal likelihood estimate returned by DNS for this data was
+$\log(\mathcal{Z}) = -1410.8$, and the information, or Kullback-Leibler
+divergence from the prior to the posterior, was estimated to be 29.7 nats.
+
 %\section{``Transit'' Example}
@@
 
+\subsection{Priors}
 For the interim prior for the total fluxes of the galaxies, and for the
@@
 there will be four hyperparameters in total: two lower limits (one for the
-fluxes and one for the widths) and two slopes. As in Section~\ref{sec:sinewaves}
+fluxes and one for the widths) and two slopes. The lower limits were assigned
+log-uniform priors between $10^{-3}$ and $10^3$, and uniform priors between 0
+and 1 were assigned to the {\it inverses} of the slopes. This latter choice
+was based on the fact that a Pareto distribution for $x$ is an exponential
+distribution for $\ln(x)$, with scale length $1/a$.
+
+The interim priors for the radius ratios $u$ and the flux ratios $w$ were both
+uniform distributions. Since $u \in [0,1]$ and $v \in [0,1]$, the upper limits
+of the uniform interim priors must be between 0 and 1. Therefore the prior
+for the upper limit was a uniform distribution between 0 and 1, and the prior
+for the lower limit given the upper limit was a uniform distribution between 0
+and the upper limit.
+
+As in Section~\ref{sec:sinewaves}
 we allowed the noise standard deviation $\sigma$ to be a free parameter with
 a log-uniform prior between $10^{-3}$ and $10^3$.
+
+\subsection{Results}
+The results from running DNS on the simulated galaxy data are shown in
+Figures~\ref{fig:galaxyfield_likelihood} and~\ref{fig:N_galaxy_result}.
+This problem, like the sinusoidal problem, exhibits a phase transition, which
+can be seen in Figure~\ref{fig:galaxyfield_likelihood}. The phase transition
+separates models which only fit the brightest galaxies from models that
+also fit the faint galaxies close to the noise level. This phase transition
+does not affect the posterior distribution, as the low-$N$ solutions are
+eventually found to have very low importance.
 
@@
 \end{figure}
+
+The marginal likelihood estimate returned by DNS for this data was
+$\log(\mathcal{Z}) = -319707.6$, and the information, or Kullback-Leibler
+divergence from the prior to the posterior, was estimated to be 550.2 nats.
 
@@
 to the user, so that the likelihood can be updated rather than computed
-from scratch.
+from scratch. Using this technique offers a speed advantage of a factor of
+$\sim 2$ for both of the example problems. The sinusoidal example could be
+solved in a few minutes on a desktop computer, whereas the galaxy example took
+much longer, about a day, to execute.
 

### 2014-08-28 Documents/Paper/ms.tex
Attempt at a conclusion

@@
 does not affect the posterior distribution, as the low-$N$ solutions are
-eventually found to have very low importance.
+eventually found to have very low importance. However it would cause difficulty
+for a thermal approach to calculating the marginal likelihood.
 
@@
 much longer, about a day, to execute.
+
+\section{Conclusions}
+In this paper, we presented a general approach to trans-dimensional Bayesian
+inference problems. To compute the posterior distribution in these contexts,
+birth and death MCMC moves are commonly used. However, the approach presented
+here replaces the posterior distribution with the alternative target
+distribution used in Diffusive Nested Sampling. This is done to facilitate
+mixing between multiple modes, along strong dependencies,
+and between different phases. Additionally,
+the marginal likelihood can be computed including the sum over the unknown
+number of components. The method was demonstrated on two illustrative examples
+inspired by astronomical data analysis, but should be applicable in other
+contexts such as mixture modelling. Software (in C++) implementing the
+general Metropolis proposals for these models, as well as the specific
+examples, is available at {\tt https://github.com/eggplantbren/RJObject}.
 

### 2014-08-28 Documents/Paper/ms.tex
Starfield is basically the same
Mention mixture models
Mention mixtures in abstract

@@
 objects in some region, along with their properties $\{\xx_i\}_{i=1}^N$,
-from a dataset $\mathcal{D}$. These problems are typically solved using one
-of the following two methods: i)
+from a dataset $\mathcal{D}$. A common statistical example is finite mixture
+modelling. In the Bayesian framework, these problems are typically solved using
+one of the following two methods: i)
 by executing a Monte Carlo algorithm (such as Nested Sampling) once for each
@@
 \citep{label_switching}, where the meaning of the model is invariant to
-switching the identities of the objects.
+switching the identities of the objects. Finite mixture modelling is a common
+statistical example.
 
@@
 intensive, so should only be applied to interesting or
-especially challenging subsets of astronomical imaging.
+especially challenging subsets of astronomical imaging. In fact, the approach
+used by \citet{starfield} was essentially an early version of the methodology
+presented in this paper.
 

### 2014-08-28 Documents/Paper/ms.tex
citations

@@
 problems with mixing. Diffusive Nested Sampling
-(DNS) replaces the posterior distribution with an alternative target
+\citep[DNS][]{dnest} replaces the posterior distribution with an alternative target
 distribution composed of a mixture of the prior distribution with other, more
@@
 models have been used in many different fields
-\citep[e.g.][]{2005PhRvD..72b2001U, 2007ApJ...654..551B, 2003AIPC..659....3B}.
+\citep[e.g.][]{2003AIPC..659....3B, 2005PhRvD..72b2001U, 2007ApJ...654..551B,
+2009MNRAS.395.2226B}.
 Note that this problem is also very closely related to the
@@
 
+\bibitem[Brewer and Stello(2009)]{2009MNRAS.395.2226B} Brewer, B.~J., 
+Stello, D.\ 2009.\ Gaussian process modelling of asteroseismic data.\ 
+Monthly Notices of the Royal Astronomical Society 395, 2226-2233.
+
 \bibitem[Brewer et al.(2013)]{starfield} Brewer, B.~J., 

### 2014-08-29 Documents/Paper/ms.tex
Extra sentence in abstract

@@
 We present two example problems, finding sinusoidal signals in
-noisy data, and finding and measuring galaxies in a noisy astronomical image.}
+noisy data, and finding and measuring galaxies in a noisy astronomical image.
+Both of the examples demonstrate phase transitions in the relationship between
+the likelihood and the cumulative prior mass.}
 

### 2014-12-01 Documents/Paper/ms.tex
affil
objects --> components

@@
 \author{Brendon J. Brewer}
+\date{Department of Statistics, The University of Auckland\\
+Private Bag 92019, Auckland 1142, New Zealand\\
+\vspace{0.3cm}
+{\tt bj.brewer@auckland.ac.nz}}
 
@@
 \abstract{Many inference problems involve inferring the number $N$ of
-objects in some region, along with their properties $\{\xx_i\}_{i=1}^N$,
+components in some region, along with their properties $\{\xx_i\}_{i=1}^N$,
 from a dataset $\mathcal{D}$. A common statistical example is finite mixture
@@
 Consider the following class of inference problems. There is some unknown
-number, $N$, of objects in a (physical or metaphorical) region. Each of the $N$ objects has a property
+number, $N$, of components in a (physical or metaphorical) region. Each of the $N$ components has a property
 $\xx$, which may be a single scalar value (for example, a mass), or a set of
@@
 \citep{label_switching}, where the meaning of the model is invariant to
-switching the identities of the objects. Finite mixture modelling is a common
+switching the identities of the components. Finite mixture modelling is a common
 statistical example.
 
-To carry out the inference, we must assign a prior to the objects' properties
+To carry out the inference, we must assign a prior to the components' properties
 $\{\xx_i\}_{i=1}^N$.
@@
 $N$ is unknown, an MCMC sampler needs to be able to jump between candidate
-solutions with different numbers of objects, using either birth-death MCMC
+solutions with different numbers of components, using either birth-death MCMC
 \citep{birthdeath} or the reversible jump framework of \citet{rjmcmc}.
@@
 of model discussed in this paper. Produced using daft ({\tt daft-pgm.org}).
-The number, $N$, of objects in the model is an unknown parameter, as are the
-properties of the objects. The data depends on the properties of the objects.
+The number, $N$, of components in the model is an unknown parameter, as are the
+properties of the components. The data depends on the properties of the components.
 Note that it is possible (and common) to have other parameters in the model
@@
 
-% Consider terminology: "objects" vs. "components"
+% Consider terminology: "components" vs. "components"
 
@@
 
-If $N' > N$, i.e. the proposal is to add objects to the model,
+If $N' > N$, i.e. the proposal is to add components to the model,
 the extra parameters needed, $\{\xx_i\}_{i=N+1}^{N'}$,
@@
 hyperparameters, i.e. the interim prior $p(\xx | \hyperparams)$.
-If $N' < N$, i.e. the proposal is to remove objects from the model,
-then $N - N'$ objects must be selected for removal. All
-of the objects have the same probability of being selected for removal.
+If $N' < N$, i.e. the proposal is to remove components from the model,
+then $N - N'$ components must be selected for removal. All
+of the components have the same probability of being selected for removal.
 
-\subsection{Proposals that modify objects}\label{sec:proposal2}
+\subsection{Proposals that modify components}\label{sec:proposal2}
 We now consider a proposal distribution that modifies one or more of the
-objects $\{\xx_i\}$, while keeping the number of objects $N$, as well as the
+components $\{\xx_i\}$, while keeping the number of components $N$, as well as the
 hyperparameters $\hyperparams$, fixed.
@@
 proposals that change more than one object at a time are possible. Hence,
-we choose the number of objects to change from a heavy tailed distribution
+we choose the number of components to change from a heavy tailed distribution
 wherer the most probable value is 1 but there is also a nontrivial probability
-of proposing to change $\sim N$ objects at once.
+of proposing to change $\sim N$ components at once.
 
 \subsection{Proposals that change the hyperparameters,
-keeping the objects fixed}\label{sec:proposal3}
+keeping the components fixed}\label{sec:proposal3}
 Another kind of proposal that we will consider is a proposal that keeps all of
-the objects fixed in place (i.e. leaves $\{\xx_i\}$) unchanged, but changes
+the components fixed in place (i.e. leaves $\{\xx_i\}$) unchanged, but changes
 the hyperparameter(s) from their current value $\hyperparams$
@@
 \end{eqnarray}
-Since this proposal leaves the objects $\{\xx_i\}$ fixed, it will usually not
+Since this proposal leaves the components $\{\xx_i\}$ fixed, it will usually not
 affect the value of the likelihood, and therefore the likelihood will not need
@@
 \subsection{Proposals that change the hyperparameters
-and all of the objects}\label{sec:proposal4}
+and all of the components}\label{sec:proposal4}
 The above proposals allow for changes to $N$, $\hyperparams$, and $\{\xx_i\}$,
@@
 Therefore, we allow an additional move that changes $\hyperparams$ to a new
-value $\hyperparams'$, but rather than leaving the objects $\{\xx_i\}$ fixed,
+value $\hyperparams'$, but rather than leaving the components $\{\xx_i\}$ fixed,
 they are ``dragged''
@@
 \end{eqnarray}
-In other words, the objects' parameters are transformed to $[0,1]$ using the
+In other words, the components' parameters are transformed to $[0,1]$ using the
 current value of the hyperparameters, and then transformed back using the
@@
 usually one or more noisy, blurred, images of the sky, we would like to know
-how many objects $N$ (such as stars or galaxies) are in the image, and their
+how many components $N$ (such as stars or galaxies) are in the image, and their
 properties (such as flux, size, orientation). Various approaches exist,
@@
 the DNS approach to a toy version of the problem of detecting and quantifying
-extended objects such as galaxies in a noisy image.
+extended components such as galaxies in a noisy image.
 We model each galaxy as a mixture of two concentric elliptical gaussian
@@
 change to the model, rather than computing the entire mock signal from scratch.
-For example, if the proposal is to add two new objects to the model, the
-mock signal can simply be updated by adding the effect of the two new objects
+For example, if the proposal is to add two new components to the model, the
+mock signal can simply be updated by adding the effect of the two new components
 to the model.

### 2014-12-01 Documents/Paper/ms.tex
Going through the introduction to add more background info

@@
 Both of the examples demonstrate phase transitions in the relationship between
-the likelihood and the cumulative prior mass.}
+the likelihood and the cumulative prior mass, highlighting the need for
+Nested Sampling.}
 
@@
 Consider the following class of inference problems. There is some unknown
-number, $N$, of components in a (physical or metaphorical) region. Each of the $N$ components has a property
+number, $N$, of components in a (physical or metaphorical) region. Each of the $N$ components has parameters
 $\xx$, which may be a single scalar value (for example, a mass), or a set of
@@
 
-To carry out the inference, we must assign a prior to the components' properties
+To carry out the inference, we must assign a prior to the components' parameters
 $\{\xx_i\}_{i=1}^N$.
-Since $N$ may be large, it is usually easier to assign an ``interim prior''
-conditional on some hyperparameters $\hyperparams$, and then assign a prior to
+Since $N$ may be large, it is usually easier to assign an ``conditional prior''
+given some hyperparameters $\hyperparams$, and then assign a prior to
 $\hyperparams$. This kind of model is usually called {\it hierarchical}.
@@
 Here we have assumed the priors for $N$ and $\hyperparams$ are independent, and
-the interim prior for $\{\xx_i\}$ is iid and does not depend on $N$.
+the conditional prior for $\{\xx_i\}$ is iid and does not depend on $N$.
 Given data $\mathcal{D}$, we usually want to calculate properties of the
@@
 difficult features, such as strong dependencies or multiple modes, that cause
-problems with mixing. Diffusive Nested Sampling
+problems with mixing.
+
+In addition to these well-known challenges,
+{\it phase transitions} \citep{skilling} can also cause difficulties for
+Bayesian computation. Phase transitions occur when the posterior distribution
+is a mixture of a ``slab'' with high volume but low density, and a ``spike''
+with low volume but high density. To understand why this causes problems,
+imagine sampling the posterior using the Metropolis algorithm. If the sampler
+is in the slab, it will only rarely propose a move into the spike, because the
+spike's volume is so small. If the sampler is in the spike, a proposal to move
+into the slab will have a very low acceptance probability because the density
+of the slab is so low relative to that of the spike.
+
+Diffusive Nested Sampling
 \citep[DNS][]{dnest} replaces the posterior distribution with an alternative target

### 2014-12-01 Documents/Paper/ms.tex
Challenges figure

@@
 
+\begin{figure}
+\begin{minipage}[lr]{\textwidth}
+\begin{center}
+\includegraphics[width=2.7in]{challenges.pdf}
+\includegraphics[width=2.7in]{dnest.png}
+\end{center}
+\end{minipage}
+\caption{{\bf Left}: An example of a two-dimensional posterior distribution
+with multiple modes, strong dependence, and a phase transition (in the top
+right mode). While this low dimensional example is not difficult to sample,
+an analogous problem in high dimensions would be. {\bf Right}: Diffusive
+Nested Sampling replaces the target posterior distribution by a mixture of
+constrained priors. The density is non-negligible everywhere (since the prior
+is a component of this distribution), and states intermediate between the two
+phases in the top-right mode are appropriately up-weighted.
+\label{fig:challenges}}
+\end{figure}
+
 Diffusive Nested Sampling
@@
 are present.
+Figure~\ref{fig:challenges} shows a mock-up of a two dimensional posterior
+density with three modes, one of which contains a strong dependence, and
+another of which contains a phase transition.
 

### 2014-12-04 Examples/SineWaves/generate_data.py
New dataset

@@ -1,24 +1,14 @@
 from pylab import *
 
-"""
-Original example with two phase transations, but they do not interfere
-with the posterior. This is here for no reason other than to make sure
-the second example is identical to what Courtney has been using.
-"""
 seed(123)
-
-t = linspace(0., 100., 1001)
-y = sin(2.*pi*t/30.) + sin(2.*pi*t/2. + 1.)
-sig = 1.
-Y = y + sig*randn(y.size)
 
 """
 This one has a phase transition at temperature = 1 :-D
 """
 
 t = linspace(0., 100., 1001)
-y = sin(2.*pi*t/30.) + 0.3*sin(2.*pi*t/2. + 1.)
-sig = 1.
+y = 7*sin(2.*pi*t/30.) + 0.155*sin(2.*pi*t/2. + 1.)
+sig = 0.5
 Y = y + sig*randn(y.size)
 
 data = empty((t.size, 2))
@@ -29,8 +19,8 @@ savetxt('fake_data.txt', data)
 rc("font", size=14, family="serif", serif="Computer Sans")
 rc("text", usetex=True)
 plot(t, y, 'r', linewidth=2, label='Noise-free signal')
-plot(t, Y, 'bo', alpha=0.4, label='Noisy measurements')
-ylim([-5., 5.])
+plot(t, Y, 'bo', markersize=3, alpha=0.5, label='Noisy measurements')
+ylim([-12., 13.])
 legend(numpoints=1)
 xlabel('Time')
 ylabel('Signal')

### 2014-12-05 Documents/Paper/ms.tex
New sinewave dataset

@@
 to detect. The true values of the parameters were
-$N=2$, $\mathbf{A} = \{1, 0.3\}$,
+$N=2$, $\mathbf{A} = \{7, 0.155\}$,
 $\mathbf{T}=\{30, 2\}$, and $\boldsymbol{\phi} = \{0, 1\}$. The signal was
@@
 \end{eqnarray}
-and the noise standard deviation was $\sigma = 1$.
+and the noise standard deviation was $\sigma = 0.5$.
 
@@
 the true signal and the points are the measurements, simulated based on a
-noise standard deviation of $\sigma = 1$.
+noise standard deviation of $\sigma = 0.5$.
 \label{fig:sinewave_data}}
@@
 entire dataset is accounted for by the noise term) from ``one-sinusoid'' models
-where $N=1$. At $\log(X) \approx -28$ nats, a second phase transition separates
+where $N=1$. At $\log(X) \approx -35$ nats, a second phase transition separates
 $N=1$ models from $N=2$ models. This phase transition causes a double-peaked

### 2014-12-05 Examples/SineWaves/Metropolis/MyModel.cpp
Straight metropolis

@@ -1,0 +1,104 @@
+#include "MyModel.h"
+#include "RandomNumberGenerator.h"
+#include "Utils.h"
+#include "Data.h"
+#include <cmath>
+
+using namespace std;
+using namespace DNest3;
+
+MyModel::MyModel()
+:objects(3, 10, false, MyDistribution(-10., 10., 1E-3, 1E3))
+,mu(Data::get_instance().get_t().size())
+{
+
+}
+
+void MyModel::fromPrior()
+{
+	objects.fromPrior();
+	objects.consolidate_diff();
+	sigma = exp(log(1E-3) + log(1E6)*randomU());
+	calculate_mu();
+	calculate_logL();
+}
+
+void MyModel::calculate_mu()
+{
+	// Get the times from the data
+	const vector<double>& t = Data::get_instance().get_t();
+
+	// Update or from scratch?
+	bool update = (objects.get_added().size() < objects.get_components().size());
+
+	// Get the components
+	const vector< vector<double> >& components = (update)?(objects.get_added()):
+				(objects.get_components());
+
+	// Zero the signal
+	if(!update)
+		mu.assign(mu.size(), 0.);
+
+	double T, A, phi;
+	for(size_t j=0; j<components.size(); j++)
+	{
+		T = exp(components[j][0]);
+		A = components[j][1];
+		phi = components[j][2];
+		for(size_t i=0; i<t.size(); i++)
+			mu[i] += A*sin(2.*M_PI*t[i]/T + phi);
+	}
+}
+
+double MyModel::perturb()
+{
+	double logH = -logL;
+
+	if(randomU() <= 0.75)
+	{
+		logH += objects.perturb();
+		objects.consolidate_diff();
+		calculate_mu();
+	}
+	else
+	{
+		sigma = log(sigma);
+		sigma += log(1E6)*randh();
+		sigma = mod(sigma - log(1E-3), log(1E6)) + log(1E-3);
+		sigma = exp(sigma);
+	}
+	calculate_logL();
+	logH += logL;
+
+	return logH;
+}
+
+double MyModel::logLikelihood() const
+{
+	return logL;
+}
+
+void MyModel::calculate_logL()
+{
+	// Get the data
+	const vector<double>& y = Data::get_instance().get_y();
+
+	logL = 0.;
+	double var = sigma*sigma;
+	for(size_t i=0; i<y.size(); i++)
[…]

### 2014-12-05 Examples/SineWaves/Metropolis/MyModel.h

@@ -1,0 +1,41 @@
+#ifndef _MyModel_
+#define _MyModel_
+
+#include "Model.h"
+#include <vector>
+#include <RJObject.h>
+#include "MyDistribution.h"
+
+class MyModel:public DNest3::Model
+{
+	private:
+		RJObject<MyDistribution> objects;
+		double sigma; // Noise standard deviation
+
+		// The signal
+		std::vector<long double> mu;
+		void calculate_mu();
+		void calculate_logL();
+		double logL;
+
+	public:
+		MyModel();
+
+		// Generate the point from the prior
+		void fromPrior();
+
+		// Metropolis-Hastings proposals
+		double perturb();
+
+		// Likelihood function
+		double logLikelihood() const;
+
+		// Print to stream
+		void print(std::ostream& out) const;
+
+		// Return string with column information
+		std::string description() const;
+};
+
+#endif
+

### 2014-12-05 Examples/SineWaves/make_plots.py
New plot

@@ -12,3 +12,10 @@ xlim([-0.5, 10.5])
 savefig('N_result.pdf', bbox_inches='tight')
 show()
 
+sample_info = loadtxt('Metropolis/sample_info.txt')
+plot(sample_info[:,1], linewidth=1)
+xlabel('Iteration / 20000')
+ylabel('Log Likelihood')
+savefig('trace_logl.pdf', bbox_inches='tight')
+show()
+

### 2014-12-05 Documents/Paper/ms.tex
New plot
references

@@
 This kind of model has many applications, and has been solved analytically
-under certain sets of assumptions \citep[see e.g.][]{bretthorst}. Similar
+under certain sets of assumptions \citep[see e.g.][]{bretthorst, 2014arXiv1412.0467M}.
+Similar
 models have been used in many different fields
@@
 attracted a lot of research attention in recent years
-\citep[e.g.][]{2007ApJ...654..551B, gregory, fengji, 2011MNRAS.415.3462F}.
+\citep[e.g.][]{gregory, fengji, 2011MNRAS.415.3462F}.
 
@@
 divergence from the prior to the posterior, was estimated to be 29.7 nats.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{trace_logl.pdf}
+\caption{A ``trace plot'' of the log likelihood in the SineWave model, for an
+MCMC chain designed to sample the posterior distribution. The two phases
+have log-likelihoods around -730 and -745 respectively, however transitions
+between the two phases are quite uncommon.
+\label{fig:trace_logl}}
+\end{center}
+\end{figure}
 
@@
 
+\bibitem[Mortier et al.(2014)]{2014arXiv1412.0467M} Mortier, A., Faria, 
+J.~P., Correia, C.~M., Santerne, A., Santos, N.~C.\ 2014.\ BGLS: A Bayesian 
+formalism for the generalised Lomb-Scargle periodogram.\ ArXiv e-prints 
+arXiv:1412.0467
+
 \bibitem[Neal(2001)]{neal} Neal, R.~M., 2001, 

### 2014-12-05 Documents/Paper/ms.tex
evidence --> marginal likelihood

@@
 doing separate runs of Nested Sampling
-(or another method),
+(or another method of calculating marginal likelihoods),
 with $N$ fixed at various trial values \citep[e.g.][]{fengji, feroz}.
 Then, the posterior for $N$ can be calculated
-based on the estimates of the evidence obtained as a function of $N$.
+based on the estimates of the marginal likelihood obtained as a function of $N$.
 This approach does not generalise well to large $N$, as it would

### 2014-12-05 Documents/Paper/ms.tex
Fixed minor notation issues

@@
 a signal is composed of $N$ sinusoids, of different periods, amplitudes,
-and phases. The signal is observed at various times $\{t_i\}_{i=1}^N$ with
+and phases. The signal is observed at various times $\{t_i\}_{i=1}^m$ with
 noise, and we want to use the resulting data to infer the number of sinusoids
@@
 Y_i | N, \{A_i\}, \{T_i\}, \{\phi_i\}, \sigma \sim
-\mathcal{N}\left(\mu(t_i), \sigma^2\right).
+\mathcal{N}\left(y(t_i), \sigma^2\right).
 \end{eqnarray}

### 2014-12-05 Documents/Paper/ms.tex
Ewan

@@
 Courtney Donovan (Auckland), Tom Loredo (Cornell), Iain Murray (Edinburgh),
+Ewan Cameron (Oxford),
 John Skilling (MaxEnt Data Consultants), and Daniela Huppenkothen

### 2014-12-05 Documents/Paper/ms.tex
Logz and H

@@
 \includegraphics{pgm.pdf}
-\caption{\it A probabilistic graphical model (PGM) depicting the kind
+\caption{A probabilistic graphical model (PGM) depicting the kind
 of model discussed in this paper. Produced using daft ({\tt daft-pgm.org}).
@@
 The marginal likelihood estimate returned by DNS for this data was
-$\log(\mathcal{Z}) = -1410.8$, and the information, or Kullback-Leibler
-divergence from the prior to the posterior, was estimated to be 29.7 nats.
+$\log(\mathcal{Z}) = -771.8$, and the information, or Kullback-Leibler
+divergence from the prior to the posterior, was estimated to be 39.8 nats.
 

### 2014-12-06 Documents/Paper/ms.tex
Working on intro

@@
 \usepackage{amsmath}
+\usepackage{dsfont}
 
@@
 
-In addition to these well-known challenges,
+\subsection{Phase Transitions and Nested Sampling}
+In addition to the well-known challenges caused by multiple modes and strong
+dependencies,
 {\it phase transitions} \citep{skilling} can also cause difficulties for
@@
 into the slab will have a very low acceptance probability because the density
-of the slab is so low relative to that of the spike.
+of the slab is so low relative to that of the spike. If the spike has a volume
+$V_{\rm spike}$ and the slab has a volume $V_{\rm slab}$ then the number of
+MCMC iterations needed to jump from the spike to the slab, or vice versa, is
+of order $V_{\rm slab}/V_{\rm spike}$ which may be very large.
+
+In many problems, the posterior distribution is a mixture of a spike component
+and a slab component, however the total posterior probability in the slab
+component is negligible. In this situation, posterior sampling is unaffected,
+however problems can arise when calculating the marginal likelihood using
+an annealing-type approach. For example, consider the thermodynamic integral,
+which gives the log of the marginal likelihood:
+
+\begin{eqnarray}
+\log(\mathcal{Z}) &=& \int_0^1
+\left<\log\left[L(\theta)\right]\right>_\beta \,d\beta
+\end{eqnarray}
+
+where $L(\theta)$ is the likelihood function, and the expectation is taken
+with respect to the ``power posterior'' distributions which are intermediate
+between the prior $\pi(\theta)$ and the posterior:
+
+\begin{eqnarray}
+p_\beta(\theta) \propto \pi(\theta)L(\theta)^\beta.
+\end{eqnarray}
+
+If the posterior distribution is a mixture of a slab and a spike,
+then for some range of inverse temperatures $\beta$ the probability mass in
+the slab and the spike will be comparable. At these temperatures the MCMC
+method will be unable to mix and will give an incorrect estimate of
+$\left<\log\left[L(\theta)\right]\right>_\beta$.
+
+Nested Sampling \citep{skilling} overcomes these problems by working with an
+alternative sequence of probability distributions, proportional to the prior
+but with a hard constraint on the value of the likelihood:
+\begin{eqnarray}
+p_{l}(\theta) \propto \pi(\theta)\mathds{1}\left[L(\theta) > l\right]
+\end{eqnarray}
+The normalising constants of these distributions are given by
+\begin{eqnarray}
+X(l) &=& \int \pi(\theta)\mathds{1}\left[L(\theta) > l\right] \, d\theta
+\end{eqnarray}
+which is the amount of prior mass that has likelihood greater than
+$l$. As Nested Sampling proceeds, $X$ shrinks geometrically with time, so
+the number of iterations required to compress from the slab to the spike is
+of order $\log(V_{\rm slab}/V_{\rm spike})$.
+
+\subsection{Diffusive Nested Sampling}
+Diffusive Nested Sampling
+\citep[DNS][]{dnest} replaces the posterior distribution with an alternative target
+distribution composed of a mixture of the prior distribution with other, more
+constrained versions of the prior, facilitating mixing between the multiple modes
+and along degeneracy curves.
+This is similar to the idea of annealed importance sampling
+\citep{neal} or parallel tempering \citep{pt},
+where the target distribution is modified from the posterior to something
+``easier''. However,
+the sequence of distributions used in Nested Sampling avoids some of the
+problems with annealing-based methods. In particular, there is no need to choose
+a temperature schedule (a potentially large set of tuning parameters),
+and the method does not fail when phase transitions
+are present.
+Figure~\ref{fig:challenges} shows a mock-up of a two dimensional posterior
+density with three modes, one of which contains a strong dependence, and
+another of which contains a phase transition.
 
@@
 \end{figure}
-
-Diffusive Nested Sampling
-\citep[DNS][]{dnest} replaces the posterior distribution with an alternative target
-distribution composed of a mixture of the prior distribution with other, more
-constrained versions of the prior, facilitating mixing between the multiple modes
-and along degeneracy curves.
-This is similar to the idea of annealed importance sampling
-\citep{neal} or parallel tempering \citep{pt},
-where the target distribution is modified from the posterior to something
-``easier''. However,
-the sequence of distributions used in Nested Sampling avoids some of the
-problems with annealing-based methods. In particular, there is no need to choose
-a temperature schedule (a potentially large set of tuning parameters),
-and the method does not fail when phase transitions
-are present.
-Figure~\ref{fig:challenges} shows a mock-up of a two dimensional posterior
-density with three modes, one of which contains a strong dependence, and
-another of which contains a phase transition.
 

### 2014-12-06 Documents/Paper/ms.tex
More writing intro

@@
 where $L(\theta)$ is the likelihood function, and the expectation is taken
-with respect to the ``power posterior'' distributions which are intermediate
+with respect to the ``thermal'' distributions which are intermediate
 between the prior $\pi(\theta)$ and the posterior:
-
 \begin{eqnarray}
-p_\beta(\theta) \propto \pi(\theta)L(\theta)^\beta.
+p_\beta(\theta) \propto \pi(\theta)L(\theta)^\beta.\label{eq:annealing}
 \end{eqnarray}
-
 If the posterior distribution is a mixture of a slab and a spike,
@@
 \begin{eqnarray}
-p_{l}(\theta) \propto \pi(\theta)\mathds{1}\left[L(\theta) > l\right]
+p_{l}(\theta) \propto \pi(\theta)\mathds{1}\left[L(\theta) > l\right]\label{eq:constrained_prior}
 \end{eqnarray}
@@
 the number of iterations required to compress from the slab to the spike is
-of order $\log(V_{\rm slab}/V_{\rm spike})$.
+of order $\log(V_{\rm slab}/V_{\rm spike})$. During this process Nested Sampling
+visits states with likelihoods intermediate between those of the slab and the
+spike, whereas no choice of $\beta$ in Equation~\ref{eq:annealing} will give
+these states substantial probability.
 
 \subsection{Diffusive Nested Sampling}
+The main difficulty with Nested Sampling is coming up with methods to sample
+the constrained prior distributions of Equation~\ref{eq:constrained_prior}.
 Diffusive Nested Sampling
-\citep[DNS][]{dnest} replaces the posterior distribution with an alternative target
-distribution composed of a mixture of the prior distribution with other, more
-constrained versions of the prior, facilitating mixing between the multiple modes
+\citep[DNS][]{dnest} is an alternative version of Nested Sampling that is
+appropriate when MCMC is used to explore the parameter space.
+
+Rather than working with a sequence of constrained prior distributions like in
+Equation~\ref{eq:constrained_prior}, DNS
+replaces the posterior distribution with an alternative target
+distribution composed of a {\it mixture} of the prior distribution with the
+constrained priors, facilitating mixing between the multiple modes
 and along degeneracy curves.

### 2014-12-06 Documents/Paper/ms.tex
Object --> component

@@
 
-Let $F(x; \hyperparams)$ be a function that takes an object $x$ and transforms it
+Let $F(x; \hyperparams)$ be a function that takes an component $x$ and transforms it
 to a value $u$ that has a uniform distribution between 0 and 1, given $\hyperparams$.
-If the object $x$ consists of a single scalar value, $F$ is the cumulative
+If the component $x$ consists of a single scalar value, $F$ is the cumulative
 distribution of the interim prior. Denote the inverse of $F$ by $G$.
@@
 A proposal
-for an object involves transforming its parameters to $[0, 1]$ using $F$,
+for an component involves transforming its parameters to $[0, 1]$ using $F$,
 making the proposal in terms of $u$, and then transforming back.
@@
 
-Choosing just one object to change is most appropriate when the DNS
+Choosing just one component to change is most appropriate when the DNS
 distribution is very constrained. When it is close to the prior, bolder
-proposals that change more than one object at a time are possible. Hence,
+proposals that change more than one component at a time are possible. Hence,
 we choose the number of components to change from a heavy tailed distribution
@@
 The ``dragging'' process works as follows, and must be carried out on
-each object:
+each component:
 \begin{eqnarray}
@@
 In the notation of Section~\ref{sec:introduction}, the parameters of the
-each object are
+each component are
 \begin{eqnarray}

### 2014-12-06 Documents/Paper/ms.tex
Talking about the trace plot

@@
 slab. Note that this phase transition still exists if we condition on $N=2$,
-so is not caused by the trans-dimensional model.
+so is not caused by the trans-dimensional model. To demonstrate this,
+Figure~\ref{fig:trace_logl} shows the value of the log likelihood versus time
+for 45 million iterations of MCMC targeting the posterior distribution.
+The sampler should visit states with log likelihood $\sim -745$ about 10\%
+of the time (corresponding with models that don't contain the small oscillation),
+and the rest of the time should be spent in states with log likelihood around
+$-730$. In this run transitions between these two phases occurred only a
+handful of times. In a practical application, computing the posterior probability
+for the existence of the small oscillation might be the whole point of the
+calculation, so mixing between models where it is present and models where it
+is absent is crucial.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{trace_logl.pdf}
+\caption{A ``trace plot'' of the log likelihood in the SineWave model, for an
+MCMC chain designed to sample the posterior distribution. The two phases
+have log-likelihoods around -730 and -745 respectively, however transitions
+between the two phases are quite uncommon.
+\label{fig:trace_logl}}
+\end{center}
+\end{figure}
 

### 2014-12-06 Documents/Paper/ms.tex
Improving caption on sinewave data

@@
 \caption{The simulated data for the sinusoidal example. The solid line shows
-the true signal and the points are the measurements, simulated based on a
+the true signal which consists of a large slow oscillation and a much smaller
+amplitude fast oscillation. The points are the measurements, simulated with a
 noise standard deviation of $\sigma = 0.5$.
@@
 \includegraphics[scale=0.5]{trace_logl.pdf}
-\caption{A ``trace plot'' of the log likelihood in the SineWave model, for an
+\caption{A ``trace plot'' of the log likelihood in the sinusoidal model, for an
 MCMC chain designed to sample the posterior distribution. The two phases

### 2014-12-07 Documents/Paper/ms.tex
Sentence about :) figure
Fixed mistake picked up by referee
DNest target distribution

@@
 constrained priors, facilitating mixing between the multiple modes
-and along degeneracy curves.
+and along degeneracy curves. The mixture of constrained priors is given by
+a mixture of distributions like in Equation~\ref{eq:constrained_prior}:
+
+\begin{eqnarray}
+p(\theta) &=& \sum_{i=0}^n w_i
+\frac{\pi(\theta)\mathds{1}\left[L(\theta) > l_i\right]}{X_i}
+\end{eqnarray}
+where $\{l_i\}$ is a set of likelihood thresholds or ``levels'',
+constructed such that the associated prior mass $X_{i+1}$ of level
+$i+1$ is approximately $e^{-1}$ times that of level $i$. The $\{w_i\}$ are a
+set of mixture weights, usually chosen to emphasise the higher levels during
+the initial stages of a run (while creating levels),
+and set to uniform once the desired number of levels $n$ has been created.
+See \citet{dnest}
+for more details about DNS.
+
 This is similar to the idea of annealed importance sampling
@@
 are present.
-Figure~\ref{fig:challenges} shows a mock-up of a two dimensional posterior
+Figure~\ref{fig:challenges} shows a mock two dimensional posterior
 density with three modes, one of which contains a strong dependence, and
-another of which contains a phase transition.
+another of which contains a phase transition. The target distribution used by
+DNS includes the prior as part of the mixture, so travel between the modes is
+possible.
 
@@
 u_i &:=& F(\xx_i; \hyperparams)\\
-\xx_i' &:=& G(\xx_i; \hyperparams')
+\xx_i' &:=& G(u_i; \hyperparams')
 \end{eqnarray}

### 2014-12-07 Documents/Paper/ms.tex
Acceptance rate

@@
 appropriate when MCMC is used to explore the parameter space.
+A C++ implementation of DNS, called {\tt DNest3}, is available at
+{https://github.com/eggplantbren/DNest3}.
 
@@
 or an analogous one, e.g. a discrete version when the proposal is to change $N$.
+The {\tt DNest3} package monitors the acceptance rate as a function of level.
+In most applications, the acceptance probability is close to 1 (or may be
+precisely 1) when sampling the prior, and decreases to 5-40\% when sampling
+high levels.
 

### 2014-12-07 Documents/Paper/ms.tex
Three new NS references

@@
 the constrained prior distributions of Equation~\ref{eq:constrained_prior}.
+Various implementations of Nested Sampling exist, with different strategies
+for sampling Equation~\ref{eq:constrained_prior}
+\citep[e.g.][]{multinest, diamonds, radfriends}.
+
 Diffusive Nested Sampling
@@
 
+\bibitem[Buchner (2014)]{radfriends}
+Buchner, Johannes.~A statistical test for Nested Sampling algorithms.~Statistics
+and Computing (2014): 1-10.
+
+\bibitem[Corsaro and De 
+Ridder(2014)]{diamonds} Corsaro, E., De Ridder, J.\ 2014.\ DIAMONDS: A new Bayesian nested sampling tool. Application to peak bagging of solar-like oscillations.\ Astronomy and Astrophysics 571, AA71.
+
 \bibitem[Dolphin(2000)]{dolphot} Dolphin, A.~E.\ 2000, PASP, 
 112, 1383
+
+\bibitem[\protect\citeauthoryear{Feroz, Hobson,
+\& Bridges}{2009}]{multinest} Feroz F., Hobson M.~P., Bridges M., 2009, MNRAS, 398, 1601
 

### 2014-12-07 Documents/Paper/ms.tex
footnote

@@
 \citep[e.g.][]{multinest, diamonds, radfriends}.
-
 Diffusive Nested Sampling
 \citep[DNS][]{dnest} is an alternative version of Nested Sampling that is
-appropriate when MCMC is used to explore the parameter space.
+appropriate when MCMC is used to explore the parameter space\footnote{
 A C++ implementation of DNS, called {\tt DNest3}, is available at
-{https://github.com/eggplantbren/DNest3}.
+{\tt https://github.com/eggplantbren/DNest3}}.
 

### 2014-12-07 Documents/Paper/ms.tex
Slight tweaks talking about logl vs logx

@@
 \begin{eqnarray}
-X(l) &=& \int \pi(\theta)\mathds{1}\left[L(\theta) > l\right] \, d\theta
+X(l) &=& \int \pi(\theta)\mathds{1}\left[L(\theta) > l\right] \, d\theta\label{eq:X}
 \end{eqnarray}
@@
 and set to uniform once the desired number of levels $n$ has been created.
-See \citet{dnest}
-for more details about DNS.
+Since DNS does not sample the posterior distribution, the samples need to be
+assigned weights (in a manner similar to importance sampling) to represent
+the posterior. See \citet{dnest} for more details about DNS.
 
@@
 We ran DNS on the simulated data shown in Figure~\ref{fig:sinewave_data}.
-The shape of the likelihood function with respect to the prior mass (the
-usual likelihood curve associated with Nested Sampling) is shown in
-Figure~\ref{fig:sinewaves_likelihood}. Two phase transitions can be seen,
-where this curve is concave up. The
+The (log) likelihood as a function of (log) enclosed prior mass $X$
+(Equation~\ref{eq:X}) is plotted in Figure~\ref{fig:sinewaves_likelihood}
+This curve is a standard output of Nested Sampling methods and can be used to
+gain insight into the structure of the problem. Particularly, when
+phase transitions are present, this curve will have concave-up
+regions \citep{skilling}.
+
+Two phase transitions can be seen in Figure~\ref{fig:sinewaves_likelihood}. The
 first phase transition, at $\log(X) \approx -10$ nats,

### 2014-12-08 Documents/Paper/ms.tex
INS paper

@@
 Various implementations of Nested Sampling exist, with different strategies
-for sampling Equation~\ref{eq:constrained_prior}
-\citep[e.g.][]{multinest, diamonds, radfriends}.
+for sampling Equation~\ref{eq:constrained_prior}, as well as other features
+to infer whether the run was successful, or to improve the accuracy of
+the marginal likelihood estimate
+\citep[e.g.][]{multinest, importance, diamonds, radfriends}.
 Diffusive Nested Sampling
@@
 
+\bibitem[Feroz et al.(2013)]{importance} Feroz, F., Hobson, M.~P., 
+Cameron, E., Pettitt, A.~N.\ 2013.\ Importance Nested Sampling and the 
+MultiNest Algorithm.\ ArXiv e-prints arXiv:1306.2144.
+
 \bibitem[\protect\citeauthoryear{Feroz, Hobson,

### 2014-12-08 Documents/Paper/ms.tex
tweaks

@@
 where $\delta_N$ is drawn from a heavy tailed
-distribution which is a discrete analogue of Equation~\ref{eqn:heavytailed}.
+distribution which is a discrete analogue of Equation~\ref{eqn:heavytailed}
+($N$ is treated as a real number in $[0, N_{\rm max} + 1]$ and then we take
+the integer part at the end).
 The most probable values of $\delta_N$ are $\pm 1$, but values
@@
 are drawn from their prior conditional on the current value of the
-hyperparameters, i.e. the interim prior $p(\xx | \hyperparams)$.
+hyperparameters, i.e. the conditional prior $p(\xx | \hyperparams)$.
 If $N' < N$, i.e. the proposal is to remove components from the model,
@@
 If the component $x$ consists of a single scalar value, $F$ is the cumulative
-distribution of the interim prior. Denote the inverse of $F$ by $G$.
+distribution of the conditional prior. Denote the inverse of $F$ by $G$.
 
@@
 current value of the hyperparameters, and then transformed back using the
-proposed value of the hyperparameters so they represent the new interim prior
+proposed value of the hyperparameters so they represent the new conditional prior
 rather than the old one.
@@
 \end{eqnarray}
-For the interim prior $p(\xx_i | \hyperparams)$, we introduced a single
+For the conditional prior $p(\xx_i | \hyperparams)$, we introduced a single
 hyperparameter $\mu$, such that $A_i$ given $\mu$ has an exponential prior
@@
 \subsection{Priors}
-For the interim prior for the total fluxes of the galaxies, and for the
+For the conditional prior for the total fluxes of the galaxies, and for the
 widths of the galaxies, we used the Pareto
@@
 was based on the fact that a Pareto distribution for $x$ is an exponential
-distribution for $\ln(x)$, with scale length $1/a$.
+distribution for $\ln(x)$, with scale length $1/a$. If we do not expect the
+scale length to be very extreme then we might want to restrict it to $[0,1]$
+with a uniform prior.
 
-The interim priors for the radius ratios $u$ and the flux ratios $w$ were both
+The conditional priors for the radius ratios $u$ and the flux ratios $w$ were both
 uniform distributions. Since $u \in [0,1]$ and $v \in [0,1]$, the upper limits
-of the uniform interim priors must be between 0 and 1. Therefore the prior
+of the uniform conditional priors must be between 0 and 1. Therefore the prior
 for the upper limit was a uniform distribution between 0 and 1, and the prior

### 2014-12-08 Documents/Paper/ms.tex
log^2 volume ratio in DNS

@@
 the slab and the spike will be comparable. At these temperatures the MCMC
-method will be unable to mix and will give an incorrect estimate of
+method will be unable to mix (the sampler will only visit the spike or the
+slab, and not both) and will give an incorrect estimate of
 $\left<\log\left[L(\theta)\right]\right>_\beta$.
@@
 the number of iterations required to compress from the slab to the spike is
-of order $\log(V_{\rm slab}/V_{\rm spike})$. During this process Nested Sampling
+of order $\log(V_{\rm slab}/V_{\rm spike})$. In Diffusive Nested Sampling
+(Section~\ref{sec:dns}), it takes of order $\log(V_{\rm slab}/V_{\rm spike})$
+iterations to pass from the slab to the spike initially, and then
+$\log(V_{\rm slab}/V_{\rm spike})^2$ time to revise this.
+During this process Nested Sampling
 visits states with likelihoods intermediate between those of the slab and the
@@
 
-\subsection{Diffusive Nested Sampling}
+\subsection{Diffusive Nested Sampling}\label{sec:dns}
 The main difficulty with Nested Sampling is coming up with methods to sample

### 2014-12-08 Documents/Paper/ms.tex
Fixed awkward sentence
cite massinf

@@
 switching the identities of the components. Finite mixture modelling is a common
-statistical example.
+statistical example, but this kind of problem appears in many other contexts
+\citep[e.g.][]{massinf}.
 
@@
 
-This is similar to the idea of annealed importance sampling
+The idea of replacing the target distribution with something other than the
+posterior is similar to annealed importance sampling
 \citep{neal} or parallel tempering \citep{pt},

### 2015-01-14 Documents/Paper/ms.tex
png figures
Italic captions

@@
 \includegraphics{pgm.pdf}
-\caption{A probabilistic graphical model (PGM) depicting the kind
+\caption{\it A probabilistic graphical model (PGM) depicting the kind
 of model discussed in this paper. Produced using daft ({\tt daft-pgm.org}).
@@
 \end{minipage}
-\caption{{\bf Left}: An example of a two-dimensional posterior distribution
+\caption{\it {\bf Left}: An example of a two-dimensional posterior distribution
 with multiple modes, strong dependence, and a phase transition (in the top
@@
 \includegraphics[scale=0.5]{sinewave_data.pdf}
-\caption{The simulated data for the sinusoidal example. The solid line shows
+\caption{\it The simulated data for the sinusoidal example. The solid line shows
 the true signal which consists of a large slow oscillation and a much smaller
@@
 \includegraphics[scale=0.5]{trace_logl.pdf}
-\caption{A ``trace plot'' of the log likelihood in the sinusoidal model, for an
+\caption{\it A ``trace plot'' of the log likelihood in the sinusoidal model, for an
 MCMC chain designed to sample the posterior distribution. The two phases
@@
 \begin{center}
-\includegraphics[scale=0.5]{sinewaves_likelihood.pdf}
-\caption{{\bf Top panel: }
+\includegraphics[scale=0.6]{sinewaves_likelihood.png}
+\caption{\it {\bf Top panel: }
 The shape of the likelihood function with respect to prior mass.
@@
 \includegraphics[scale=0.5]{N_result.pdf}
-\caption{The inference for $N$, the number of sinusoids, based on the
+\caption{\it The inference for $N$, the number of sinusoids, based on the
 sinusoid data. The true value was $N=2$, which is also the most probable value
@@
 \includegraphics[scale=0.5]{trace_logl.pdf}
-\caption{A ``trace plot'' of the log likelihood in the SineWave model, for an
+\caption{\it A ``trace plot'' of the log likelihood in the SineWave model, for an
 MCMC chain designed to sample the posterior distribution. The two phases
@@
 \includegraphics[scale=0.7]{galaxyfield_data.pdf}
-\caption{The simulated image for the GalaxyField example. The image is
+\caption{\it The simulated image for the GalaxyField example. The image is
 $200\times 200$ pixels in size and contains $N=47$ ``galaxies''.
@@
 \begin{center}
-\includegraphics[scale=0.5]{galaxyfield_likelihood.pdf}
-\caption{{\bf Top panel: }
+\includegraphics[scale=0.6]{galaxyfield_likelihood.png}
+\caption{\it {\bf Top panel: }
 The shape of the likelihood function with respect to prior mass for the
@@
 \includegraphics[scale=0.5]{N_galaxy_result.pdf}
-\caption{The inference for $N$, the number of galaxies, based on the
+\caption{\it The inference for $N$, the number of galaxies, based on the
 GalaxyField data. The true value was $N=47$.

### 2015-01-25 Documents/Talk/content.tex
Projects
Wokring on the talk

@@
 
+\begin{frame}[t]{Motivating Problem I}
+Problems like this arise both inside and outside of astronomy,
+for example:
+\vspace{20pt}
+\begin{itemize}
+\item Asteroseismology
+\item Extrasolar planets (radial velocity technique)
+\end{itemize}
+\end{frame}
+
 \begin{frame}[t]{Motivating Problem II}
@@
 \end{center}
+\end{frame}
+
+\begin{frame}[t]{Motivating Problem II}
+This is a fundamental task in astronomy!
 \end{frame}
@@
 \begin{frame}[t]{Sinusoid data movie}
-
+{\tt https://www.youtube.com/watch?v=q4s8NJyUVoY}
 \end{frame}
@@
 {\tt https://www.youtube.com/watch?v=s1udKctW2HY}
+\end{frame}
+
+\begin{frame}[t]{Real projects}
+\begin{itemize}
+\item Magnetar bursts
+\item Making probabilistic catalogs
+\item Gravitational lens time delays in unresolved systems
+\item Gravitational lens inversion
+\item Modelling the progress of parallel Nested Sampling runs (!)
+\end{itemize}
 \end{frame}

### 2015-01-31 Documents/Talk/content.tex
starfield image

@@
 \begin{itemize}
-\item How many galaxies are in this data?
-\item Also, what are their positions, sizes, orientations, etc?
+\item How many stars are in these images?
+\item Also, what are their positions, fluxes, etc?
 \end{itemize}
 \begin{center}
-\includegraphics[scale=0.35]{../Paper/galaxyfield_data.pdf}
+\includegraphics[scale=0.35]{test_cases.pdf}
 \end{center}
@@
 \end{center}
+\end{frame}
+
+\begin{frame}[t]{View of data analysis}
+All conclusions are probabilities.
+
+The only difference between data analysis methods is what prior information
+they assume.
 \end{frame}

### 2015-02-01 Documents/Talk/content.tex
Exoplanet figure

@@
 All conclusions are probabilities.
+\vspace{20pt}
 
 The only difference between data analysis methods is what prior information
-they assume.
+they assume, and what numerical methods are used to calculate the result.
 \end{frame}
@@
 that starts with the prior distribution, and shrinks volume at a controlled
-rate. Maps big space into easy 1-D problem.
+rate.
+\vspace{20pt}
 
-Resolves phase transition problem, and helps a little with multiple modes \&
-strong degeneracies. Figure from Skilling (2006).
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.2]{ns.png}
-\end{center}
-\end{figure}
-
+{\bf Nested Sampling solves the problem of phase transitions}.
 \end{frame}
@@
 
+\begin{frame}[t]{Exoplanets}
+Radial velocity data is considered challenging
+\end{frame}
+
+\begin{frame}[t]{Exoplanets: Simulated Dataset}
+How many planets?
+\begin{center}
+\includegraphics[scale=0.35]{ExoplanetFigures/fake_data.pdf}
+\end{center}
+\end{frame}
+
 \begin{frame}[t]{Real projects}

### 2015-02-01 Documents/Talk/content.tex
StarField results
Working on the slides

@@
 
+\begin{frame}[t]{Marginal Likelihood and Model Selection}
+The marginal likelihood is more than a normalising constant. Consider
+estimating parameters $\theta$ within a model $M$. The marginal likelihood
+is really
+\begin{equation}
+p(D|M) = \int p(\theta | M)p(D|\theta, M) \, d\theta
+\end{equation}
+which can
+be used as a likelihood if we want to calculate the plausibility of $M$ vs.
+an alternative.
+\end{frame}
+
 \begin{frame}[t]{Posterior Distribution vs. Maximum Likelihood}
@@
 \begin{frame}[t]{View of data analysis}
-All conclusions are probabilities.
+All conclusions are probabilities, and probabilities are plausibilities.
 \vspace{20pt}
@@
 
-\begin{frame}[t]{Birth and Death}
+\begin{frame}[t]{Popular but slow approach}
+The marginal posterior distribution for $N$ be written down directly
+using Bayes' rule with $N$ as the only unknown parameter:
+\begin{eqnarray}
+p(N | D) &=& \frac{p(N)p(D | N)}{\sum_N p(N)p(D | N)}.
+\end{eqnarray}
+where
+\begin{eqnarray}
+p(D | N) &=& \int p(\theta | N) p(D | \theta, N) \, d^n \theta
+\end{eqnarray}
+\end{frame}
+
+\begin{frame}[t]{Popular but slow approach}
+Use many trial values of $N$ and use some Monte Carlo method to do the integral
+for each $N$.
+\end{frame}
+
+\begin{frame}[t]{Alternative: Birth and Death}
 For problems of unknown dimensionality, the hypothesis space is the union
@@
 
-\begin{frame}[t]{Results from the galaxy data}
+\begin{frame}[t]{Motivating Problem II}
+\begin{itemize}
+\item How many stars are in these images?
+\item Also, what are their positions, fluxes, etc?
+\end{itemize}
 \begin{center}
-\includegraphics[scale=0.45]{../Paper/N_galaxy_result.pdf}
+\includegraphics[scale=0.35]{test_cases.pdf}
 \end{center}
@@
 
-\begin{frame}[t]{Galaxy data movie}
-{\tt https://www.youtube.com/watch?v=s1udKctW2HY}
+\begin{frame}[t]{Probabilistic catalogs}
+\begin{center}
+\includegraphics[scale=0.25]{StarFieldFigures/catalogs.pdf}
+\end{center}
+\end{frame}
+
+\begin{frame}[t]{Hyperparameters}
+\begin{center}
+\includegraphics[scale=0.3]{StarFieldFigures/inference2.pdf}
+\end{center}
+\end{frame}
+
+\begin{frame}[t]{Luminosity Function}
+\begin{center}
+\includegraphics[scale=0.35]{StarFieldFigures/luminosity_function.pdf}
+\end{center}
 \end{frame}
@@
 \begin{frame}[t]{Exoplanets}
-Radial velocity data is considered challenging
+Radial velocity data is considered challenging. There are many systems where
+controversy rages over $N$.
+\end{frame}
+
+\begin{frame}[t]{Exoplanets}
+Radial velocity data is considered challenging. There are many systems where
+controversy rages over $N$.
 \end{frame}

### 2015-02-02 Documents/Talk/content.tex
More work

@@
   \begin{itemize}
-  \item Generate {\bf proposal} from some distribution $q(\theta' | \theta)$
+  \item Generate {\bf proposal} from some {\it proposal distribution} $q(\theta' | \theta)$
   (e.g. slightly perturb the current position).
-  \item With probability$^*$ $\alpha = \min\left(1, \frac{p(\theta')p(D|\theta')}{p(\theta)p(D|\theta)}\right)$, accept the proposal (i.e. replace $\theta$ with $\theta'$).
+  \item With probability $\alpha = \min\left(1, \frac{q(\theta | \theta')}{q(\theta' | \theta)}\frac{p(\theta')p(D|\theta')}{p(\theta)p(D|\theta)}\right)$, accept the proposal (i.e.
+replace $\theta$ with $\theta'$).
   \item Otherwise, stay in the same place.
@@
 
+\begin{frame}[t]{Alternative: Birth and Death}
+Birth and death moves are surprisingly easy to implement. Design them to
+explore the prior (\& satisfy detailed balance), and then use
+\begin{eqnarray}
+\alpha = \min\left(1, \frac{p(D|\theta')}{p(D|\theta)}\right)
+\end{eqnarray}
+as the acceptance probability: i.e. just the ratio of likelihoods.
+\end{frame}
+
+\begin{frame}[t]{Alternative: Birth and Death}
+Birth proposal: Increment $N$, generate the new 'object' from the prior
+(given current values of hyperparameters)
+
+\vspace{20pt}
+
+Death proposal: Choose one of the $N$ objects, remove it, decrement $N$.
+\end{frame}
+
 \begin{frame}[t]{What's the issue?}
@@
 
-{\bf Nested Sampling solves the problem of phase transitions}.
+{\bf Nested Sampling solves the problem of phase transitions and (as far as I
+know) no other Monte Carlo method does.}
 \end{frame}
@@
   \item Replace the posterior distribution with a {\bf mixture of constrained priors}
-  \item Can cross valleys
+  \item Can cross valleys (the prior is a mixture component)
   \item Posterior is available via re-weighting
@@
 \setlength{\itemsep}{20pt}
-\item I've implemented the sinewave problem and the galaxy problem
-\item The common features (what proposals to use?) are delegated to
-a separate library called {\tt RJObject}
+\item I've implemented a library, {\tt RJObject}, that handles all the
+Metropolis proposals I need for these kinds of problems.
 \item Uses birth-death MCMC, but exploring the DNS distribution, rather than
@@
 \begin{frame}[t]{Sinusoid data movie}
-{\tt https://www.youtube.com/watch?v=q4s8NJyUVoY}
+{\tt https://www.youtube.com/watch?v=i-f4VqPPeGg}
 \end{frame}
@@
 
+\begin{frame}[t]{Recommendations}
+It is infeasable to apply this approach wholesale.
+Instead I would use it on small subsets of imaging that are of specific interest.
+\end{frame}
+
 \begin{frame}[t]{Exoplanets}
@@
 controversy rages over $N$.
+\end{frame}
+
+\begin{frame}[t]{Exoplanets: Simulated Dataset}
+How many planets?
+\begin{center}
+\includegraphics[scale=0.35]{ExoplanetFigures/fake_data.pdf}
+\end{center}
 \end{frame}

### 2015-02-02 Documents/Talk/content.tex
Small edits
More figs etc

@@
 \begin{column}{0.35\textwidth}
-  \vspace{30pt}
+  \vspace{10pt}
   \begin{itemize}
@@
 \begin{center}
-\includegraphics[scale=0.3]{StarFieldFigures/inference2.pdf}
+\includegraphics[scale=0.28]{StarFieldFigures/inference2.pdf}
 \end{center}
@@
 \begin{frame}[t]{Recommendations}
-It is infeasable to apply this approach wholesale.
+It is infeasible to apply this approach wholesale.
 Instead I would use it on small subsets of imaging that are of specific interest.
-\end{frame}
 
-\begin{frame}[t]{Exoplanets}
-Radial velocity data is considered challenging. There are many systems where
-controversy rages over $N$.
+\vspace{20pt}
+
+\begin{center}
+{\it Any takers?}
+\end{center}
 \end{frame}
@@
 
-\begin{frame}[t]{Exoplanets: Simulated Dataset}
-How many planets?
+\begin{frame}[t]{Exoplanets: Simulated Dataset Results}
+The true value was $N$=7.
 \begin{center}
-\includegraphics[scale=0.35]{ExoplanetFigures/fake_data.pdf}
+\includegraphics[scale=0.35]{ExoplanetFigures/fake_data_N.pdf}
+\end{center}
+\end{frame}
+
+\begin{frame}[t]{Exoplanets: Simulated Dataset Results}
+This is the posterior distribution for $P_1$. Keep in mind the {\bf
+label-switching degeneracy}.
+\begin{center}
+\includegraphics[scale=0.35]{ExoplanetFigures/fake_data_periods.pdf}
+\end{center}
+\end{frame}
+
+\begin{frame}[t]{Exoplanets: Simulated Dataset Results}
+Here, we can see why one of the planets was not ``detected'':
+\begin{center}
+\includegraphics[scale=0.35]{ExoplanetFigures/fake_data_posterior.pdf}
+\end{center}
+\end{frame}
+
+\begin{frame}[t]{Time}
+It takes about 15-20 minutes to obtain some posterior samples.
+
+Compare with Phil Gregory's code (30 minutes per planet, no trans dimensional
+moves).
+\end{frame}
+
+\begin{frame}[t]{Gliese 581}
+Gliese 581 is an M dwarf which has been observed extensively with the HARPS
+and UVES spectrographs.
+
+\vspace{20pt}
+
+There is controversy about $N$, some people only
+believe {\bf two} of the claimed planets, while other analyses suggest
+{\bf up to 6} (with a few of these in the habitable zone).
+\end{frame}
+
+\begin{frame}[t]{Giant Disclaimer}
+\begin{center}
+\includegraphics[scale=0.25]{robertson.png}
+\end{center}
+\end{frame}
+
+\begin{frame}[t]{Giant Disclaimer}
+\begin{center}
+\includegraphics[scale=0.25]{robertson.png}
 \end{center}

### 2015-02-02 Documents/Talk/content.tex
Amplitudes comment
More figures

@@
 \end{itemize}
+\end{frame}
+
+\begin{frame}[t]{Motivating Problem I}
+\begin{itemize}
+\item How many sinusoids are in this data?
+\item {\it Note that this must depend on what we believe about their properties,
+particularly the amplitudes}
+\end{itemize}
+\begin{center}
+\includegraphics[scale=0.35]{sinewave_data.pdf}
+\end{center}
 \end{frame}
@@
 \begin{frame}[t]{Recommendations}
-It is infeasible to apply this approach wholesale.
-Instead I would use it on small subsets of imaging that are of specific interest.
+It is infeasible to apply this approach wholesale (i.e. to a whole survey).
+Instead I would use it on very small subsets of imaging that are of specific
+interest.
 
@@
 
-\begin{frame}[t]{Giant Disclaimer}
+\begin{frame}[t]{Gliese 581 -- Number of Signals}
 \begin{center}
-\includegraphics[scale=0.25]{robertson.png}
+\includegraphics[scale=0.35]{ExoplanetFigures/gliese581_N.pdf}
+\end{center}
+\end{frame}
+
+\begin{frame}[t]{Gliese 581 -- Periods}
+\begin{center}
+\includegraphics[scale=0.4]{ExoplanetFigures/gliese581_periods.pdf}
+\end{center}
+\end{frame}
+
+\begin{frame}[t]{Gliese 581 -- Model Fit (Part of Dataset)}
+\begin{center}
+\includegraphics[scale=0.4]{ExoplanetFigures/gliese581.pdf}
+\end{center}
+\end{frame}
+
+\begin{frame}[t]{Gliese 581 -- Likelihood vs Enclosed Prior Mass}
+\begin{center}
+\includegraphics[scale=0.4]{ExoplanetFigures/logl.pdf}
 \end{center}

### 2015-02-02 Documents/Talk/content.tex
References

@@
 
-\begin{frame}[t]{Real projects}
+\begin{frame}[t]{References}
 \begin{itemize}
-\item Magnetar bursts
-\item Making probabilistic catalogs
-\item Gravitational lens time delays in unresolved systems
-\item Gravitational lens inversion
-\item Modelling the progress of parallel Nested Sampling runs (!)
+\item Diffusive Nested Sampling (0912.2380)
+\item Birth-death MCMC in DNS (1411.3921)
+\item Radial Velocity Data (1501.06952)
+\item Probabilistic Catalogs (1211.5805)
+\item Magnetar Variability (1501.05251)
 \end{itemize}
@@
 \begin{frame}[t]{Software}
-\begin{itemize}
-\setlength{\itemsep}{20pt}
-\item Diffusive Nested Sampling in C++
-{\tt github.com/eggplantbren/DNest3}
-\item Birth-death models in C++
-{\tt github.com/eggplantbren/RJObject}
-\item These are free software (GPLd).
-\item The learning curve is steep (sorry) but I'm open to collaborations.
-\end{itemize}
+Everything I've discussed today is GPLd and is on Github.
 \end{frame}

### 2015-09-10 Documents/manual.tex

% Every Latex document starts with a documentclass command
\documentclass[a4paper, 11pt]{article}

% Load some packages
\usepackage{graphicx} % This allows you to put figures in
\usepackage{natbib}   % This allows for relatively pain-free reference lists
\usepackage[left=2cm,top=3cm,right=2cm]{geometry} % The way I like the margins
\usepackage{framed}

\newcommand{\rjobject}{{\tt RJObject}}
\newcommand{\dnest}{{\tt DNest3}}

% This helps with figure placement
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\parindent=0cm

% Set values so you can have a title
\title{RJObject Manual}
\author{Brendon J. Brewer}
\date{\today}

% Document starts here
\begin{document}

% Actually put the title in
\maketitle

\section{Introduction}
RJObject is a C++ library for implementing certain kinds of hierarchical
models in DNest3.

\end{document}

### 2015-09-10 Documents/Paper/ms.tex
Removed duplicate figure
updated numbers

@@
 \begin{center}
-\includegraphics[scale=0.5]{trace_logl.pdf}
-\caption{\it A ``trace plot'' of the log likelihood in the sinusoidal model, for an
-MCMC chain designed to sample the posterior distribution. The two phases
-have log-likelihoods around -730 and -745 respectively, however transitions
-between the two phases are quite uncommon.
-\label{fig:trace_logl}}
-\end{center}
-\end{figure}
-
-\begin{figure}
-\begin{center}
 \includegraphics[scale=0.6]{sinewaves_likelihood.png}
@@
 The marginal likelihood estimate returned by DNS for this data was
-$\log(\mathcal{Z}) = -771.8$, and the information, or Kullback-Leibler
-divergence from the prior to the posterior, was estimated to be 39.8 nats.
+$\log(\mathcal{Z}) = -772.7$, and the information, or Kullback-Leibler
+divergence from the prior to the posterior, was estimated to be 41.4 nats.
 

### 2015-09-13 Documents/Paper/ms.tex
portillo
Some notation changes

@@
 \newcommand{\hyperparams}{\boldsymbol{\alpha}}
-\newcommand{\xx}{\mathbf{x}}
+\newcommand{\xx}{\boldsymbol{x}}
+\newcommand{\data}{\boldsymbol{D}}
 
@@
 components in some region, along with their properties $\{\xx_i\}_{i=1}^N$,
-from a dataset $\mathcal{D}$. A common statistical example is finite mixture
+from a dataset $\data$. A common statistical example is finite mixture
 modelling. In the Bayesian framework, these problems are typically solved using
@@
 the conditional prior for $\{\xx_i\}$ is iid and does not depend on $N$.
-Given data $\mathcal{D}$, we usually want to calculate properties of the
+Given data $\data$, we usually want to calculate properties of the
 posterior distribution, given by:
 \begin{eqnarray}
-p(N, \hyperparams, \{\xx_i\} | \mathcal{D}) \propto
+p(N, \hyperparams, \{\xx_i\} | \data) \propto
 p(N, \hyperparams, \{\xx_i\})
-p(\mathcal{D} | N, \hyperparams, \{\xx_i\})\label{eq:bayes}
+p(\data | N, \hyperparams, \{\xx_i\})\label{eq:bayes}
 \end{eqnarray}
@@
 \begin{eqnarray}
-\mathcal{Z} &=& p(\mathcal{D})\\
+\mathcal{Z} &=& p(\data)\\
 &=& \sum_{i=0}^{N_{\rm max}} \int
-p(N, \hyperparams, \{\xx_i\})p(\mathcal{D} | N, \hyperparams, \{\xx_i\})
+p(N, \hyperparams, \{\xx_i\})p(\data | N, \hyperparams, \{\xx_i\})
 \, d^N \xx_i \, d\hyperparams \label{eq:evidence}
@@
 John Skilling (MaxEnt Data Consultants), and Daniela Huppenkothen
-(Amsterdam, NYU).
+(Amsterdam, NYU). Stephen Portillo (Harvard) discovered a significant bug.
 

### 2015-09-13 Examples/NDMixture/MyDistribution.cpp
Comments and from_uniform method

@@ -98,16 +98,18 @@ double MyDistribution::log_pdf(const std::vector<double>& vec) const
 	return logp;
 }
 
-//void MyDistribution::from_uniform(std::vector<double>& vec) const
-//{
-//	vec[0] = center_locations
-//			+ diversity_locations*gsl_cdf_ugaussian_Pinv(vec[0]);
+void MyDistribution::from_uniform(std::vector<double>& vec) const
+{
+	for(size_t i=0; i<min.size(); i++)
+	{
+		vec[i] = center_locations[i]
+				+ diversity_locations[i]*gsl_cdf_ugaussian_Pinv(vec[i]);
 
-//	vec[1] = center_logwidths
-//			+ diversity_logwidths*gsl_cdf_ugaussian_Pinv(vec[1]);
-
-//	vec[2] = diversity_logweights*gsl_cdf_ugaussian_Pinv(vec[2]);
-//}
+		vec[min.size() + i] = center_logwidths[i]
+				+ diversity_logwidths[i]*gsl_cdf_ugaussian_Pinv(vec[min.size() + i]);
+	}
+	vec[min.size()] = diversity_logweights*gsl_cdf_ugaussian_Pinv(vec[min.size()]);
+}
 
 //void MyDistribution::to_uniform(std::vector<double>& vec) const
 //{

### 2015-09-13 Examples/NDMixture/MyDistribution.h

@@ -6,15 +6,18 @@
 class MyDistribution:public Distribution
 {
 	private:
-		// Limits
+		// Limits for the mean vector
 		std::vector<double> min, max, range;
 
+		// These determine the conditional prior for a mean vector
 		std::vector<double> center_locations;
 		std::vector<double> diversity_locations;
 
+		// These determine the conditional prior for gaussian sigmas
 		std::vector<double> center_logwidths;
 		std::vector<double> diversity_logwidths;
 
+		// Determines the conditional prior for the weights
 		double diversity_logweights;
 
 		double perturb_parameters();

### 2015-09-13 Examples/NDMixture/MyDistribution.cpp
to_uniform method

@@ -111,12 +111,15 @@ void MyDistribution::from_uniform(std::vector<double>& vec) const
 	vec[min.size()] = diversity_logweights*gsl_cdf_ugaussian_Pinv(vec[min.size()]);
 }
 
-//void MyDistribution::to_uniform(std::vector<double>& vec) const
-//{
-//	vec[0] = gsl_cdf_ugaussian_P((vec[0] - center_locations)/diversity_locations);
-//	vec[1] = gsl_cdf_ugaussian_P((vec[1] - center_logwidths)/diversity_logwidths);
-//	vec[2] = gsl_cdf_ugaussian_P(vec[2]/diversity_logweights);
-//}
+void MyDistribution::to_uniform(std::vector<double>& vec) const
+{
+	for(size_t i=0; i<min.size(); i++)
+	{
+		vec[i] = gsl_cdf_ugaussian_P((vec[i] - center_locations[i])/diversity_locations[i]);
+		vec[min.size() + i] = gsl_cdf_ugaussian_P((vec[min.size() + i] - center_logwidths[i])/diversity_logwidths[i]);
+	}
+	vec[min.size()] = gsl_cdf_ugaussian_P(vec[min.size()]/diversity_logweights);
+}
 
 //void MyDistribution::print(std::ostream& out) const
 //{

### 2015-09-13 Examples/NDMixture/MyDistribution.cpp
print method

@@ -121,10 +121,14 @@ void MyDistribution::to_uniform(std::vector<double>& vec) const
 	vec[min.size()] = gsl_cdf_ugaussian_P(vec[min.size()]/diversity_logweights);
 }
 
-//void MyDistribution::print(std::ostream& out) const
-//{
-//	out<<center_locations<<' '<<diversity_locations<<' ';
-//	out<<center_logwidths<<' '<<diversity_logwidths<<' ';
-//	out<<diversity_logweights<<' ';
-//}
+void MyDistribution::print(std::ostream& out) const
+{
+	for(size_t i=0; i<min.size(); i++)
+		out<<center_locations[i]<<' ';
+	for(size_t i=0; i<min.size(); i++)
+		out<<diversity_locations[i]<<' ';
+	for(size_t i=0; i<min.size(); i++)
+		out<<center_logwidths[i]<<' ';
+	out<<diversity_logweights<<' ';
+}
 

### 2015-09-13 Examples/NDMixture/MyDistribution.cpp
Use Data::num_dimensions instead of vector sizes

@@ -12,22 +12,22 @@ using namespace DNest3;
 MyDistribution::MyDistribution(const vector<double>& min, const vector<double>& max)
 :min(min)
 ,max(max)
-,range(max.size())
-,center_locations(max.size())
-,diversity_locations(max.size())
-,center_logwidths(max.size())
-,diversity_logwidths(max.size())
+,range(Data::num_dimensions)
+,center_locations(Data::num_dimensions)
+,diversity_locations(Data::num_dimensions)
+,center_logwidths(Data::num_dimensions)
+,diversity_logwidths(Data::num_dimensions)
 {
 	assert(min.size() == Data::num_dimensions);
 	assert(max.size() == Data::num_dimensions);
 
-	for(size_t i=0; i<range.size(); i++)
+	for(int i=0; i<Data::num_dimensions; i++)
 		range[i] = max[i] - min[i];
 }
 
 void MyDistribution::fromPrior()
 {
-	for(size_t i=0; i<min.size(); i++)
+	for(int i=0; i<Data::num_dimensions; i++)
 	{
 		center_locations[i] = min[i] + range[i]*randomU();
 		diversity_locations[i] = exp(log(1E-3) + log(1E3)*randomU())*range[i];
@@ -44,7 +44,7 @@ double MyDistribution::perturb_parameters()
 	double logH = 0.;
 
 	int which = randInt(5);
-	int i = randInt(min.size());
+	int i = randInt(Data::num_dimensions);
 
 	if(which == 0)
 	{
@@ -85,12 +85,12 @@ double MyDistribution::log_pdf(const std::vector<double>& vec) const
 {
 	double logp = 0.;
 
-	for(size_t i=0; i<min.size(); i++)
+	for(int i=0; i<Data::num_dimensions; i++)
 	{
 		logp += -log(diversity_locations[i])
 			-0.5*pow((vec[i] - center_locations[i])/diversity_locations[i], 2);
 		logp += -log(diversity_logwidths[i])
-			-0.5*pow((vec[min.size() + i] - center_logwidths[i])/diversity_logwidths[i], 2);
+			-0.5*pow((vec[Data::num_dimensions + i] - center_logwidths[i])/diversity_logwidths[i], 2);
 	}
 	logp += -log(diversity_logweights)
 		-0.5*pow(vec[2]/diversity_logweights, 2);
[…]

### 2015-09-13 Examples/NDMixture/MyModel.cpp
Starting to work on MyModel

@@ -1,0 +1,85 @@
+#include "MyModel.h"
+#include "RandomNumberGenerator.h"
+#include "Utils.h"
+#include "Data.h"
+#include <cmath>
+
+using namespace std;
+using namespace DNest3;
+
+MyModel::MyModel()
+:objects(1 + 2*Data::num_dimensions, 10, false,
+		MyDistribution(vector<double>(Data::num_dimensions, -10.), 
+						vector<double>(Data::num_dimensions, 10.)))
+{
+
+}
+
+void MyModel::fromPrior()
+{
+	objects.fromPrior();
+}
+
+double MyModel::perturb()
+{
+	double logH = 0.;
+
+	logH += objects.perturb();
+
+	return logH;
+}
+
+//double MyModel::logLikelihood() const
+//{
+//	// Get the data
+//	const vector<double>& x = Data::get_instance().get_x();
+
+//	// Get mixture components
+//	const vector< vector<double> >& components = objects.get_components();
+
+//	// Can't have zero mixture components
+//	if(components.size() == 0)
+//		return -1E200;
+
+//	// Calculate normalised log weights of mixture components
+//	vector<double> logw(components.size());
+//	double tot = -1E300;
+//	for(size_t i=0; i<components.size(); i++)
+//	{
+//		logw[i] = components[i][2];
+//		tot = logsumexp(tot, logw[i]);
+//	}
+//	for(size_t i=0; i<components.size(); i++)
+//		logw[i] -= tot;
+
+//	double logL = 0.;
+
+//	// Loop over datapoints
+//	for(size_t i=0; i<x.size(); i++)
+//	{
+//		// Loop over mixture components
+//		double temp = -1E300;
+//		double gaussian;
+//		for(size_t j=0; j<components.size(); j++)
+//		{
+//			gaussian = -0.5*log(2.*M_PI) - components[j][1]
+//				-0.5*pow((x[i] - components[j][0])/exp(components[j][1]), 2);
+//			temp = logsumexp(temp, logw[j] + gaussian);
+//		}
+
+//		logL += temp;
+//	}
+
+//	return logL;
+//}
+
+//void MyModel::print(std::ostream& out) const
+//{
+//	objects.print(out); out<<' ';
+//}
+
+//string MyModel::description() const
+//{
+//	return string("objects");
+//}
+

### 2015-09-13 Examples/NDMixture/MyModel.cpp
log likelihood etc

@@ -29,57 +29,61 @@ double MyModel::perturb()
 	return logH;
 }
 
-//double MyModel::logLikelihood() const
-//{
-//	// Get the data
-//	const vector<double>& x = Data::get_instance().get_x();
+double MyModel::logLikelihood() const
+{
+	// Get the data
+	const vector< vector<double> >& x = Data::get_instance().get_x();
 
-//	// Get mixture components
-//	const vector< vector<double> >& components = objects.get_components();
+	// Get mixture components
+	const vector< vector<double> >& components = objects.get_components();
 
-//	// Can't have zero mixture components
-//	if(components.size() == 0)
-//		return -1E200;
+	// Can't have zero mixture components
+	if(components.size() == 0)
+		return -1E200;
 
-//	// Calculate normalised log weights of mixture components
-//	vector<double> logw(components.size());
-//	double tot = -1E300;
-//	for(size_t i=0; i<components.size(); i++)
-//	{
-//		logw[i] = components[i][2];
-//		tot = logsumexp(tot, logw[i]);
-//	}
-//	for(size_t i=0; i<components.size(); i++)
-//		logw[i] -= tot;
+	// Calculate normalised log weights of mixture components
+	vector<double> logw(components.size());
+	double tot = -1E300;
+	for(size_t i=0; i<components.size(); i++)
+	{
+		logw[i] = components[i].back();
+		tot = logsumexp(tot, logw[i]);
+	}
+	for(size_t i=0; i<components.size(); i++)
+		logw[i] -= tot;
 
-//	double logL = 0.;
+	double logL = 0.;
 
-//	// Loop over datapoints
-//	for(size_t i=0; i<x.size(); i++)
-//	{
-//		// Loop over mixture components
-//		double temp = -1E300;
-//		double gaussian;
-//		for(size_t j=0; j<components.size(); j++)
-//		{
-//			gaussian = -0.5*log(2.*M_PI) - components[j][1]
-//				-0.5*pow((x[i] - components[j][0])/exp(components[j][1]), 2);
-//			temp = logsumexp(temp, logw[j] + gaussian);
-//		}
+	double temp, gaussian;
+	// Loop over datapoints
+	for(size_t i=0; i<x.size(); i++)
+	{
+		// Loop over dimensions
+		for(size_t j=0; j<x[i].size(); j++)
+		{
+			// Loop over mixture components
+			temp = -1E300;
[…]

### 2015-09-13 Examples/NDMixture/main.cpp
Ready to roll

@@ -1,18 +1,18 @@
 #include <iostream>
-//#include "Start.h"
-//#include "MyModel.h"
+#include "Start.h"
+#include "MyModel.h"
 #include "Data.h"
 
 using namespace std;
-//using namespace DNest3;
+using namespace DNest3;
 
 int main(int argc, char** argv)
 {
 	// Load the data
 	Data::get_instance().load("data.txt");
 
-//	MTSampler<MyModel> sampler = setup_mt<MyModel>(argc, argv);
-//	sampler.run();
+	MTSampler<MyModel> sampler = setup_mt<MyModel>(argc, argv);
+	sampler.run();
 
 	return 0;
 }

### 2015-09-13 Examples/NDMixture/MyDistribution.cpp
Get rid of warnings

@@ -18,8 +18,8 @@ MyDistribution::MyDistribution(const vector<double>& min, const vector<double>& 
 ,center_logwidths(Data::num_dimensions)
 ,diversity_logwidths(Data::num_dimensions)
 {
-	assert(min.size() == Data::num_dimensions);
-	assert(max.size() == Data::num_dimensions);
+	assert(static_cast<int>(min.size()) == Data::num_dimensions);
+	assert(static_cast<int>(max.size()) == Data::num_dimensions);
 
 	for(int i=0; i<Data::num_dimensions; i++)
 		range[i] = max[i] - min[i];

### 2015-09-13 Examples/NDMixture/display.py
Scripts and things

@@ -1,0 +1,24 @@
+from pylab import *
+import os
+
+data = loadtxt('data.txt')
+posterior_sample = atleast_2d(loadtxt('posterior_sample.txt'))
+
+ion()
+
+for i in range(0, posterior_sample.shape[0]):
+	hold(False)
+	plot(data[:,0], data[:,1], 'b.')
+	hold(True)
+
+	x = posterior_sample[i, 10:20]
+	y = posterior_sample[i, 20:30]
+	x = x[x != 0.]
+	y = y[y != 0.]
+
+	plot(x, y, 'ro')
+	draw()
+
+ioff()
+show()
+

### 2015-09-13 Examples/NDMixture/generate_data.py
Clearer data

@@ -1,0 +1,15 @@
+from pylab import *
+
+seed(0)
+
+data = randn(1000, 2)
+data[0:500, 0] += 1.
+data[0:500, 1] += 2.
+data[500:, :] *= 0.5
+data[500:, 0] += 3.
+data[500:, 1] += 4.
+
+savetxt('data.txt', data)
+plot(data[:,0], data[:,1], 'b.')
+show()
+

### 2015-09-13 Examples/NDMixture/MyDistribution.cpp
Fixed one bug

@@ -77,7 +77,7 @@ double MyDistribution::perturb_parameters()
 	return logH;
 }
 
-// vec[0], ..., vec[num_dimensions-1] = central position
+// vec[0], ..., vec[num_dimensions-1] = position
 // vec[num_dimensions, ..., vec[2*num_dimensions-1] = log_width
 // vec[2*num_dimensions] = log_weight
 
@@ -93,7 +93,7 @@ double MyDistribution::log_pdf(const std::vector<double>& vec) const
 			-0.5*pow((vec[Data::num_dimensions + i] - center_logwidths[i])/diversity_logwidths[i], 2);
 	}
 	logp += -log(diversity_logweights)
-		-0.5*pow(vec[2]/diversity_logweights, 2);
+		-0.5*pow(vec[Data::num_dimensions]/diversity_logweights, 2);
 
 	return logp;
 }

### 2015-09-13 Examples/NDMixture/MyDistribution.cpp
OOPS! wrong element of vector

@@ -93,7 +93,7 @@ double MyDistribution::log_pdf(const std::vector<double>& vec) const
 			-0.5*pow((vec[Data::num_dimensions + i] - center_logwidths[i])/diversity_logwidths[i], 2);
 	}
 	logp += -log(diversity_logweights)
-		-0.5*pow(vec[Data::num_dimensions]/diversity_logweights, 2);
+		-0.5*pow(vec.back()/diversity_logweights, 2);
 
 	return logp;
 }
@@ -108,7 +108,7 @@ void MyDistribution::from_uniform(std::vector<double>& vec) const
 		vec[Data::num_dimensions + i] = center_logwidths[i]
 				+ diversity_logwidths[i]*gsl_cdf_ugaussian_Pinv(vec[Data::num_dimensions + i]);
 	}
-	vec[Data::num_dimensions] = diversity_logweights*gsl_cdf_ugaussian_Pinv(vec[Data::num_dimensions]);
+	vec.back() = diversity_logweights*gsl_cdf_ugaussian_Pinv(vec.back());
 }
 
 void MyDistribution::to_uniform(std::vector<double>& vec) const
@@ -118,7 +118,7 @@ void MyDistribution::to_uniform(std::vector<double>& vec) const
 		vec[i] = gsl_cdf_ugaussian_P((vec[i] - center_locations[i])/diversity_locations[i]);
 		vec[Data::num_dimensions + i] = gsl_cdf_ugaussian_P((vec[Data::num_dimensions + i] - center_logwidths[i])/diversity_logwidths[i]);
 	}
-	vec[Data::num_dimensions] = gsl_cdf_ugaussian_P(vec[Data::num_dimensions]/diversity_logweights);
+	vec.back() = gsl_cdf_ugaussian_P(vec.back()/diversity_logweights);
 }
 
 void MyDistribution::print(std::ostream& out) const

### 2015-09-13 Examples/NDMixture/MyModel.cpp
logl rewrite

@@ -54,24 +54,25 @@ double MyModel::logLikelihood() const
 
 	double logL = 0.;
 
-	double temp, gaussian;
-	// Loop over datapoints
+	double temp1, temp2;
+	// For all datapoints
 	for(size_t i=0; i<x.size(); i++)
 	{
-		// Loop over dimensions
-		for(size_t j=0; j<x[i].size(); j++)
+		// Evaluate the mixture of gaussians
+		// at this datapoint
+		temp1 = -1E300;
+		for(size_t j=0; j<components.size(); j++)
 		{
-			// Loop over mixture components
-			temp = -1E300;
-			for(size_t k=0; k<components.size(); k++)
+			temp2 = 0.;
+			for(int k=0; k<Data::num_dimensions; k++)
 			{
-				gaussian = -0.5*log(2.*M_PI) - components[k][j + Data::num_dimensions]
-					-0.5*pow((x[i][j] - components[k][j])/exp(components[k][j + Data::num_dimensions]), 2);
-				temp = logsumexp(temp, logw[k] + gaussian);
+				temp2 += -components[j][Data::num_dimensions+k]
+							-0.5*pow((x[i][k] - components[j][k])/
+								exp(components[j][Data::num_dimensions+k]), 2);
 			}
-
-			logL += temp;
+			temp1 = logsumexp(temp1, logw[j] + temp2);
 		}
+		logL += temp1;
 	}
 
 	return logL;

### 2015-09-19 Examples/GalaxyField/postprocess.py
Better OPTIONS and proper ESS

@@ -205,7 +205,7 @@ def postprocess(temperature=1., numResampleLogX=1, plot=True, loaded=[], \
 	print("Effective sample size = " + str(ESS))
 
 	# Resample to uniform weight
-	N = 300#int(ESS)
+	N = int(ESS)
 	posterior_sample = np.zeros((N, sample.shape[1]))
 	w = P_samples
 	w = w/np.max(w)

### 2015-09-20 Documents/Talk/content.tex
Lensing ref
Pre-talk edits

@@
 \item How many sinusoids are in this data?
-\item {\it Note that this must depend on what we believe about their properties,
+\item {\it This must depend on what we believe about their properties,
 particularly the amplitudes}
@@
 
-The only difference between data analysis methods is what prior information
-they assume, and what numerical methods are used to calculate the result.
+The only differences between data analysis methods:\\
+\vspace{20pt}
+\begin{enumerate}
+\item What prior information is assumed\\
+\item Which numerical methods are used to calculate the results.
+\end{enumerate}
 \end{frame}
@@
 \item Magnetar Variability (1501.05251)
+\item Strong gravitational lensing (1508.00662)
 \end{itemize}
-\end{frame}
 
-\begin{frame}[t]{Software}
 Everything I've discussed today is GPLd and is on Github.

### 2015-10-28 Documents/Paper/ms.tex
Updated numbers

@@
 The marginal likelihood estimate returned by DNS for this data was
-$\log(\mathcal{Z}) = -319707.6$, and the information, or Kullback-Leibler
-divergence from the prior to the posterior, was estimated to be 550.2 nats.
+$\log(\mathcal{Z}) = -319690.5$, and the information, or Kullback-Leibler
+divergence from the prior to the posterior, was estimated to be 593.7 nats.
 

### 2015-10-28 Examples/GalaxyField/postprocess.py
png

@@ -190,7 +190,7 @@ def postprocess(temperature=1., numResampleLogX=1, plot=True, loaded=[], \
 			if numResampleLogX > 1:
 				plt.draw()
 
-			plt.savefig('galaxyfield_likelihood.pdf', bbox_inches='tight')
+			plt.savefig('galaxyfield_likelihood.png', bbox_inches='tight', dpi=500)
 
 	P_samples = np.mean(P_samples, 1)
 	P_samples = P_samples/np.sum(P_samples)

### 2015-11-16 Documents/manual.tex
A bit of work on the manual

@@
 \section{Introduction}
-RJObject is a C++ library for implementing certain kinds of hierarchical
-models in DNest3.
+RJObject is a C++ library for implementing certain kinds of trans-dimensional
+hierarchical models (basically `mixture models')in DNest3. I developed this
+library after realising that many models I was implementing all had this
+common structure.
+Throughout this manual, I will assume you're proficient at
+DNest3. If not, please read the manual for that project first.
+
+\section{Dependencies}
+As well as all the DNest3 dependencies, it's best if you add some of DNest3's
+directories to your various PATH variables. Specifically, things will work best
+if you put the location of the DNest3 libraries (the location of
+the {\tt libdnest3.so} file) to your {\tt LIBRARY\_PATH} and
+{\tt LD\_LIBRARY\_PATH}. Also add the directory containing the DNest3 header
+files ({\tt *.h}) to your {\tt CPLUS\_INCLUDE\_PATH}. Finally, add the
+directory containing DNest3's {\tt postprocess.py} and {\tt dnest\_plots.py} to
+your {\tt PYTHON\_PATH}.
 

### 2015-11-16 Documents/elliptical.py

@@ -1,26 +1,0 @@
-from pylab import *
-
-N = 5000
-
-# Parameters
-q = 0.5
-
-# Defines Radial profile
-def R(u):
-	return -log(1. - u)
-
-# The uniforms
-u1 = rand(N)
-u2 = rand(N)
-
-# The actual positions
-x = R(u1)/sqrt(q)*cos(2.*pi*u2)
-y = R(u1)*sqrt(q)*sin(2.*pi*u2)
-
-plot(x, y, 'k.', markersize=1)
-axis('equal')
-xlabel('$x$')
-ylabel('$y$')
-show()
-
-

### 2015-12-07 Documents/manual.tex
Compiling
Specifying models

@@
 
+\section{Compiling}
+If you have all the dependencies, and have all the paths set up correctly as
+specified above, you should be able to compile RJObject by going into the
+root RJObject directory and invoking {\tt make}. Alternatively you can follow
+the {\tt cmake} instructions in the README.
+
+\section{Specifying models}
+When using RJObject, you're really using DNest3. So to implement a model, you
+need to write a DNest3 model class and specify the fromPrior, perturb,
+logLikelihood methods etc. All RJObject provides is a quick and relatively
+painless way of incorporating a trans-dimensional object into your model and
+specifying heirarchical priors for its parameters. For example, consider
+the SineWave example from the RJObject paper. There are an unknown number of
+sinusoids $N$, and each of the sinusoids has an unknown period $T_i$, amplitude
+$A_i$, and phase $\phi_i$. To use DNest3 on this problem, you'd have to write
+the fromPrior and perturb methods for all of these parameters, which is
+potentially very annoying.
+
 \end{document}

### 2015-12-07 Documents/manual.tex
More writing

@@
 the fromPrior and perturb methods for all of these parameters, which is
-potentially very annoying.
+potentially very annoying. You'd have to decide such things as how to propose
+to change $N$, how frequently to propose a move for one or a few of the
+amplitudes, periods, and phases, and the nature of those moves. This is even
+more complicated when you want a heirarchical prior for the amplitudes, periods,
+and phases.
+
+RJObject provides sensible default moves, drastically reducing the amount of
+work required to write the model class.
 

### 2015-12-07 Documents/manual.tex
Using SineWaves example

@@
 RJObject provides sensible default moves, drastically reducing the amount of
-work required to write the model class.
+work required to write the model class. In the Examples/SineWaves directory,
+the model is specified in the MyModel class. If you open MyModel.h you'll
+notice the MyModel class doesn't have very many data members:
+
+\begin{framed}
+\begin{verbatim}
+class MyModel:public DNest3::Model
+{
+    private:
+        RJObject<MyDistribution> objects;
+        double sigma;
+
+        std::vector<long double> mu;
+}
+\end{verbatim}
+\end{framed}
+
+The variable {\tt objects}, of type {\tt RJObject<MyDistribution>}, represents
+$\left\{N, \{A_i, T_i, \phi_i\}_{i=1}^N\right\}$. The only other parameter
+is the noise standard deviation $\sigma$. The vector of long doubles is
+the ``mock signal'' used in the likelihood function.
 

### 2015-12-07 Documents/manual.tex
Refer to paper

@@
 RJObject is a C++ library for implementing certain kinds of trans-dimensional
-hierarchical models (basically `mixture models')in DNest3. I developed this
+hierarchical models (basically `mixture models') in DNest3. I developed this
 library after realising that many models I was implementing all had this
@@
 Throughout this manual, I will assume you're proficient at
-DNest3. If not, please read the manual for that project first.
+DNest3. If not, please read the manual for that project first and get some
+experience using the software. I'll also assume you've read the RJObject
+paper (available at {\tt http://arxiv.org/abs/1411.3921}).
 

### 2015-12-08 Documents/manual.tex
formatting

@@
 \section{Introduction}
-RJObject is a C++ library for implementing certain kinds of trans-dimensional
-hierarchical models (basically `mixture models') in DNest3. I developed this
+\rjobject~is a C++ library for implementing certain kinds of trans-dimensional
+hierarchical models (basically `mixture models') in \dnest. I developed this
 library after realising that many models I was implementing all had this
@@
 Throughout this manual, I will assume you're proficient at
-DNest3. If not, please read the manual for that project first and get some
-experience using the software. I'll also assume you've read the RJObject
-paper (available at {\tt http://arxiv.org/abs/1411.3921}).
+\dnest. If not, please read the manual for that project first and get some
+experience using the software. I'll also assume you've read the \rjobject~paper
+(available at {\tt http://arxiv.org/abs/1411.3921}).
 
 \section{Dependencies}
-As well as all the DNest3 dependencies, it's best if you add some of DNest3's
-directories to your various PATH variables. Specifically, things will work best
-if you put the location of the DNest3 libraries (the location of
+As well as \dnest~and its dependencies, it's best if you add some of
+\dnest's directories to your various PATH variables.
+Specifically, things will work best
+if you put the location of the \dnest~library (the location of
 the {\tt libdnest3.so} file) to your {\tt LIBRARY\_PATH} and
-{\tt LD\_LIBRARY\_PATH}. Also add the directory containing the DNest3 header
+{\tt LD\_LIBRARY\_PATH}. Also add the directory containing the \dnest~header
 files ({\tt *.h}) to your {\tt CPLUS\_INCLUDE\_PATH}. Finally, add the
-directory containing DNest3's {\tt postprocess.py} and {\tt dnest\_plots.py} to
+directory containing \dnest's {\tt postprocess.py} and {\tt dnest\_plots.py} to
 your {\tt PYTHON\_PATH}.
@@
 If you have all the dependencies, and have all the paths set up correctly as
-specified above, you should be able to compile RJObject by going into the
-root RJObject directory and invoking {\tt make}. Alternatively you can follow
+specified above, you should be able to compile \rjobject~by going into the
+root \rjobject~directory and invoking {\tt make}. Alternatively you can follow
 the {\tt cmake} instructions in the README.
@@
 \section{Specifying models}
-When using RJObject, you're really using DNest3. So to implement a model, you
-need to write a DNest3 model class and specify the fromPrior, perturb,
-logLikelihood methods etc. All RJObject provides is a quick and relatively
+When using \rjobject, you're really using \dnest. So to implement a model, you
+need to write a \dnest~model class and specify the fromPrior, perturb,
+logLikelihood methods etc. All \rjobject provides is a quick and relatively
 painless way of incorporating a trans-dimensional object into your model and
 specifying heirarchical priors for its parameters. For example, consider
-the SineWave example from the RJObject paper. There are an unknown number of
+the SineWave example from the \rjobject paper. There are an unknown number of
 sinusoids $N$, and each of the sinusoids has an unknown period $T_i$, amplitude
-$A_i$, and phase $\phi_i$. To use DNest3 on this problem, you'd have to write
+$A_i$, and phase $\phi_i$. To use \dnest~on this problem, you'd have to write
 the fromPrior and perturb methods for all of these parameters, which is
@@
 
-RJObject provides sensible default moves, drastically reducing the amount of
+\rjobject provides sensible default moves, drastically reducing the amount of
 work required to write the model class. In the Examples/SineWaves directory,
